---
title: "Industry-Health"
format: html
editor: source
---

```{r setup, include=FALSE}

install.packages("vtable")
install.packages("dplyr")
install.packages("tidyverse")
install.packages("ggplot2")
install.packages("stringr")
install.packages("spdep")
install.packages("tidycensus")
install.packages("tigris")
install.packages("viridis")
install.packages("cowplot")
install.packages("estimatr")
install.packages("huxtable")
#install.packages("margins")
install.packages(c("lmtest", "sandwich"))
install.packages("spatialreg")



library(vtable)
library(dplyr)
library(tidyverse)
library(ggplot2)
library(spdep)
library(tidycensus)
library(tigris)
library(viridis)
library(cowplot)
library(lmtest)
library(sandwich)
library(estimatr)
library(huxtable)
#library(margins)
library(spatialreg)
library(logger)
library(moments)
library(PerformanceAnalytics)
library(correlation)
library(estimatr)


# scripts
source("C:/Users/saraa/rurec/projects/industry_health/dataprep.R", local = (ih_dp <- new.env()))
file.exists("projects/industry_health/dataprep.R")
getwd()
file.exists("C:/Users/saraa/rurec/projects/industry_health/dataprep.R")
source("datasets/circularity/circularity.R", local = (circ <- new.env()))

# chunk behavior 
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 

# nonscientific notation
options(scipen=999)

# log output 
log_threshold(DEBUG)



##Load Hard Coded Data Sets 

##Median Age of County 2012
age = get_acs(geography = "county",
                   survey = "acs5",
                   variables = c("S0101_C01_030E"
                   ),
                   year = params$year) %>%
  select("place"=GEOID, "age"=estimate)



##ERS County Typologies 

ers = read.csv("https://www.ers.usda.gov/webdocs/DataFiles/48652/2015CountyTypologyCodes.csv?v=8831.8") %>%
  mutate(place = sprintf("%05d", FIPStxt)) %>%
  select(place, "ers_ag"=Farming_2015_Update, "ers_mine"=Mining_2015.Update, "ers_mfg"=Manufacturing_2015_Update, "ers_gov"=Government_2015_Update, "ers_rec"=Recreation_2015_Update, "ers_nonsp"=Nonspecialized_2015_Update)



##Industry Structure

industry = get_acs(geography = "county",
                   survey = "acs5",
                   variables = c("DP03_0032E","DP03_0033E","DP03_0034E"
                                        ,"DP03_0035E","DP03_0036E","DP03_0037E"
                                        ,"DP03_0038E","DP03_0039E","DP03_0040E"
                                        ,"DP03_0041E","DP03_0042E","DP03_0043E"
                                        ,"DP03_0044E","DP03_0045E"
                   ),
                   year = params$year) %>%
  select("place"=GEOID, variable, estimate) %>%
  pivot_wider(names_from = variable, values_from = estimate) %>%
  mutate(pct_afm = DP03_0033/DP03_0032, pct_mfg = DP03_0034/DP03_0032, pct_fire = DP03_0039/DP03_0032, pct_edmed = DP03_0041/DP03_0032, pct_rec = DP03_0042/DP03_0032, pct_gov = DP03_0044/DP03_0032) %>%
  select(place, pct_afm, pct_mfg, pct_fire, pct_edmed, pct_rec, pct_gov)


```

```{r zip pack, include = FALSE}

# uncomment and pack/unpack as necessary
# dp$zip_pack("tmp/eca_paa_data_v240324.zip", dp$opath)
# dp$zip_unpack("tmp/eca_paa_data_v240324.zip", overwrite = TRUE)

```



# data

```{r data cashe, include = FALSE}

# uncomment as necessary
# dp$create_complete_cache()

```

### Premature Death Change over time (2012-2017)

```{r}

##Construct typology of premature death change over time between 2012-2017 

#Calculate change over time

# Construct typology E.g. 4 types:
    
# High premature death in t0 & stable/reduction in premature death from t-1 to t0
    #   E.g. maybe southeast?
# High premature death in t0 & material increase in premature death from t-1 to t0
    # E.g. maybe Appalachia?
# Low premature death in t0 & stable in premature death from t-1 to t0
# Low premature death in t0 & material increase in premature death from t-1 to t0

```

### Access to Healthcare

```{r}
# Access to health care----------

# Normalize the data
# data$Uninsured_Ratio <- data$Uninsured_Percentage / 100
# data$Primary_Care_Ratio <- data$Population / data$Primary_Care_Physicians
# data$Dentists_Ratio <- data$Population / data$Dentists
# data$Mental_Health_Ratio <- data$Population / data$Mental_Health_Providers
# 
# # Create a composite score for each county
# data$Healthcare_Access_Score <- (1 - data$Uninsured_Ratio) * (1 - data$Primary_Care_Ratio / max(data$Primary_Care_Ratio)) *
#   (1 - data$Dentists_Ratio / max(data$Dentists_Ratio)) * (1 - data$Mental_Health_Ratio / max(data$Mental_Health_Ratio))

```

\`\`\`

## Summary Statistics

### Descriptive Statistics

### ESDA

### Density Plots

### Maps

## Analysis - Multinomial Logit Model

For Statistics

### Model 1-

## Analysis- Exploratory Regressions

Factors

Trade

Autonomy

etention

Supply chain value added

\% AFM % value added

\% worker share of value added

\% worker share of output

Industrial structure

ERS dependency typologies

\% composition (output)

\% composition (employment)

Prosperity

Poverty,

education,

income,

etc.

Covariates

Access to healthcare,

workforce,

population,

age

## Analysis- Premature Death

Death = trade Death = trade + covariates

Death = trade + industrial + covariates

Death = trade + supply chain + covariates

Death = trade + industry + supply chain + covariates

Death = prosperity + covariates

Death = prosperity + industrial + covariates

Death = trade + prosperity + covariates

Death = trade + prosperity + industrial + covariates

Death = trade + prosperity + industrial + supply chain + covariates

Specifications

Error

Error + interactions

Error + regimes

Pat

## Analysis- Spatial Regime

### 