---
title: "isolation"
format:
  html:
    date: today
    toc: true
    toc-depth: 2
    code-fold: true
    code-overflow: wrap
    embed-resources: true
    df-print: paged
editor: 
  markdown: 
    wrap: 72
---

# Header

```{r}
library(tidyverse)
library(ggridges)
library(tidycensus)
library(tigris)
library(tmap)
library(ipumsr)
library(economiccomplexity)
library(scales)
library(migraph)
library(igraph)
library(scoredec)
library(brainGraph)
library(sf)
library(purrr)
library(spatstat)
library(boot)
library(furrr)
library(purrr)
library(viridis)

source("R/basic_utilities.R", local = (util <- new.env()))
source("R/geography.R", local = (geography <- new.env()))
source("R/place_io.R", local = (place_io <- new.env()))
source("R/circularity.R", local = (circularity <- new.env()))
source("R/trade_flows_wip.R", local = (trade_flows <- new.env()))


input_ptile = 0.50

# paths within project subfolder
ppath <- function(...) {
  file.path("projects/isolation", ...)
}

# input data files
ipath <- list(
  dist = str_glue(geography$opath$dist_mat_, from = "center", year = 2013, cbsa = FALSE),
  outsupdem = str_glue(place_io$opath$outsupdem_, year = 2012, ilevel = "det", bus_data = "cbp_imp"),
  tf = str_glue(trade_flows$opath$flows_, bus_data = "cbp_imp", ilevel = "det", year = 2012, ind_code = "all_industries"),
  tf17 = str_glue(trade_flows$opath$flows_, bus_data = "cbp_imp", ilevel = "det", year = 2017, ind_code = "all_industries"),
  tf_com = str_glue(trade_flows$opath$flows_all_, bus_data = "cbp_imp", ilevel = "det", year = 2012)
)

# output data files
opath <- list(
  
)

# pack all data files
# util$zip_pack(ppath("data/datacache_2025-05-28.zip"), files = c(ipath, opath))
# unpack
util$zip_unpack(ppath("data/datacache_2025-05-28.zip"))

```

# Data

```{r}
# county-to-county centroid distance in miles
df_dist <- local({
  # convert from matrix in meters to unitless dataframe in miles
  x <- readRDS(ipath$dist)
  x1 <- x %>%
    units::set_units("mi") %>%
    units::set_units(NULL)
  dimnames(x1) <- dimnames(x)
  x1 %>%
    as_tibble(rownames = "from") %>%
    pivot_longer(!from, names_to = "to", values_to = "distance")
})

# county-commodity output, supply and demand
df_outsupdem <- arrow::read_parquet(ipath$outsupdem) %>% 
    mutate(exsup = pmax(0, supply - demand), exdem = pmax(0, demand - supply))


# county circularity indicators, depends on outsupdem data file
df_circ <- circularity$circularity_indicators(year = 2012, bus_data = "cbp_imp")

# county-to-county trade flows across all commodities
df_tf <- readRDS(ipath$tf) %>%
  as_tibble(rownames = "from") %>%
  pivot_longer(!from, names_to = "to", values_to = "flow") %>%
  filter(flow > 0)

# county-to-county trade flows by commodity
df_tf_com <- arrow::read_parquet(ipath$tf_com)

df_dist %>% head()
df_outsupdem %>% head()
df_circ %>% head()
df_tf %>% head()
df_tf_com %>% head()

```

```{r}

census_map <- get_acs(
  geography = "county",
  variables = "B01002_001",
  year = 2019,
  survey = "acs5",
  geometry = TRUE,
  resolution = "20m"
) %>%
  shift_geometry() %>%
  dplyr::select(place="GEOID")
states  <- get_acs(
  geography = "state",
  variables = "B01002_001",
  year = 2019,
  survey = "acs5",
  geometry = TRUE,
  resolution = "20m"
) %>%
  shift_geometry() %>%
  dplyr::select(state="GEOID")

```

## poverty

#### puma-level

```{r}
inf_clean = read.csv("projects/isolation/data/raw/inflation.csv")

inf23 = inf_clean %>%
  filter(Year == 2023)
inf23 = inf23$inflation

inf = inf_clean %>% 
  mutate(inf23 = inf23,
         deflate = inflation/inf23) %>% 
  dplyr::select("year"=Year, deflate)

```

```{r}
cola_clean = read.csv("projects/isolation/data/raw/cola2.csv") %>% 
  dplyr::select(-X) %>% 
  mutate(FIPS = sprintf("%05d", FIPS))

```

```{r}
xwalk12 = read.csv("projects/isolation/data/raw/county_puma12.csv") %>% 
  filter(state != "State code") %>% 
  mutate(FIPS = sprintf("%05d", as.numeric(county)),
         puma12 = sprintf("%05d", as.numeric(puma12)),
         state = sprintf("%02d", as.numeric(state)),
         PUMA = str_c(state, puma12), afact = as.numeric(afact)) %>% 
  dplyr::select(PUMA, FIPS, afact)
xwalk22 = read.csv("projects/isolation/data/raw/county_puma22.csv") %>% 
  filter(state != "State code") %>% 
  mutate(county = sprintf("%05d", as.numeric(county)),
         puma22 = sprintf("%05d", as.numeric(puma22)),
         state = sprintf("%02d", as.numeric(state)),
         PUMA = str_c(state, puma22), afact = as.numeric(afact)) %>% 
  dplyr::select(PUMA, "FIPS"=county, afact)
```

```{r}
ddi <- read_ipums_ddi("projects/isolation/data/raw/usa_00048.xml")
pums_clean <- read_ipums_micro(ddi, data_file = "projects/isolation/data/raw/usa_00048.dat.gz")

```

```{r}
puma1222 = read.csv("projects/isolation/data/raw/puma2010-to-puma2020.csv") %>% 
  filter(state != "State code") %>% 
  mutate(puma12 = sprintf("%05d", as.numeric(puma12)),
         puma22 = sprintf("%05d", as.numeric(puma22)),
         state = sprintf("%02d", as.numeric(state)),
         puma12 = str_c(state, puma12),
         puma22 = str_c(state, puma22)
         ) %>% 
  dplyr::select(puma12, puma22, afact) %>% 
  distinct() %>% 
  mutate(afact = as.numeric(afact))
```

```{r}

pums12 = pums_clean %>% 
  filter(FTOTINC < 9999993 & PERNUM == 1) %>% 
  mutate(PUMA = str_c(sprintf("%02d", STATEFIP), sprintf("%05d", PUMA)),
         NADULT = ifelse(MARST == 1, 2, 1),
         fsizeadj = ifelse(NCHILD == 0, NADULT^0.5,
                    ifelse(NCHILD > 0 & NADULT == 1, (NADULT+0.8*1+0.5*(NCHILD-1))^0.7,
                    ifelse(NCHILD > 0 & NADULT > 1, (NADULT+0.5*NCHILD)^0.7, NA))),
         mar = ifelse(MARST == 1 | MARST == 2, "2p", "1p"),
         ch = ifelse(NCHILD == 0, "0c",
              ifelse(NCHILD == 1, "1c",
                     ifelse(NCHILD == 2, "2c",
                     ifelse(NCHILD == 3, "3c",
                     ifelse(NCHILD > 3, "4c", NA
                     ))))),
         fsize = str_c(mar, ch)) %>% 
  dplyr::select(PUMA, fsize, FTOTINC, "year"=YEAR, PERWT, fsizeadj) %>% 
  left_join(y=xwalk12, by="PUMA") %>% 
  left_join(y=inf, by="year") %>% 
  left_join(y=cola_clean, by=c("FIPS","fsize","year")) %>% 
  mutate(wgt = PERWT*afact,
         cola = FTOTINC/col/fsizeadj/deflate) %>% 
  ungroup() %>% 
  mutate(usavg = weighted.median(x=cola, w=wgt, na.rm=TRUE),
         ccapov = ifelse(cola < (0.60*usavg), 1,
                         ifelse(cola >= (0.60*usavg), 0, NA))) %>%
  group_by("puma12"=PUMA) %>% 
  dplyr::summarize(ccapov = weighted.mean(x=ccapov, w=wgt, na.rm=TRUE)) %>% 
  mutate(ccapov = ifelse(ccapov == "NaN", NA, ccapov)) %>% 
  mutate(year = 2012)


pums22 = pums_clean %>% 
  filter(FTOTINC < 9999993 & PERNUM == 1) %>% 
  mutate(PUMA = str_c(sprintf("%02d", STATEFIP), sprintf("%05d", PUMA)),
         NADULT = ifelse(MARST == 1, 2, 1),
         fsizeadj = ifelse(NCHILD == 0, NADULT^0.5,
                    ifelse(NCHILD > 0 & NADULT == 1, (NADULT+0.8*1+0.5*(NCHILD-1))^0.7,
                    ifelse(NCHILD > 0 & NADULT > 1, (NADULT+0.5*NCHILD)^0.7, NA))),
         mar = ifelse(MARST == 1 | MARST == 2, "2p", "1p"),
         ch = ifelse(NCHILD == 0, "0c",
              ifelse(NCHILD == 1, "1c",
                     ifelse(NCHILD == 2, "2c",
                     ifelse(NCHILD == 3, "3c",
                     ifelse(NCHILD > 3, "4c", NA
                     ))))),
         fsize = str_c(mar, ch)) 
pums22 = pums22 %>% 
  dplyr::select(PUMA, fsize, FTOTINC, "year"=YEAR, PERWT, fsizeadj) %>% 
  left_join(y=xwalk22, by="PUMA") %>% 
  left_join(y=inf, by="year") %>% 
  left_join(y=cola_clean, by=c("FIPS","fsize","year")) %>% 
  mutate(wgt = PERWT*afact,
         cola = FTOTINC/col/fsizeadj/deflate) %>% 
  ungroup() %>% 
  mutate(usavg = weighted.median(x=cola, w=wgt, na.rm=TRUE),
         ccapov = ifelse(cola < (0.60*usavg), 1,
                         ifelse(cola >= (0.60*usavg), 0, NA))) %>%
  group_by("puma22"=PUMA) %>% 
  dplyr::summarize(ccapov = weighted.mean(x=ccapov, w=wgt, na.rm=TRUE)) %>% 
  left_join(y=puma1222, by="puma22") %>% 
  group_by(puma12) %>% 
  mutate(ccapov = ifelse(ccapov == "NaN", NA, ccapov)) %>% 
  dplyr::summarize(ccapov = weighted.mean(x=ccapov, w=afact, na.rm=TRUE)) %>% 
  mutate(year = 2022)

```

```{r}

ccapov = rbind(pums12, pums22) %>% 
  mutate(year = str_c("ccapov",year)) %>% 
  pivot_wider(names_from = year, values_from = ccapov) %>% 
  mutate(filt = ifelse(is.na(ccapov2012) & is.na(ccapov2022), "drop", "keep")) %>% 
  filter(filt != "drop") %>% 
  dplyr::select(-filt)


```

```{r}
# write.csv(ccapov, "ccapov.csv")
ccapov = read.csv("ccapov.csv")
```

#### tract-level

take tract10-puma12 xwalk, allocate ccapov to tracts based on tract:puma
poverty ratio

-   seems a little complicated, since not only have partial tracts
    within pumas but they have different poverty ratios

```{r}
puma1222 = read.csv("projects/isolation/data/raw/puma2010-to-puma2020.csv") %>% 
  filter(state != "State code") %>% 
  mutate(puma12 = sprintf("%05d", as.numeric(puma12)),
         puma22 = sprintf("%05d", as.numeric(puma22)),
         state = sprintf("%02d", as.numeric(state)),
         puma12 = str_c(state, puma12),
         puma22 = str_c(state, puma22)
         ) %>% 
  dplyr::select(puma12, puma22, afact) %>% 
  distinct() %>% 
  mutate(afact = as.numeric(afact))
```

```{r}
xwalk_tpuma = read.csv("projects/isolation/data/raw/2010_Census_Tract_to_2010_PUMA.txt") %>% 
  mutate(puma12 = str_c(sprintf("%02d", STATEFP), sprintf("%05d", PUMA5CE)),
         tract12 = str_c(sprintf("%02d", STATEFP), sprintf("%03d", COUNTYFP), 
                         sprintf("%06d", TRACTCE))
         ) %>% 
  dplyr::select(puma12, tract12) %>% 
  distinct()
```

```{r}
nhgis_clean = read.csv("projects/isolation/data/raw/nhgis0031_ts_nominal_tract.csv")
nhgis2_clean = read.csv("projects/isolation/data/raw/nhgis0032_ds192_20125_puma.csv")
nhgis3_clean = read.csv("projects/isolation/data/raw/nhgis0032_ds268_20235_puma.csv")

```

```{r}
xwalk_tract = read.csv("projects/isolation/data/raw/tract1222.csv") %>% 
  mutate(tract12 = str_pad(GEOID_TRACT_10, width = 11, pad = "0", side = "left"),
         tract22 = str_pad(GEOID_TRACT_20, width = 11, pad = "0", side = "left"))
```

```{r}
altwalk_tract = read.csv("projects/isolation/data/raw/xwalk_tract.csv") %>% 
  mutate(tract12 = str_pad(TRTID2010, width = 11, pad = "0", side = "left"),
         tract22 = str_pad(TRTID2020, width = 11, pad = "0", side = "left")) %>% 
  dplyr::select(tract12, tract22, "afact"=TR20_TO_TR10_WEIGHT) %>% 
  distinct()
```

```{r}

```

```{r}
nhgis_tract = nhgis_clean %>% 
  mutate(tract12 = str_c(sprintf("%02d", STATEFP), sprintf("%03d", COUNTYFP), 
                         sprintf("%06d", TRACTA))) %>% 
  dplyr::select(tract12, 
                YEAR,
                "inlf_2529"=BT7AU,
                "outlf_2529"=BT7AX,
                "emp_2529"=BT7AV, 
                "inlf_3034"=BT7BA,
                "outlf_3034"=BT7BD,
                "emp_3034"=BT7BB,
                "inlf_3544"=BT7BG,
                "outlf_3544"=BT7BJ,
                "emp_3544"=BT7BH,
                "inlf_4554"=BT7BM,
                "outlf_4554"=BT7BP,
                "emp_4554"=BT7BN,
                "pov"=CL6AA,
                "povpop"=AX6AA
                ) %>% 
  mutate(pop = inlf_2529+outlf_2529+inlf_3034+outlf_3034+inlf_3544+outlf_3544+inlf_4554+outlf_4554,
         emp = emp_2529+emp_3034+emp_3544+emp_4554,
         emprt = emp/pop,
         povrt = pov/povpop,
         year = ifelse(YEAR == "2008-2012", 2012,
                       ifelse(YEAR == "2019-2023", 2023, NA))) %>% 
  dplyr::select(tract12, year, emprt, povrt)
```

```{r}
nhgis_tract12 = nhgis_tract %>% 
  filter(year == 2012)

nhgis_tract23 = nhgis_tract %>% 
  filter(year == 2023) %>% 
  rename("tract22"="tract12") %>% 
  left_join(y=altwalk_tract, by="tract22") %>% 
  group_by(tract12, year) %>% 
  dplyr::summarize(emprt = weighted.mean(x=emprt, w=afact, na.rm=TRUE),
                   povrt = weighted.mean(x=povrt, w=afact, na.rm=TRUE),
                   )

nhgis_tract_alt = rbind(nhgis_tract12, nhgis_tract23) %>% 
  mutate(povrt = ifelse(povrt == "NaN", NA, povrt),
         emprt = ifelse(emprt == "NaN", NA, emprt)
         )

```

```{r}
nhgis_puma12 = nhgis2_clean %>% 
  mutate(povrt = RAZE002/RAZE001) %>% 
  mutate(puma12 = str_c(sprintf("%02d", STATEA), 
                         sprintf("%05d", PUMAA))) %>% 
  dplyr::select(puma12, povrt) %>% 
  mutate(year = 2012)

nhgis_puma22 = nhgis3_clean %>% 
  mutate(povrt = AS7TE002/AS7TE001) %>% 
  mutate(puma22 = str_c(sprintf("%02d", STATEA), 
                         sprintf("%05d", PUMAA))) %>% 
  dplyr::select(puma22, "year"=YEAR, povrt) %>% 
  left_join(y=puma1222, by="puma22") %>% 
  group_by(puma12) %>% 
  dplyr::summarize(povrt = weighted.mean(x=povrt, w=afact, na.rm=TRUE)) %>% 
  mutate(year = 2023)

nhgis_puma = rbind(nhgis_puma12, nhgis_puma22)
```

```{r}
nhgis = xwalk_tpuma %>% 
  left_join(y=nhgis_puma, by=c("puma12")) %>% 
  rename("pov_puma"="povrt") %>% 
  left_join(y=nhgis_tract_alt, by=c("tract12","year")) %>% 
  rename("pov_tract"="povrt") %>% 
  mutate(tpratio = pov_tract/pov_puma)
```

```{r}

temp = nhgis %>% 
  dplyr::select(tract12, year, emprt)

socex = nhgis %>% 
  dplyr::select(puma12, tract12, year, tpratio) %>% 
  drop_na(year) %>% 
  pivot_wider(names_from = year, values_from = tpratio) %>% 
  rename("tpratio_2012"="2012",
         "tpratio_2023"="2023") %>% 
  left_join(y=ccapov, by=c("puma12")) %>% 
  mutate(pov_2012 = ccapov2012*tpratio_2012,
         pov_2023 = ccapov2022*tpratio_2023
         ) %>% 
  dplyr::select(tract12, pov_2012, pov_2023) %>% 
  pivot_longer(cols=!c(tract12), names_to = "year", values_to = "pov") %>% 
  mutate(year = as.numeric(substr(year, 5, 8))) %>% 
  left_join(y=temp, by=c("tract12","year")) %>% 
  pivot_wider(names_from = year, values_from = c(pov, emprt)) %>% 
  rename("tract"="tract12") %>% 
  mutate(povch = (pov_2023-pov_2012)/pov_2012,
         emprtch = (emprt_2023-emprt_2012)/emprt_2012,
         povch = ifelse(povch > 2, 2,
                        ifelse(povch == "Inf", NA,
                        ifelse(povch == "NaN", NA,
                        povch))),
         emprtch = ifelse(emprtch > 1, 1,
                        ifelse(emprtch == "Inf", NA,
                        ifelse(emprtch == "NaN", NA,
                        emprtch))))
```

```{r}
test = nhgis %>% dplyr::select(tract12, year, tpratio) %>%
  drop_na() %>% 
  pivot_wider(names_from=year, values_from=tpratio) %>% 
  rename("a2012"="2012", "a2023"="2023") %>% 
  mutate(test = a2023-a2012,
         test = ifelse(test > 1, 1,
                       ifelse(test < -1, -1, test)))

hist(test$test)

nhgis %>% 
  drop_na() %>% 
  group_by(year) %>% 
  dplyr::summarize(d = n())

```

# core-periphery

```{r}
# Create a sample trade flow data frame
trade_data <- data.frame(
  from = c("USA", "USA", "USA", "CAN", "CAN", "MEX"),
  to = c("CAN", "MEX", "BRA", "MEX", "USA", "USA"),
  value = c(100, 150, 50, 80, 200, 75)
)
```

##### using

added log, see what that looks like

double check with anton what units are on trade volume

```{r}

df_tf_clean = df_tf
df_tf2 = df_tf_clean

df_tf = df_tf2 %>% 
  filter(flow > 100)

```

```{r}
xwalk = read.csv("C:/Users/adams/Downloads/xwalk_cz.csv") %>% 
  mutate(place = sprintf("%05d", cty)) %>% 
  dplyr::select(cz, place) %>% 
  mutate(cz = as.character(cz))
```

```{r}



temp1 = xwalk %>% 
  rename("from"="place",
         "cz_from"="cz")

temp2 = xwalk %>% 
  rename("to"="place",
         "cz_to"="cz")

df_tf4 = df_tf2 %>% 
  left_join(y=temp1, by="from") %>% 
  left_join(y=temp2, by="to") %>% 
  dplyr::select(-to, -from) %>% 
  mutate(filt = ifelse(cz_from == cz_to, "same", "diff")) %>% 
  drop_na() %>% 
  filter(filt == "diff") %>% 
  group_by(cz_from, cz_to) %>% 
  dplyr::summarize(flow = sum(flow, na.rm=TRUE)) %>% 
  rename("from"="cz_from",
         "to"="cz_to")

```

```{r}
# df = df_tf4 %>% 
#   filter(flow > 50000)
# df = df_tf4 %>%
#   filter(flow > 5000)
# df = df_tf4 %>%
#   filter(flow > 50)

df = df_tf4
vertices <- data.frame(cz = unique(c(df$from, df$to)))

# Build graph explicitly
g <- igraph::graph_from_data_frame(
  d = df %>% dplyr::select(from, to, flow),
  directed = FALSE,
  vertices = vertices
)

# Assign weights
E(g)$weight <- df$flow

# Compute weighted s-core
weighted_coreness_scores <- brainGraph::s_core(g)

# Convert to dataframe with vertex names
brain_cz <- data.frame(
  cz = igraph::V(g)$name,
  alt_core = as.numeric(weighted_coreness_scores)
) 

core_brain3 = brain_cz %>%
  left_join(xwalk, by = "cz") %>%
  dplyr::select(place, alt_core)

```

```{r}

```

## iso

```{r}
ruc = read.csv("projects/isolation/data/raw/ruralurbancodes2013.csv") %>% 
  dplyr::select("place"=FIPS, "metro"=RUCC_2013) %>% 
  mutate(metro = ifelse(metro > 3, 0,
                        ifelse(metro <= 3, 1, NA)),
         place = sprintf("%05d", place))
```

```{r}
isolation = census_map  %>% 
  left_join(y=core_brain3, by="place") %>% 
  left_join(y=ruc, by="place")

iso2 = isolation%>%
  st_drop_geometry()
  
  
```

```{r}
# write.csv(iso2, "temp_data_v2.csv")
# isolation = read.csv("temp_data_v2.csv") %>% dplyr::select(-X)
```

moran's i

```{r}

```

## geo

```{r}
ruca_clean = read.csv("projects/isolation/data/raw/usda_ruca.csv")
far_clean = read.csv("projects/isolation/data/raw/usda_far.csv")
xwalk_trzip = read.csv("projects/isolation/data/raw/zip_tract_092016.csv") %>% 
  dplyr::select(ZIP, "tract"=TRACT)

```

```{r}
far = far_clean %>% 
  dplyr::select(ZIP, far1) %>% 
  distinct() %>% 
  left_join(y=xwalk_trzip, by="ZIP") %>% 
  group_by(tract) %>% 
  dplyr::summarize(far = sum(far1, na.rm=TRUE)) %>% 
  mutate(far = ifelse(far >= 1, 1,
                      ifelse(far == 0, 0, NA)),
         tract = str_pad(tract, width = 11, pad = "0", side = "left"))
```

```{r}
ruca = ruca_clean %>% 
  dplyr::select(tract, "ruca"=primary) %>% 
  mutate(noncore = ifelse(ruca == 1, 0,
                        ifelse(ruca == 99, NA,
                        ifelse(ruca > 1, 1, NA))),
         tract = str_pad(tract, width = 11, pad = "0", side = "left")) %>% 
  left_join(y=far, by="tract")
```

```{r}

```

```{r}
options(tigris_use_cache = TRUE)
us_states <- states(class = "sf", progress_bar = FALSE)
state_list <- us_states$STUSPS
all_tracts_sf <- map_dfr(state_list, ~ {
  tracts(state = .x, class = "sf", progress_bar = TRUE, year = 2010)
}) %>% 
  shift_geometry()
```

```{r}
ptile = brain_cz %>%
  ungroup() %>% 
  dplyr::summarize(
    ptile = quantile(alt_core, probs = input_ptile, na.rm = TRUE)
  )

ptile = ptile$ptile
```

```{r}
geo = all_tracts_sf %>% 
  mutate(FIPS = str_c(STATEFP, COUNTYFP)) %>% 
  dplyr::select("tract"=GEOID10, FIPS) %>%
  left_join(y=ruca, by="tract") %>% 
  mutate(geoiso = ifelse(far == 1 & noncore == 1, "iso",
                         ifelse(far == 1 & noncore == 0, "core far",
                                ifelse(far == 0 & noncore == 1, "noncore nonfar",
                                       ifelse(far == 0 & noncore == 0, "core nonfar",
                                       NA)))),
         st = as.numeric(substr(FIPS, 1, 2))) %>%
  filter(st <= 56)

temp = iso2 %>% 
  dplyr::select("FIPS"=place, alt_core, metro)  %>%
  mutate(alt_core = ifelse(is.na(alt_core), 0, alt_core),
         econiso = ifelse(alt_core >= ptile, 0,
                       ifelse(alt_core < ptile, 1, NA)))

data = geo %>% 
  left_join(y=temp, by="FIPS") %>% 
  left_join(y=socex, by="tract")

temp = data %>%
  st_drop_geometry() %>% 
  filter(noncore == 1 & far == 1)

test = data %>%
  st_drop_geometry() %>% 
  filter(noncore == 1 & far == 1) %>%
  mutate(d = 1) %>% 
  group_by(econiso) %>% 
  dplyr::summarize(d = sum(d, na.rm=TRUE))
```

# data-clean

```{r}
saveRDS(data, "data.rds")
# data <- readRDS("data.rds")

states  <- get_acs(
  geography = "state",
  variables = "B01002_001",
  year = 2019,
  survey = "acs5",
  geometry = TRUE,
  resolution = "20m"
) %>%
  shift_geometry() %>%
  dplyr::select(state="GEOID")
```

# descriptives

## scatter

```{r}

ggplot(data = t2, aes(x = alt_core, y = pov_2023, color = factor(far, levels = c(1, 0)))) +
  geom_point(alpha = 0.7, size = 1.7) +
  scale_color_manual(
    values = c("0" = "#cfe2f3", "1" = "#008080"),
    labels = c("0" = "non geo iso", "1" = "geo iso"),
    name = "Far Status"
  ) +
  labs(
    title = "Poverty by Economic Coreness",
    x = "Economic coreness",
    y = "Poverty Rate (2023)"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 14, face = "bold"),
    axis.title.x = element_text(size = 12, face = "bold"),
    axis.title.y = element_text(size = 12, face = "bold"),
    legend.title = element_text(size = 12, face = "bold"),
    legend.text = element_text(size = 10, face = "italic"),
    legend.position = "right"
  )

ggplot(data = t2, aes(x = alt_core, y = emprt_2023, color = factor(far, levels = c(1, 0)))) +
  geom_point(alpha = 0.7, size = 1.7) +
  scale_color_manual(
    values = c("0" = "#cfe2f3", "1" = "#008080"),
    labels = c("0" = "non geo iso", "1" = "geo iso"),
    name = "Far Status"
  ) +
  labs(
    title = "Poverty by Economic Coreness",
    x = "Economic coreness",
    y = "Employment Rate (2023)"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 14, face = "bold"),
    axis.title.x = element_text(size = 12, face = "bold"),
    axis.title.y = element_text(size = 12, face = "bold"),
    legend.title = element_text(size = 12, face = "bold"),
    legend.text = element_text(size = 10, face = "italic"),
    legend.position = "right"
  )


```
```{r}
t2_clean <- t2 %>%
  filter(is.finite(alt_core), is.finite(pov_2023)) %>%
  mutate(
    alt_core = as.numeric(alt_core),
    pov_2023 = as.numeric(pov_2023),
    # Properly recode far into a labeled factor
    far = factor(far, levels = c(1, 0), labels = c("geo iso", "non geo iso"))
  )

ggplot(t2_clean, aes(x = alt_core, y = pov_2023, color = far)) +
  geom_smooth(method = "gam", formula = y ~ s(x), se = TRUE, linewidth = 1.2) +
  scale_color_manual(
    values = c("geo iso" = "#008080", "non geo iso" = "#cfe2f3"),
    name = "Far Status"
  ) +
  labs(
    title = "Smoothed Poverty Trends by Economic Coreness",
    x = "Economic Coreness (alt_core)",
    y = "Poverty Rate (2023)"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 14, face = "bold"),
    axis.title.x = element_text(size = 12, face = "bold"),
    axis.title.y = element_text(size = 12, face = "bold"),
    legend.title = element_text(size = 12, face = "bold"),
    legend.text = element_text(size = 10, face = "italic"),
    legend.position = "right"
  )
```



## geo map
```{r}
data %>%
  drop_na(geoiso) %>%
  mutate(geoiso = case_when(
    geoiso == "core nonfar" ~ "metro core",
    geoiso == "noncore nonfar" ~ "noncore",
    geoiso == "iso" ~ "iso"
  ),
  geoiso = factor(geoiso, levels = c("metro core", "noncore", "iso"))) %>%
  tm_shape() +
  tm_polygons(
    col = "geoiso",
    border.alpha = 0,
    style = "cat",
    palette = c("metro core" = "white", "noncore" = "#BEE8CA", "iso" = "#3F498E"),
    title = "geo iso?"
  ) +
  tm_shape(states) +
  tm_borders(lwd = 0.6) +
  tm_layout(
    frame = FALSE,
    legend.position = c("LEFT", "top"),
    legend.frame = FALSE,
    legend.bg.alpha = 0
  )


```


```{r}
isolation %>%
  mutate(alt_core = ifelse(place == "12086", 681, alt_core),
         econiso = ifelse(alt_core >= ptile, 0,
                          ifelse(alt_core < ptile, 1, NA))) %>% 
  tm_shape() +
  tm_polygons(col = "econiso", border.alpha = 0, style = "cat",
              palette = c("#BEE8CA","#3F498E"),
              title = "econ iso?") +
  tm_shape(states) +
  tm_borders(lwd = 0.6) +
  tm_layout(frame = FALSE,                # removes box around map
            legend.position = c("RIGHT", "bottom"),  # bottom-right
            legend.frame = FALSE,              # removes legend box
    legend.bg.alpha = 0)  # legend in bottom right

isolation %>%
  mutate(alt_core = ifelse(place == "12086", 681, alt_core)) %>% 
  tm_shape() +
  tm_polygons(col = "alt_core", midpoint = 200, border.alpha = 0, style = "cont",
              palette = viridis(100, option = "mako", direction = 1),
              title = "s-cores") +
  tm_shape(states) +
  tm_borders(lwd = 0.6) +
  tm_layout(
    title = "",
    title.size = 2,
    frame = FALSE,                  # removes box around map
    legend.position = c("LEFT", "top"),  # bottom-right
    legend.frame = FALSE,              # removes legend box
    legend.bg.alpha = 0             # removes box around legend
  )


# data %>%
#   filter(noncore == 1 & far == 1) %>% 
#   tm_shape() + 
#   tm_polygons(col = "econiso", midpoint = 0.5, border.alpha = 0.0, style = "cont", 
#               palette = c("#6FA8DC","#20124d"), 
#               legend.show = FALSE) +
#   tm_layout(title = "", title.size = 2) +
#   tm_shape(states) + 
#   tm_borders(lwd=0.6)
```

```{r}

# data %>%
#   filter(noncore == 1 & far == 1) %>% 
#   tm_shape() + 
#   tm_polygons(col = "povch", midpoint = 0, border.alpha = 0.0, style = "cont", 
#               palette = c("#d9d2e9","#351c75"), 
#               legend.show = FALSE) +
#   tm_layout(title = "", title.size = 2) +
#   tm_shape(states) + 
#   tm_borders(lwd=0.6)

# data %>%
#   mutate(povch = ifelse(noncore != 1, NA, povch)) %>% 
#   filter(!is.na(far)) %>% 
#   tm_shape() + 
#   tm_polygons(col = "povch", midpoint = 0, border.alpha = 0.0, style = "cont", 
#               palette = c("#133337","#fff2cc","#660000")) +
#   tm_layout(title = "", title.size = 2) +
#   tm_shape(states) + 
#   tm_borders(lwd=0.6)
```

```{r}


temp = data %>% 
  mutate(quad_combo = ifelse(geoiso == "iso" & econiso == 1, "geo & econ",
                      ifelse(geoiso == "iso" & econiso == 0, "geo",
                      ifelse(geoiso == "noncore nonfar" & econiso == 1, "econ (non-core metro)",
                      ifelse(geoiso == "noncore nonfar" & econiso == 0, "neither (non-core metro)",
                      ifelse(geoiso == "core nonfar" & econiso == 1, "econ (core metro)",
                      ifelse(geoiso == "core nonfar" & econiso == 0, "neither (core metro)", NA
                      
  )))))),
  quad_nonmetro = ifelse(noncore == 0, "core metro", quad_combo),
  quad_metro = ifelse(noncore == 1, "non-core metro", quad_combo)) %>% 
  drop_na(econiso, noncore, geoiso)


temp_nonmetro = temp %>% 
  mutate(quad_nonmetro = ifelse(quad_nonmetro == "econ (non-core metro)", "econ", quad_nonmetro),
         quad_nonmetro = ifelse(quad_nonmetro == "neither (non-core metro)", "neither", quad_nonmetro))
temp_metro = temp %>% 
  mutate(quad_metro = ifelse(quad_metro == "econ (core metro)", "econ", quad_metro),
         quad_metro = ifelse(quad_metro == "neither (core metro)", "neither", quad_metro))
         

  quad_nonmetro_colors <- c(
    "geo & econ"   = "#FFEEAD",
    "geo"  = "#52BF90",
    "econ"  = "#20124d",
    "core metro"  = "white",
    "neither" = "#6FA8DC"
  )

    quad_metro_colors <- c(
    "econ"  = "#20124d",
    "non-core metro"  = "white",
    "non" = "#6FA8DC"
  )


  
tmap_mode(mode = "plot")

  # Map
  tm_shape(temp_nonmetro) + 
    tm_polygons(
      col = "quad_nonmetro",
      border.alpha = 0.0,
      style = "cat",
      palette = quad_nonmetro_colors,
      legend.show = FALSE
    ) +
    tm_layout(
      title = "",
      title.size = 2
    ) +
    tm_shape(states) + 
    tm_borders(lwd = 0.6)
```

##### moran's i

```{r}
temp = xwalk %>% 
  rename("FIPS"="place")

czdata = data %>% 
  left_join(y=temp, by="FIPS") %>% 
  group_by(cz) %>% 
  dplyr::summarize(econiso = mean(econiso, na.rm=TRUE))
```

```{r}
temp = czdata

nb <- poly2nb(temp, queen = TRUE)
listw <- nb2listw(nb, style = "W", zero.policy = TRUE)

moran_test_result <- moran.test(temp$econiso, listw)
print(moran_test_result)
```

```{r}
temp = data %>% 
  dplyr::select(tract, far) %>% 
  drop_na()

nb <- poly2nb(temp, queen = TRUE)
listw <- nb2listw(nb, style = "W", zero.policy = TRUE)

moran_test_result <- moran.test(temp$far, listw)
print(moran_test_result)
```

## distributions

```{r}

temp = data %>% 
  mutate(quad_combo = ifelse(geoiso == "iso" & econiso == 1, "geo & econ",
                      ifelse(geoiso == "iso" & econiso == 0, "geo",
                      ifelse(geoiso == "noncore nonfar" & econiso == 1, "econ (non-core metro)",
                      ifelse(geoiso == "noncore nonfar" & econiso == 0, "neither (non-core metro)",
                      ifelse(geoiso == "core nonfar" & econiso == 1, "econ (core metro)",
                      ifelse(geoiso == "core nonfar" & econiso == 0, "neither (core metro)", NA
                      
  )))))),
  quad_nonmetro = ifelse(noncore == 0, "core metro", quad_combo),
  quad_metro = ifelse(noncore == 1, "non-core metro", quad_combo)) %>% 
  drop_na(econiso, noncore, geoiso) %>% 
  mutate(pov_2012 = ifelse(pov_2012 > 1, 1, pov_2012),
         pov_2023 = ifelse(pov_2023 > 1, 1, pov_2023)
         )


temp_nonmetro = temp %>% 
  mutate(quad_nonmetro = ifelse(quad_nonmetro == "econ (non-core metro)", "econ", quad_nonmetro),
         quad_nonmetro = ifelse(quad_nonmetro == "neither (non-core metro)", "neither", quad_nonmetro)) %>% 
  filter(noncore == 1)
temp_metro = temp %>% 
  mutate(quad_metro = ifelse(quad_metro == "econ (core metro)", "econ", quad_metro),
         quad_metro = ifelse(quad_metro == "neither (core metro)", "neither", quad_metro)) %>% 
  filter(noncore == 0)

t2 = temp %>% 
  mutate(quad_nonmetro = ifelse(quad_nonmetro == "econ (non-core metro)", "econ", quad_nonmetro),
         quad_nonmetro = ifelse(quad_nonmetro == "neither (non-core metro)", "neither", quad_nonmetro),
         quad_nonmetro = ifelse(noncore == 0, "metro", quad_nonmetro))

t2 <- t2 %>%
  mutate(quad_nonmetro = factor(
    quad_nonmetro,
    levels = c("metro", "neither", "geo", "econ", "geo & econ")
  ))


ggplot(data = t2, aes(x = pov_2023, y = quad_nonmetro, fill = quad_nonmetro)) +
  stat_density_ridges(
    geom = "density_ridges_gradient",
    alpha = 0.8,
    quantile_lines = TRUE,
    quantiles = c(0.25, 0.50, 0.75),
    quantile_colour = "white" # Keeps quantile lines visible on dark colors
  ) +
  # Use scale_fill_viridis_d() for discrete data
  scale_fill_viridis_d(
    option = "mako", 
    begin = 0.4 # Avoids the darkest colors at the start of the palette
  ) +
  labs(
    title = "Distributions of tract poverty rate",
    subtitle = "",
    x = "Poverty rate, 2023",
    y = "Tract typology"
  ) +
  theme_ridges() +
  theme(
    plot.title.position = "plot",
    plot.title = element_text(size = 14, hjust = 0),
    plot.subtitle = element_text(hjust = 0),
    axis.text.y = element_text(size = 12, face = "bold"),
    legend.position = "none"
  )

ggplot(data = t2, aes(x = emprt_2023, y = quad_nonmetro, fill = quad_nonmetro)) +
  stat_density_ridges(
    geom = "density_ridges_gradient",
    alpha = 0.8,
    quantile_lines = TRUE,
    quantiles = c(0.25, 0.50, 0.75),
    quantile_colour = "white" # Keeps quantile lines visible on dark colors
  ) +
  # Use scale_fill_viridis_d() for discrete data
  scale_fill_viridis_d(
    option = "mako", 
    begin = 0.4 # Avoids the darkest colors at the start of the palette
  ) +
  labs(
    title = "Distributions of tract employment rate",
    subtitle = "",
    x = "Employment rate, 2023",
    y = "Tract typology"
  ) +
  theme_ridges() +
  theme(
    plot.title.position = "plot",
    plot.title = element_text(size = 14, hjust = 0),
    plot.subtitle = element_text(hjust = 0),
    axis.text.y = element_text(size = 12, face = "bold"),
    legend.position = "none"
  )
```

# tests

```{r}
temp = data %>% 
  mutate(quad_combo = ifelse(geoiso == "iso" & econiso == 1, "geo & econ",
                      ifelse(geoiso == "iso" & econiso == 0, "geo",
                      ifelse(geoiso == "noncore nonfar" & econiso == 1, "econ (non-core metro)",
                      ifelse(geoiso == "noncore nonfar" & econiso == 0, "neither (non-core metro)",
                      ifelse(geoiso == "core nonfar" & econiso == 1, "econ (core metro)",
                      ifelse(geoiso == "core nonfar" & econiso == 0, "neither (core metro)", NA
                      
  )))))),
  quad_nonmetro = ifelse(noncore == 0, "core metro", quad_combo),
  quad_metro = ifelse(noncore == 1, "non-core metro", quad_combo)) %>% 
  drop_na(econiso, noncore, geoiso) %>% 
  mutate(pov_2012 = ifelse(pov_2012 > 1, 1, pov_2012),
         pov_2023 = ifelse(pov_2023 > 1, 1, pov_2023)
         )


temp_nonmetro = temp %>% 
  mutate(quad_nonmetro = ifelse(quad_nonmetro == "econ (non-core metro)", "econ", quad_nonmetro),
         quad_nonmetro = ifelse(quad_nonmetro == "neither (non-core metro)", "neither", quad_nonmetro)) %>% 
  filter(noncore == 1)
temp_metro = temp %>% 
  mutate(quad_metro = ifelse(quad_metro == "econ (core metro)", "econ", quad_metro),
         quad_metro = ifelse(quad_metro == "neither (core metro)", "neither", quad_metro)) %>% 
  filter(noncore == 0)

```

## PS

### emprt

```{r}
# 
# 
# # --- Data prep (important: prevent factor issues) ---
# temp2 <- temp_nonmetro %>%
#   mutate(quad_nonmetro = as.character(quad_nonmetro))
# 
# # --- PS function ---
# ps_stat <- function(data, indices, g1, g2) {
#   d <- data[indices, ]
#   d$quad_nonmetro <- as.character(d$quad_nonmetro)  # ensure character
#   
#   x <- d$pov_2023[d$quad_nonmetro == g1]
#   y <- d$pov_2023[d$quad_nonmetro == g2]
#   
#   # guard: if one group vanishes in resample, return NA
#   if (length(x) == 0 || length(y) == 0) return(NA_real_)
#   
#   mean(outer(x, y, ">")) + 0.5 * mean(outer(x, y, "=="))
# }
# 
# # --- Safe boot wrapper ---
# boot_ps <- function(data, g1, g2, R = 2000) {
#   out <- tryCatch({
#     b <- boot(
#       data = data,
#       statistic = function(d, idx) ps_stat(d, idx, g1, g2),
#       R = R,
#       parallel = "snow",
#       ncpus = 1
#     )
#     
#     ci <- tryCatch(
#       boot.ci(b, type = "perc")$percent[4:5],
#       error = function(e) c(NA, NA)
#     )
#     
#     tibble(
#       group1 = g1,
#       group2 = g2,
#       ps_est = b$t0,
#       ci_low = ci[1],
#       ci_high = ci[2]
#     )
#   }, error = function(e) {
#     tibble(group1 = g1, group2 = g2,
#            ps_est = NA_real_, ci_low = NA_real_, ci_high = NA_real_)
#   })
#   
#   out
# }
# 
# # --- Parallel plan ---
# plan(multisession, workers = parallel::detectCores())
# 
# # --- Generate pairwise group list ---
# groups <- unique(temp2$quad_nonmetro)
# 
# pairwise_results <- expand.grid(group1 = groups, group2 = groups,
#                                 stringsAsFactors = FALSE) %>%
#   filter(match(group1, groups) < match(group2, groups))
# 
# # --- Run all pairwise PS bootstraps in parallel ---
# results <- future_pmap_dfr(
#   list(pairwise_results$group1, pairwise_results$group2),
#   ~ boot_ps(temp2, ..1, ..2, R = 2000)
# )
# 
# # --- Display results ---
# results

```

```{r}
# 
# temp2 <- temp_nonmetro
# 
# # --- PS function ---
# ps_stat <- function(data, indices, g1, g2) {
#   d <- data[indices, ]
#   x <- d$emprt_2023[d$quad_nonmetro == g1]
#   y <- d$emprt_2023[d$quad_nonmetro == g2]
#   
#   # guard: if one group vanishes in resample, return NA
#   if (length(x) == 0 || length(y) == 0) return(NA_real_)
#   
#   mean(outer(x, y, ">")) + 0.5 * mean(outer(x, y, "=="))
# }
# 
# # --- Safe boot wrapper ---
# boot_ps <- function(data, g1, g2, R = 2000) {
#   b <- boot(
#     data = data,
#     statistic = function(d, idx) ps_stat(d, idx, g1, g2),
#     R = R,
#     parallel = "snow",
#     ncpus = 1
#   )
#   
#   # try CI
#   ci <- tryCatch(
#     boot.ci(b, type = "perc")$percent[4:5],
#     error = function(e) c(NA, NA)
#   )
#   
#   tibble(
#     group1 = g1,
#     group2 = g2,
#     ps_est = b$t0,
#     ci_low = ci[1],
#     ci_high = ci[2]
#   )
# }
# 
# # --- Parallel plan ---
# plan(multisession, workers = parallel::detectCores())
# 
# # --- Run all pairwise ---
# groups <- unique(temp2$quad_nonmetro) |> as.character()
# 
# pairwise_results <- expand.grid(group1 = groups, group2 = groups, 
#                                 stringsAsFactors = FALSE) %>%
#   filter(group1 < group2)
# 
# results <- future_pmap_dfr(
#   list(pairwise_results$group1, pairwise_results$group2),
#   ~ boot_ps(temp2, ..1, ..2, R = 2000)
# )
# 
# results
```

```{r}
library(dplyr)
library(purrr)
library(furrr)
library(boot)
library(tibble)

# --- Prepare data ---
# Use your dataset here; must contain:
#  - quad_nonmetro: grouping variable
#  - emprt_2023: numeric outcome variable
temp2 <- temp_nonmetro %>%
  mutate(quad_nonmetro = as.character(quad_nonmetro)) %>%  # ensure character
  filter(!is.na(quad_nonmetro), !is.na(emprt_2023))

# --- Optional: filter groups with too few observations ---
group_counts <- temp2 %>% count(quad_nonmetro)
valid_groups <- group_counts$quad_nonmetro[group_counts$n >= 5]

temp2 <- temp2 %>% filter(quad_nonmetro %in% valid_groups)

# --- PS function ---
ps_stat <- function(data, indices, g1, g2) {
  d <- data[indices, ]
  d$quad_nonmetro <- as.character(d$quad_nonmetro)  # ensure character

  x <- d$emprt_2023[d$quad_nonmetro == g1]
  y <- d$emprt_2023[d$quad_nonmetro == g2]

  if (length(x) == 0 || length(y) == 0) return(NA_real_)

  mean(outer(x, y, ">")) + 0.5 * mean(outer(x, y, "=="))
}

# --- Safe boot wrapper ---
boot_ps <- function(data, g1, g2, R = 2000) {
  out <- tryCatch({
    b <- boot(
      data = data,
      statistic = function(d, idx) ps_stat(d, idx, g1, g2),
      R = R,
      parallel = "snow",
      ncpus = 1
    )

    ci <- tryCatch(
      boot.ci(b, type = "perc")$percent[4:5],
      error = function(e) c(NA, NA)
    )

    tibble(
      group1 = g1,
      group2 = g2,
      ps_est = b$t0,
      ci_low = ci[1],
      ci_high = ci[2]
    )
  }, error = function(e) tibble(
    group1 = g1,
    group2 = g2,
    ps_est = NA_real_,
    ci_low = NA_real_,
    ci_high = NA_real_
  ))

  out
}

# --- Parallel plan ---
plan(multisession, workers = parallel::detectCores())

# --- Create all pairwise combinations safely ---
groups <- sort(unique(temp2$quad_nonmetro))

pairwise_results <- expand.grid(group1 = groups, group2 = groups,
                                stringsAsFactors = FALSE) %>%
  filter(match(group1, groups) < match(group2, groups))

# --- Run all pairwise bootstraps in parallel ---
results <- future_pmap_dfr(
  list(pairwise_results$group1, pairwise_results$group2),
  ~ boot_ps(temp2, ..1, ..2, R = 2000)
)

# --- Display results ---
results
```

### poverty rate

```{r}
library(dplyr)
library(purrr)
library(furrr)
library(boot)
library(tibble)

# --- Prepare data ---
# Use your dataset here; must contain:
#  - quad_nonmetro: grouping variable
#  - emprt_2023: numeric outcome variable
temp2 <- temp_nonmetro %>%
  mutate(quad_nonmetro = as.character(quad_nonmetro)) %>%  # ensure character
  filter(!is.na(quad_nonmetro), !is.na(pov_2023))

# --- Optional: filter groups with too few observations ---
group_counts <- temp2 %>% count(quad_nonmetro)
valid_groups <- group_counts$quad_nonmetro[group_counts$n >= 5]

temp2 <- temp2 %>% filter(quad_nonmetro %in% valid_groups)

# --- PS function ---
ps_stat <- function(data, indices, g1, g2) {
  d <- data[indices, ]
  d$quad_nonmetro <- as.character(d$quad_nonmetro)  # ensure character

  x <- d$pov_2023[d$quad_nonmetro == g1]
  y <- d$pov_2023[d$quad_nonmetro == g2]

  if (length(x) == 0 || length(y) == 0) return(NA_real_)

  mean(outer(x, y, ">")) + 0.5 * mean(outer(x, y, "=="))
}

# --- Safe boot wrapper ---
boot_ps <- function(data, g1, g2, R = 2000) {
  out <- tryCatch({
    b <- boot(
      data = data,
      statistic = function(d, idx) ps_stat(d, idx, g1, g2),
      R = R,
      parallel = "snow",
      ncpus = 1
    )

    ci <- tryCatch(
      boot.ci(b, type = "perc")$percent[4:5],
      error = function(e) c(NA, NA)
    )

    tibble(
      group1 = g1,
      group2 = g2,
      ps_est = b$t0,
      ci_low = ci[1],
      ci_high = ci[2]
    )
  }, error = function(e) tibble(
    group1 = g1,
    group2 = g2,
    ps_est = NA_real_,
    ci_low = NA_real_,
    ci_high = NA_real_
  ))

  out
}

# --- Parallel plan ---
plan(multisession, workers = parallel::detectCores())

# --- Create all pairwise combinations safely ---
groups <- sort(unique(temp2$quad_nonmetro))

pairwise_results <- expand.grid(group1 = groups, group2 = groups,
                                stringsAsFactors = FALSE) %>%
  filter(match(group1, groups) < match(group2, groups))

# --- Run all pairwise bootstraps in parallel ---
results <- future_pmap_dfr(
  list(pairwise_results$group1, pairwise_results$group2),
  ~ boot_ps(temp2, ..1, ..2, R = 2000)
)

# --- Display results ---
results
```

```{r}
# # --- Libraries ---
# library(dplyr)
# library(purrr)
# library(furrr)
# library(boot)
# library(tibble)
# 
# # --- Prepare data ---
# temp2 <- temp_nonmetro %>%
#   st_drop_geometry() %>%
#   mutate(quad_nonmetro = as.character(quad_nonmetro)) %>%
#   filter(!is.na(quad_nonmetro), !is.na(pov_2023))
# 
# # --- Keep groups with enough observations ---
# min_group_size <- 5
# group_counts <- temp2 %>%
#   group_by(quad_nonmetro) %>%
#   summarise(n = n(), .groups = "drop") %>%
#   filter(n >= min_group_size)
# valid_groups <- group_counts$quad_nonmetro
# 
# temp2 <- temp2 %>% filter(quad_nonmetro %in% valid_groups)
# 
# # --- Stratified PS bootstrap ---
# boot_ps_strat <- function(data, g1, g2, R = 2000, ncpus_inner = 2) {
#   tryCatch({
#     x <- data$pov_2023[data$quad_nonmetro == g1]
#     y <- data$pov_2023[data$quad_nonmetro == g2]
#     
#     # skip pairs where either group is empty (safety check)
#     if(length(x) == 0 || length(y) == 0) {
#       return(tibble(group1 = g1, group2 = g2, ps_est = NA_real_,
#                     ci_low = NA_real_, ci_high = NA_real_))
#     }
# 
#     # adaptive R: smaller groups → fewer replicates to save time
#     R_adj <- R
#     small_thresh <- 20
#     if(length(x) < small_thresh || length(y) < small_thresh) R_adj <- min(R, 1000)
# 
#     # stratified bootstrap: use separate vectors, inner parallelization
#     d <- data.frame(x = x, y = y)
#     stat_fun <- function(d, i) {
#       xs <- sample(d$x, replace = TRUE)
#       ys <- sample(d$y, replace = TRUE)
#       mean(outer(xs, ys, ">")) + 0.5 * mean(outer(xs, ys, "=="))
#     }
# 
#     b <- boot(
#       data = d,
#       statistic = stat_fun,
#       R = R_adj,
#       parallel = if (.Platform$OS.type == "windows") "snow" else "multicore",
#       ncpus = ncpus_inner
#     )
# 
#     ci <- tryCatch(boot.ci(b, type = "perc")$percent[4:5],
#                    error = function(e) c(NA, NA))
# 
#     tibble(group1 = g1, group2 = g2, ps_est = b$t0, ci_low = ci[1], ci_high = ci[2])
#   }, error = function(e) {
#     tibble(group1 = g1, group2 = g2, ps_est = NA_real_, ci_low = NA_real_, ci_high = NA_real_)
#   })
# }
# 
# # --- Generate pairwise combinations ---
# groups <- sort(unique(temp2$quad_nonmetro))
# pairwise_results <- expand.grid(group1 = groups, group2 = groups,
#                                 stringsAsFactors = FALSE) %>%
#   filter(match(group1, groups) < match(group2, groups))
# 
# # --- Automatic core allocation ---
# total_cores <- parallel::detectCores()
# outer_cores <- max(1, floor(total_cores / 2))   # outer layer uses half cores
# inner_cores <- max(1, total_cores - outer_cores) # inner bootstrap uses the rest
# 
# plan(multisession, workers = outer_cores)
# 
# # --- Run bootstraps in parallel ---
# results <- future_pmap_dfr(
#   list(pairwise_results$group1, pairwise_results$group2),
#   ~ boot_ps_strat(temp2, ..1, ..2, R = 2000, ncpus_inner = inner_cores),
#   .progress = TRUE
# )
# 
# # --- Final output ---
# results


```

## permutation

```{r}
# 
# # --- Function to compute nonparametric effect stats ---
# compute_effects <- function(data, outcome) {
#   # group means
#   means <- data %>%
#     group_by(econiso, far) %>%
#     summarise(mean_outcome = mean(.data[[outcome]], na.rm = TRUE), .groups = "drop")
#   
#   # main effects (marginal means)
#   mean_econ1 <- means %>% filter(econiso == 1) %>% summarise(m = mean(mean_outcome)) %>% pull(m)
#   mean_econ0 <- means %>% filter(econiso == 0) %>% summarise(m = mean(mean_outcome)) %>% pull(m)
#   f_econ <- mean_econ1 - mean_econ0
#   
#   mean_far1 <- means %>% filter(far == 1) %>% summarise(m = mean(mean_outcome)) %>% pull(m)
#   mean_far0 <- means %>% filter(far == 0) %>% summarise(m = mean(mean_outcome)) %>% pull(m)
#   f_far <- mean_far1 - mean_far0
#   
#   # interaction: difference of differences
#   g11 <- means %>% filter(econiso==1, far==1) %>% pull(mean_outcome)
#   g10 <- means %>% filter(econiso==1, far==0) %>% pull(mean_outcome)
#   g01 <- means %>% filter(econiso==0, far==1) %>% pull(mean_outcome)
#   g00 <- means %>% filter(econiso==0, far==0) %>% pull(mean_outcome)
#   f_interact <- (g11 - g10) - (g01 - g00)
#   
#   c(main_econ = f_econ, main_far = f_far, interact = f_interact)
# }
# 
# # --- Permutation test ---
# perm_test_np <- function(data, outcome, n_perm = 2000, seed = 123) {
#   set.seed(seed)
#   f_obs <- compute_effects(data, outcome)
#   
#   f_perm <- replicate(n_perm, {
#     data_perm <- data
#     data_perm[[outcome]] <- sample(data_perm[[outcome]])  # shuffle outcome
#     compute_effects(data_perm, outcome)
#   })
#   
#   pvals <- sapply(1:3, function(i) mean(abs(f_perm[i, ]) >= abs(f_obs[i])))
#   
#   tibble(
#     outcome = outcome,
#     effect  = names(f_obs),
#     effect_obs = f_obs,
#     p_perm = pvals
#   )
# }
# 
# # --- Run for your outcomes ---
# results_pov <- perm_test_np(temp_nonmetro, "pov_2023", n_perm = 2000)
# results_emprt <- perm_test_np(temp_nonmetro, "emprt_2023", n_perm = 2000)
# 
# results_perm <- bind_rows(results_pov, results_emprt)
# results_perm

```

```{r}
# 
# temp_nonmetro2 = temp_nonmetro %>% 
#   st_drop_geometry()
# 
# # --- Set up parallel plan ---
# plan(multisession, workers = parallel::detectCores())
# 
# # --- Parallel permutation test function ---
# perm_test_np_parallel <- function(data, outcome, n_perm = 2000, seed = 123) {
#   set.seed(seed)
#   f_obs <- compute_effects(data, outcome)
#   
#   # Run permutations in parallel using future_map
#   f_perm_list <- future_map(1:n_perm, ~{
#     data_perm <- data
#     data_perm[[outcome]] <- sample(data_perm[[outcome]])
#     compute_effects(data_perm, outcome)
#   })
#   
#   # Convert list to matrix: rows = effects, cols = permutations
#   f_perm <- do.call(cbind, f_perm_list)
#   
#   # Compute permutation p-values
#   pvals <- map_dbl(1:3, ~mean(abs(f_perm[., ]) >= abs(f_obs[.])) )
#   
#   tibble(
#     outcome = outcome,
#     effect  = names(f_obs),
#     effect_obs = f_obs,
#     p_perm = pvals
#   )
# }
# 
# # --- Multiple outcomes in parallel ---
# outcomes <- c("pov_2023", "emprt_2023")
# 
# results_perm <- future_map_dfr(outcomes, ~perm_test_np_parallel(temp_nonmetro2, .x, n_perm = 2000))
# results_perm
```

interaction plot - haven't tested if it works

```{r}
# # --- Compute observed group means ---
# means <- temp_nonmetro %>%
#   group_by(econiso, far) %>%
#   summarise(poverty_rate = mean(pov_2023, na.rm = TRUE), .groups = "drop")
# 
# # --- Compute additive expectations ---
# grand_mean <- mean(means$poverty_rate)
# 
# econ_effect <- means %>%
#   group_by(econiso) %>%
#   summarise(econ_mean = mean(poverty_rate)) %>%
#   mutate(effect = econ_mean - grand_mean) %>%
#   select(econiso, effect)
# 
# far_effect <- means %>%
#   group_by(far) %>%
#   summarise(far_mean = mean(poverty_rate)) %>%
#   mutate(effect = far_mean - grand_mean) %>%
#   select(far, effect)
# 
# means <- means %>%
#   left_join(econ_effect, by = "econiso") %>%
#   rename(econ_effect = effect) %>%
#   left_join(far_effect, by = "far") %>%
#   rename(far_effect = effect) %>%
#   mutate(expected_additive = grand_mean + econ_effect + far_effect)
# 
# # --- Plot observed vs additive ---
# ggplot(means, aes(x = factor(far), y = poverty_rate,
#                   group = factor(econiso), color = factor(econiso))) +
#   geom_line(aes(linetype = "Observed")) +
#   geom_point() +
#   geom_line(aes(y = expected_additive, linetype = "Additive")) +
#   labs(x = "Geographic Isolation (far)",
#        y = "Poverty Rate",
#        color = "Economic Isolation",
#        linetype = "Line Type",
#        title = "Observed vs Expected (Additive) Interaction: Poverty") +
#   theme_minimal()
```

```{r}
# --- Libraries ---
library(dplyr)
library(purrr)
library(furrr)
library(tibble)
library(sf)  # assuming temp_nonmetro has geometry

# --- Prepare data ---
temp_nonmetro2 <- temp_nonmetro %>%
  st_drop_geometry()

# --- Helper function: compute nonparametric effects ---
compute_effects <- function(data, outcome) {
  means <- data %>%
    group_by(econiso, far) %>%
    summarise(mean_outcome = mean(.data[[outcome]], na.rm = TRUE), .groups = "drop")
  
  mean_econ1 <- means %>% filter(econiso == 1) %>% summarise(m = mean(mean_outcome)) %>% pull(m)
  mean_econ0 <- means %>% filter(econiso == 0) %>% summarise(m = mean(mean_outcome)) %>% pull(m)
  f_econ <- mean_econ1 - mean_econ0
  
  mean_far1 <- means %>% filter(far == 1) %>% summarise(m = mean(mean_outcome)) %>% pull(m)
  mean_far0 <- means %>% filter(far == 0) %>% summarise(m = mean(mean_outcome)) %>% pull(m)
  f_far <- mean_far1 - mean_far0
  
  g11 <- means %>% filter(econiso == 1, far == 1) %>% pull(mean_outcome)
  g10 <- means %>% filter(econiso == 1, far == 0) %>% pull(mean_outcome)
  g01 <- means %>% filter(econiso == 0, far == 1) %>% pull(mean_outcome)
  g00 <- means %>% filter(econiso == 0, far == 0) %>% pull(mean_outcome)
  
  f_interact <- (g11 - g10) - (g01 - g00)
  
  c(main_econ = f_econ, main_far = f_far, interact = f_interact)
}

# --- Parallel permutation test function ---
perm_test_np_parallel <- function(data, outcome, n_perm = 2000, seed = 123) {
  set.seed(seed)
  
  # observed effects
  f_obs <- compute_effects(data, outcome)
  
  # reproducible random seeds
  seeds <- sample.int(1e6, n_perm, replace = FALSE)
  
  # parallel map with package loading
  f_perm_list <- future_map(
    seeds,
    ~{
      set.seed(.x)
      data_perm <- data
      data_perm[[outcome]] <- sample(data_perm[[outcome]])
      compute_effects(data_perm, outcome)
    },
    .packages = c("dplyr", "purrr", "tibble"),
    .progress = TRUE
  )
  
  # combine permutations
  f_perm <- do.call(cbind, f_perm_list)
  
  # compute two-sided permutation p-values
  pvals <- map_dbl(1:3, ~mean(abs(f_perm[., ]) >= abs(f_obs[.])))
  
  tibble(
    outcome = outcome,
    effect  = names(f_obs),
    effect_obs = f_obs,
    p_perm = pvals
  )
}

# --- Set up parallel plan ---
plan(multisession, workers = parallel::detectCores())

# --- Run across outcomes ---
outcomes <- c("pov_2023", "emprt_2023")

results_perm <- future_map_dfr(
  outcomes,
  ~perm_test_np_parallel(temp_nonmetro2, .x, n_perm = 2000),
  .progress = TRUE
)

# --- Output ---
results_perm


```
