---
title: "Mortality Decomposition Data"
author: "Sara Peters"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#Set your working directory to where the mortality data is saved.

setwd("C:/Users/saraa/OneDrive - UW-Madison/Mortality Data-CDC Wonder/IH2") 


# Load necessary libraries
library(readr)
library(dplyr)
library(purrr)
library(stringr)
library(tidyr)
library(readxl)



```

## Industry Health 2 Race-Specific Mortality Data 

##Define File Types and Functions

```{r}

# Define race suffixes instead of file types
race_suffixes <- c("Black", "White", "AIAN", "API")  # Reflects your filenames
file_types <- "AllCause"

# RUCC data
ruccdata <- read_excel("Rural.Urban.Continuum.Codes.1974.1983.1993.2003.2013.xls") %>%
  select(FIPS, 
         `Rural-Urban Continuum Code 1974`, 
         `Rural-Urban Continuum Code 1983`, 
         `Rural-Urban Continuum Code 1993`, 
         `Rural-Urban Continuum Code 2003`, 
         `Rural-Urban Continuum Code 2013`)

# Extract Race Group from File name
extract_race_group <- function(file_name) {
  str_extract(file_name, "(Black|White|AIAN|API)(?=\\.txt$)")
}

# Main processing function
process_region_files <- function(region, year_range, year) {
  race_suffixes <- c("Black", "White", "AIAN", "API")
  file_paths <- paste0(year_range, "-All/", year_range, "-", region, "-", file_types, "-", race_suffixes, ".txt")

  for (file_path in file_paths) {
    if (!file.exists(file_path)) next

    df <- read_delim(file_path, delim = "\t", quote = '"', col_types = cols(.default = "c"), trim_ws = TRUE)

    if ("Notes" %in% names(df)) {
      metadata_row <- which(df$Notes == "---")[1]
      if (!is.na(metadata_row)) {
        df <- df[1:(metadata_row - 1), ]
      }
    }

    df <- df %>%
      mutate(
        Deaths = as.numeric(na_if(as.character(Deaths), "Suppressed")),
        Population = as.numeric(na_if(as.character(Population), "Suppressed")),
        across(where(~ all(str_detect(., "^-?\\d+(\\.\\d+)?$"), na.rm = TRUE)), as.numeric)
      )

    race_group <- str_extract(file_path, "(Black|White|AIAN|API)")
    df <- df %>% mutate(race_group = race_group)

    df <- df %>% filter(rowSums(is.na(.)) != ncol(.))

    name <- paste0(race_group, "-", region, "-", year)
    if (exists(name, envir = .GlobalEnv)) {
      existing_df <- get(name, envir = .GlobalEnv)
      assign(name, bind_rows(existing_df, df), envir = .GlobalEnv)
    } else {
      assign(name, df, envir = .GlobalEnv)
    }
  }

  message(paste("All dataframes for", region, "created successfully!"))
}

```


##2013-2015


###Define dynamic values
```{r}
# Define dynamic year values
year_range <- "13-15"
year <- "2013"



```



####East South Central

```{r}

# Define Regions
regions <- c("AL-KY-MS-TN", "AL-KY", "MS-TN")

# Process and bind
lapply(regions, process_region_files, year_range = year_range, year = year)

df_names <- unlist(lapply(regions, function(r) {
  paste0(race_suffixes, "-", r, "-", year)
})) %>% as.vector()

ESC_combined_13 <- bind_rows(
  mget(df_names, ifnotfound = as.list(rep(list(NULL), length(df_names)))),
  .id = "Source"
)

# Add region labels
ESC_combined_13 <- ESC_combined_13 %>%
  mutate(
    `Sub-Region` = "East-South-Central",
    `Region` = "South",
    `Year` = year
  ) 


```

####West South Central 

```{r}

# Define regions
regions <- c("AR-LA-OK", "TX")


# Process and bind
lapply(regions, process_region_files, year_range = year_range, year = year)

df_names <- unlist(lapply(regions, function(r) {
  paste0(race_suffixes, "-", r, "-", year)
})) %>% as.vector()

WSC_combined_13 <- bind_rows(
  mget(df_names, ifnotfound = as.list(rep(list(NULL), length(df_names)))),
  .id = "Source"
)

# Add new columns for Sub-Region and Region
WSC_combined_13 <- WSC_combined_13 %>%
  mutate(
    `Sub-Region` = "West-South-Central",
    `Region` = "South", 
    `Year` = "2013"
  ) 




```

####South Atlantic

```{r}

# Define regions
regions <- c("DE-DC", "FL-GA-MD", "NC-SC", "VA-WV")


## Process and bind
lapply(regions, process_region_files, year_range = year_range, year = year)

df_names <- unlist(lapply(regions, function(r) {
  paste0(race_suffixes, "-", r, "-", year)
})) %>% as.vector()

SA_combined_13 <- bind_rows(
  mget(df_names, ifnotfound = as.list(rep(list(NULL), length(df_names)))),
  .id = "Source"
)


# Add new columns for Sub-Region and Region
SA_combined_13 <- SA_combined_13 %>%
  mutate(
    `Sub-Region` = "South-Atlantic",
    `Region` = "South", 
    `Year` = "2013"
  ) 


```

####New England

```{r}

# Define regions
regions <- c("CT-ME-MA-NH-RI-VT")


# Process and bind
lapply(regions, process_region_files, year_range = year_range, year = year)

df_names <- unlist(lapply(regions, function(r) {
  paste0(race_suffixes, "-", r, "-", year)
})) %>% as.vector()

NewEngland_combined_13 <- bind_rows(
  mget(df_names, ifnotfound = as.list(rep(list(NULL), length(df_names)))),
  .id = "Source"
)

# Add new columns for Sub-Region and Region
NewEngland_combined_13 <- NewEngland_combined_13 %>%
  mutate(
    `Sub-Region` = "New-England",
    `Region` = "Northeast", 
    `Year` = "2013"
  )


```




####Middle Atlantic

```{r}



# Define regions
regions <- c("NJ-NY-PA")


# Process and bind
lapply(regions, process_region_files, year_range = year_range, year = year)

df_names <- unlist(lapply(regions, function(r) {
  paste0(race_suffixes, "-", r, "-", year)
})) %>% as.vector()

MiddleAtlantic_combined_13 <- bind_rows(
  mget(df_names, ifnotfound = as.list(rep(list(NULL), length(df_names)))),
  .id = "Source"
)

# Add new columns for Sub-Region and Region
MiddleAtlantic_combined_13 <- MiddleAtlantic_combined_13 %>%
  mutate(
    `Sub-Region` = "Middle-Atlantic",
    `Region` = "Northeast", 
    `Year` = "2013"
  ) 


```

####East North Central
```{r}

# Define regions
regions <- c("IL-IN-OH", "MI-WI")


# Process and bind
lapply(regions, process_region_files, year_range = year_range, year = year)

df_names <- unlist(lapply(regions, function(r) {
  paste0(race_suffixes, "-", r, "-", year)
})) %>% as.vector()

ENC_combined_13 <- bind_rows(
  mget(df_names, ifnotfound = as.list(rep(list(NULL), length(df_names)))),
  .id = "Source"
)

# Add new columns for Sub-Region and Region
ENC_combined_13 <- ENC_combined_13 %>%
  mutate(
    `Sub-Region` = "East-North-Central",
    `Region` = "MidWest", 
    `Year` = "2013"
  ) 


```

####West North Central

```{r}

# Define regions
regions <- c("IA-MN-MO", "KS-NE-ND-SD")


# Process and bind
lapply(regions, process_region_files, year_range = year_range, year = year)

df_names <- unlist(lapply(regions, function(r) {
  paste0(race_suffixes, "-", r, "-", year)
})) %>% as.vector()

WNC_combined_13 <- bind_rows(
  mget(df_names, ifnotfound = as.list(rep(list(NULL), length(df_names)))),
  .id = "Source"
)

# Add new columns for Sub-Region and Region
WNC_combined_13  <- WNC_combined_13 %>%
  mutate(
    `Sub-Region` = "West-North-Central",
    `Region` = "MidWest", 
    `Year` = "2013"
  ) 



```

####Pacific

```{r}

# Define regions
regions <- c("AK-CA-HI-OR-WA")



# Process and bind
lapply(regions, process_region_files, year_range = year_range, year = year)

df_names <- unlist(lapply(regions, function(r) {
  paste0(race_suffixes, "-", r, "-", year)
})) %>% as.vector()

Pacific_combined_13 <- bind_rows(
  mget(df_names, ifnotfound = as.list(rep(list(NULL), length(df_names)))),
  .id = "Source"
)

# Add new columns for Sub-Region and Region
Pacific_combined_13 <- Pacific_combined_13 %>%
  mutate(
    `Sub-Region` = "Pacific",
    `Region` = "West", 
    `Year` = "2013"
  ) 



```

####Mountain

```{r}

# Define regions
regions <- c("ID-MT-NV-WY", "AZ-CO-NM-UT")


# Process and bind
lapply(regions, process_region_files, year_range = year_range, year = year)

df_names <- unlist(lapply(regions, function(r) {
  paste0(race_suffixes, "-", r, "-", year)
})) %>% as.vector()

Mountain_combined_13 <- bind_rows(
  mget(df_names, ifnotfound = as.list(rep(list(NULL), length(df_names)))),
  .id = "Source"
)

# Add new columns for Sub-Region and Region
Mountain_combined_13 <- Mountain_combined_13 %>%
  mutate(
    `Sub-Region` = "Mountain",
    `Region` = "West", 
    `Year` = "2013"
  ) 


```

####South

```{r}

#Combined Region
# List of dynamically created dataframe names
df_names <- c(
   "SA_combined_13",
  "ESC_combined_13",
  "WSC_combined_13"
)

# Bind all dataframes together
South_combined_13 <- bind_rows(mget(df_names), .id = "Source_SubRegion")




```


####Northeast

```{r}
#Combined Region
# List of dynamically created dataframe names
df_names <- c(
   "NewEngland_combined_13",
  "MiddleAtlantic_combined_13"
)

# Bind all dataframes together 
Northeast_combined_13 <- bind_rows(mget(df_names), .id = "Source_SubRegion")


```


####Midwest

```{r}

#Combined Region MidWest
# List of dynamically created dataframe names
df_names <- c(
   "ENC_combined_13",
  "WNC_combined_13"
)

# Bind all dataframes together
Midwest_combined_13 <- bind_rows(mget(df_names), .id = "Source_SubRegion")



```


####West

```{r}

#Combined Region West
# List of dataframe names
df_names <- c(
   "Pacific_combined_13",
  "Mountain_combined_13"
)

# Bind all dataframes together
West_combined_13 <- bind_rows(mget(df_names), .id = "Source_SubRegion")



```

####All 2013-2015

```{r}

#Pooled 2013 Data
# List of dataframe names
df_names <- c(
   "South_combined_13",
  "Northeast_combined_13", 
  "Midwest_combined_13",
  "West_combined_13"
)

# Bind all dataframes together
All_13_combined <- bind_rows(mget(df_names), .id = "Source_SubRegion")



```


####Validate Parsing
```{r}
# Define valid values
valid_sex <- c("Female", "Male")  # If "Sex Code" contains different labels, update this list
valid_race_group <- c("Black", "White", "AIAN", "API")

# Find rows where "Sex Code" is not in the valid list
invalid_sex <- All_13_combined %>%
  mutate(row_number = row_number()) %>%
  filter(!(Sex %in% valid_sex) & !is.na(Sex)) %>%
  select(row_number, Sex)

# Find rows where "Broad-Type" has unexpected values
invalid_race_group <- All_13_combined %>%
  mutate(row_number = row_number()) %>%
  filter(!(race_group %in% valid_race_group) & !is.na(race_group)) %>%
  select(row_number, race_group)

# Print results or "No invalid values" message
if (nrow(invalid_sex) == 0) {
  cat("✅ No invalid values found in 'Sex Code' column.\n")
} else {
  cat("🔍 Invalid values in 'Sex Code' column:\n")
  print(invalid_sex)
}

if (nrow(invalid_race_group) == 0) {
  cat("✅ No invalid values found in 'Brace column.\n")
} else {
  cat("\n🔍 Invalid values in 'Race_Group' column:\n")
  print(invalid_race_group)
}


```


####Full File Clean up

```{r}

#Drop Total and remaining notes rows


# Keep only rows where "Notes" is NA
All_13_combined <- All_13_combined %>%
  filter(is.na(Notes))

non_na_notes <- All_13_combined %>%
  filter(!is.na(Notes))

# Print the number of non-NA rows
cat("Number of rows where 'Notes' is NOT NA:", nrow(non_na_notes), "\n")

# Print unique values in "Notes" (if any exist)
if (nrow(non_na_notes) > 0) {
  print(unique(non_na_notes$Notes))
} else {
  cat("All values in 'Notes' are NA.\n")
}

# Remove the "Notes" column if no values in Notes
All_13_combined <- All_13_combined %>%
  select(-Notes)


#Create State Column and Rename County Code to FIPS Code

# Separate "County" column into "County" and "State"
All_13_combined <- All_13_combined %>%
  separate(County, into = c("County", "State"), sep = ", ", remove = FALSE) %>% 
  rename(FIPS = `County Code`)


#Add RUCC Codes


All_13_combined <- left_join(ruccdata, All_13_combined, by=c("FIPS"))

##preparing rucc metro/nonmetro 
All_13_combined = All_13_combined %>% 
  mutate(metro_nonmetro_13 = if_else(`Rural-Urban Continuum Code 2013` < 4, 0, 1)) %>% 
  mutate(rurality_13 = if_else(metro_nonmetro_13 == 0, "metro", "nonmetro"))

## Recategorize "Hispanic-Latino"

#Vector for hispanic origin values == "Hispanic Origin"; values == "Hispanic or Latino", "Not Hispanic or Latino", or "Not Stated"

#Vector for racialized category == "race_group"; values == "AIAN", "API", "Black", "White"

#can mutate "race_group" to NHBlack, etc., or use a combo of "Hispanic Origin" and "race_group" in sample selection. 

#Will need to decide what to do about cases where Hispanic Origin is "Not Stated", exclude or assume not hispanic? 


```
##Get Current Popoulation Estimates 2013-2015


#Calculate race-specific age-adjusted premature mortality for each county



