---
title: "Industry-Health_new"
editor: visual
format:
  html:
    self-contained: true
    page-layout: full
    code-fold: true
    code-tools: true
    code_download: yes
    latex_engine: pdflatex
params:
  year: 2012
  year_range: !expr c(2012, 2017) # or c(2012, 2015, 2017) or c(2010:2020)
  show_temporal_density: TRUE
  show_descriptive_density: TRUE
  show_descriptive_density_rural: TRUE
  show_descriptive_maps: TRUE
  interactive_maps: FALSE
  show_corr: FALSE
  show_test_regression: FALSE
  show_lm_regression: TRUE
  show_lm_cluster_regression: TRUE
  reg_factors: !expr c("entry_rate", "exit_rate", "establishments_gr", "wage", "wage_gr", "jobs_gr", "payroll_gr", "unemp_rate", "laborforce_part_rate", "employment_gr", "poverty_rate", "income_rate", "poverty_rate_gr", "income_rate_gr", "population_gr",  "ypll75", "highschool_attainment_rate", "net_migration_rate") 
  reg_factor_funs: !expr c("call_establishments", "call_wage", "call_jobs", "call_payroll", "call_employment", "call_poverty_rate", "call_income_rate", "call_population")
  example_coef_index:  2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

install.packages('tinytex')
tinytex::install_tinytex()
tinytex:::install_prebuilt()

require(kableExtra)
library(tinytex)
library(knitr)
library(tidycensus)
```

## Industry and the Rural Mortality Penalty: A spatial regime analysis of economic structure and early life mortality in the rural US

### Abstract:

The last half century in the rural United States has seen stagnant life expectancy, an increasing mortality penalty, and industrial restructuring as a result of increasing globalization and automation. The relationships between mortality, industry, and rurality are well established. Prior literature has not fully investigated the relationship between early life mortality and rurality, including whether industrial restructuring contributes to the rural mortality penalty. Additionally, while prior studies have emphasized the importance of place in understanding the mortality penalty, the attributes of place including spatial location are not well-understood. We empirically investigate the influence of early life mortality on life expectancy and the spatial patterning of these relationships according to economic structure. Leveraging novel data, we use spatial regime analysis to assess the spatial patterning of relationships between premature death and measures of dominant industry and trade of intermediate commodities for all US counties. This study offers new insights on place attributes and the spatial dimensions of the rural mortality penalty. 

## Economic Dynamism

### Conceptual definition

#### Trade

#### Industrial Structure

#### Value Added + Supply Chain

Total industry output (basic prices) = Total Intermediate + **Value Added (basic prices)**

**Value Added (basic prices)** = Compensation of employees + Other taxes on production - Other subsidies on production + Gross operating surplus

## Health

### Conceptual definition

## **Analytical plan**

### **List of Data and measures**

Outcome: 

1.  premature death rate (period average of (3-year ave) for 2012-17 from CHRR

Key independent variables: 

Trade

1.  autonomy (2012) from Rurec circularity 
2.  retention (2012) from Rurec

Supply chain

1.  Proportion value-added (all industries)
2.  Proportion AFM value-added?

Industrial structure (pick one to present but explore all for sensitivity analysis)

1.  industry-specific employment rate from ACS/BLS
    -   Ag, mining, mftg, FIRE, eds/meds, recreation
    -   ?? regression on premature death = share of every sector
    -   Which loads?
2.  industry-specific output from Rurec
3.  county typologies from USDA ERS (dependency)

Covariates

1.  labor force participation (xxx at period start) from ACS
2.  population size (logged population at period start) from ACS
3.  population change ((period end - period start)/period start) from ACS
4.  age structure (median age) from ACS
5.  Education distribution
6.  Access to health care

### Regressions

```{r setup, include=FALSE}

install.packages(vtable)
install.packages(dplyr)
install.packages(tidyverse)
install.packages(ggplot2)
install.packages(stringr)
install.packages(spdep)
install.packages(tidycensus)
install.packages(tigris)
install.packages(viridis)
install.packages(cowplot)
# install.packages("estimatr")
# install.packages("huxtable")
# install.packages("margins")
#install.packages(c("lmtest", "sandwich"))
#install.packages("spatialreg")



#library(vtable)
library(dplyr)
library(tidyverse)
library(ggplot2)
library(spdep)
library(tidycensus)
library(tigris)
library(viridis)
library(cowplot)
library(lmtest)
library(sandwich)
library(estimatr)
library(huxtable)
library(margins)
library("spatialreg")

my_color_palette <- viridis(9, option = "D")

palette(my_color_palette)

# scripts
source("projects/industry_health/dataprep.R", local = (ih_dp <- new.env()))
source("datasets/circularity/circularity.R", local = (circ <- new.env()))

# chunk behavior 
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 

# nonscientific notation
options(scipen=999)

# log output 
log_threshold(DEBUG)

# initialize parameters
params



## Data
#Load Data from Working Directory


##Load Data Sets 
#TODO: For CLAYTON


##Median Age of County 2012
age = get_acs(geography = "county",
                   survey = "acs5",
                   variables = c("S0101_C01_030E"
                   ),
                   year = params$year) %>%
  select("place"=GEOID, "age"=estimate)



##ERS County Typologies 

ers = read.csv("https://www.ers.usda.gov/webdocs/DataFiles/48652/2015CountyTypologyCodes.csv?v=8831.8") %>%
  mutate(place = sprintf("%05d", FIPStxt)) %>%
  select(place, "ers_ag"=Farming_2015_Update, "ers_mine"=Mining_2015.Update, "ers_mfg"=Manufacturing_2015_Update, "ers_gov"=Government_2015_Update, "ers_rec"=Recreation_2015_Update, "ers_nonsp"=Nonspecialized_2015_Update)



##Industry Structure

industry = get_acs(geography = "county",
                   survey = "acs5",
                   variables = c("DP03_0032E","DP03_0033E","DP03_0034E"
                                        ,"DP03_0035E","DP03_0036E","DP03_0037E"
                                        ,"DP03_0038E","DP03_0039E","DP03_0040E"
                                        ,"DP03_0041E","DP03_0042E","DP03_0043E"
                                        ,"DP03_0044E","DP03_0045E"
                   ),
                   year = params$year) %>%
  select("place"=GEOID, variable, estimate) %>%
  pivot_wider(names_from = variable, values_from = estimate) %>%
  mutate(pct_afm = DP03_0033/DP03_0032, pct_mfg = DP03_0034/DP03_0032, pct_fire = DP03_0039/DP03_0032, pct_edmed = DP03_0041/DP03_0032, pct_rec = DP03_0042/DP03_0032, pct_gov = DP03_0044/DP03_0032) %>%
  select(place, pct_afm, pct_mfg, pct_fire, pct_edmed, pct_rec, pct_gov)


```

```{r zip pack, include = FALSE}

# uncomment and pack/unpack as necessary
# dp$zip_pack("tmp/eca_paa_data_v240324.zip", dp$opath)
# dp$zip_unpack("tmp/eca_paa_data_v240324.zip", overwrite = TRUE)

```

::: panel-tabset

# data

```{r data cashe, include = FALSE}

# uncomment as necessary
# dp$create_complete_cache()

```

### Premature Death Change over time (2012-2017)

```{r}

##Construct typology of premature death change over time between 2012-2017 

#Calculate change over time

# Construct typology E.g. 4 types:
    
# High premature death in t0 & stable/reduction in premature death from t-1 to t0
    #   E.g. maybe southeast?
# High premature death in t0 & material increase in premature death from t-1 to t0
    # E.g. maybe Appalachia?
# Low premature death in t0 & stable in premature death from t-1 to t0
# Low premature death in t0 & material increase in premature death from t-1 to t0

```

### Access to Healthcare

```{r}
# Access to health care----------

# Normalize the data
# data$Uninsured_Ratio <- data$Uninsured_Percentage / 100
# data$Primary_Care_Ratio <- data$Population / data$Primary_Care_Physicians
# data$Dentists_Ratio <- data$Population / data$Dentists
# data$Mental_Health_Ratio <- data$Population / data$Mental_Health_Providers
# 
# # Create a composite score for each county
# data$Healthcare_Access_Score <- (1 - data$Uninsured_Ratio) * (1 - data$Primary_Care_Ratio / max(data$Primary_Care_Ratio)) *
#   (1 - data$Dentists_Ratio / max(data$Dentists_Ratio)) * (1 - data$Mental_Health_Ratio / max(data$Mental_Health_Ratio))

```

\`\`\`

## Summary Statistics

### Descriptive Statistics

### ESDA

### Density Plots

### Maps

## Analysis - Multinomial Logit Model

For Statistics

### Model 1-

## Analysis- Exploratory Regressions

Factors

Trade

Autonomy

etention

Supply chain value added

\% AFM % value added

\% worker share of value added

\% worker share of output

Industrial structure

ERS dependency typologies

\% composition (output)

\% composition (employment)

Prosperity

Poverty,

education,

income,

etc.

Covariates

Access to healthcare,

workforce,

population,

age

## Analysis- Premature Death

Death = trade Death = trade + covariates

Death = trade + industrial + covariates

Death = trade + supply chain + covariates

Death = trade + industry + supply chain + covariates

Death = prosperity + covariates

Death = prosperity + industrial + covariates

Death = trade + prosperity + covariates

Death = trade + prosperity + industrial + covariates

Death = trade + prosperity + industrial + supply chain + covariates

Specifications

Error

Error + interactions

Error + regimes

Pat

## Analysis- Spatial Regime

### 
