---
title: "Rural in BDS"
format:
  html:
    toc: true
    toc-depth: 2
    embed-resources: true
    code-fold: true
---


```{r}
#| output: false

library(tidyverse)

theme_set(theme_minimal())

param <- list(
  rural_ruc = 4:9
)
```

# Data

```{r}
df_ruc_2003 <- pubdata::ers_rural_get("ruc_2003")

df_ruc_2023 <- pubdata::ers_rural_get("ruc_2023") %>%
  mutate(name = case_match(
    attribute,
    "Population_2020" ~ "population_2020",
    "RUCC_2023" ~ "ruc_code_2023",
    "Description" ~ "ruc_desc_2023"
    )) %>%
  pivot_wider(id_cols = c("fips", "state", "county_name")) %>%
  mutate(across(c(population_2020, ruc_code_2023), as.numeric))

df_bds <- pubdata::bds_get("2022_st_cty")

df_main <- df_bds %>%
  mutate(fips = paste0(st, cty), .before = "st") %>%
  left_join(df_ruc_2003 %>% select(fips, ruc_code_2003), by = "fips") %>%
  left_join(df_ruc_2023 %>% select(fips, ruc_code_2023), by = "fips") %>%
  mutate(
    rural_2003 = (ruc_code_2003 %in% param$rural_ruc),
    rural_2023 = (ruc_code_2023 %in% param$rural_ruc),
    rural_tr = case_when(
      rural_2003 & !rural_2023 ~ "fr",
      !rural_2023 ~ "urban",
      .default = "rural"
    )
  )

df_ruc_2003 %>% head()
df_ruc_2023 %>% head()
df_main %>% head()


```


# Rurality definition

```{r}
df_ruc_2003 %>%
  summarize(counties = n(), population_2000 = sum(population_2000, na.rm = TRUE), .by = c(ruc_code_2003, ruc_desc_2003)) %>%
  arrange(ruc_code_2003)

df_ruc_2023 %>%
  summarize(counties = n(), population_2020 = sum(population_2020), .by = c(ruc_code_2023, ruc_desc_2023)) %>%
  arrange(ruc_code_2023)
```

Defining rural as "non-metro".

```{r}
param["rural_ruc"]

df_main %>%
  filter(year == 2022) %>%
  count(rural_2003, rural_2023)

```


# Establishment and employment growth

From 2000 to 2019, urban greatly outperforms rural.
Rural looks worse under more recent definition.

```{r}

comp_gr <- function(rur_col, y0, y1) {
df_main %>%
  filter(between(year, y0, y1)) %>%
  summarize(estabs = sum(estabs, na.rm = TRUE), emp = sum(emp, na.rm = TRUE), .by = c(year, .data[[rur_col]])) %>%
  mutate(
    est_gr = estabs / first(if_else(year == y0, estabs, NA)),
    emp_gr = emp / first(if_else(year == y0, emp, NA)),
    .by = .data[[rur_col]]
  ) %>%
  pivot_longer({{rur_col}}, names_to = "rurality", values_to = "rural")
}

bind_rows(
  comp_gr("rural_2003", 2000, 2019),
  comp_gr("rural_2023", 2000, 2019)
) %>%
  pivot_longer(c(est_gr, emp_gr)) %>%
  ggplot() + 
  geom_line(aes(year, value, color = rurality, linetype = rural)) +
  facet_wrap("name") +
  theme(legend.position = "bottom")

```

From Great Recession to Covid, rural employment growth lagged, and number of establishments remained stagnant.

```{r}

comp_gr("rural_2023", 2010, 2019) %>%
  pivot_longer(c(est_gr, emp_gr)) %>%
  ggplot() + 
  geom_line(aes(year, value, linetype = rural)) +
  facet_wrap("name") +
  theme(legend.position = "bottom")

```


# Job creation and destruction

Rural employment is less dynamic: both job creation and destruction rates are below urban.

```{r}
y0 = 2010
y1 = 2019

d <- df_main %>%
  filter(between(year, y0, y1)) %>%
  summarize(across(
    c(denom, job_creation, job_creation_births, job_creation_continuers, job_destruction, job_destruction_deaths, job_destruction_continuers),
    \(x) sum(x, na.rm = TRUE)), 
    .by = c(year, rural_2023)) %>%
  mutate(across(
    c(job_creation, job_creation_births, job_creation_continuers, job_destruction, job_destruction_deaths, job_destruction_continuers),
    \(x) x / denom))

d %>%
  pivot_longer(c(job_creation, job_destruction)) %>%
  ggplot() +
  geom_line(aes(year, value, color = name, linetype = rural_2023)) +
  ylim(c(0, NA))

```


Job destruction rate by continuing establishments in rural is similar to urban, but creation is much lower.
Net creation by continuers is thus much higher in urban than in rural.

```{r}
d %>%
  pivot_longer(c(job_creation_births, job_creation_continuers, job_destruction_deaths, job_destruction_continuers)) %>%
  mutate(name = ordered(name, c("job_creation_births", "job_destruction_deaths", "job_creation_continuers", "job_destruction_continuers"))) %>%
  ggplot() +
  geom_line(aes(year, value, linetype = rural_2023)) +
  ylim(c(0, NA)) +
  facet_wrap("name", ncol = 2)


```

"Creative destruction" (birth and death) accounts for about a third of all job dynamism in both rural and urban.
In other words, about half as many jobs are created and destroyed by entry and exit.

If we combine young establishments with birth, the role of start-ups in job creation will be even higher. 
*No figure here to support this claim, but it is almost certainly true.*


```{r}
d %>%
  pivot_longer(c(job_creation_births, job_creation_continuers, job_destruction_deaths, job_destruction_continuers)) %>%
  mutate(continuer = str_detect(name, "continuers"), creation = str_detect(name, "creation")) %>%
  mutate(creation = case_match(creation, TRUE ~ "creation", FALSE ~ "destruction"),
         continuer = case_match(continuer, TRUE ~ "continuer", FALSE ~ "death birth"),
         rural_2023 = case_match(rural_2023, TRUE ~ "rural", FALSE ~ "urban")) %>%
  ggplot() +
  geom_line(aes(year, value, color = creation, linetype = continuer)) +
  ylim(c(0, NA)) +
  facet_wrap("rural_2023", ncol = 2)

```

# Creative destruction across counties

Looking at un-weighted correlation between county establishment and employment growth rate from 2010 to 2019, employment dynamism plays smaller role in job creation in rural areas.
But if we weight counties by employment, rural effect becomes positive.

```{r}
d <- df_main %>%
  filter(year %in% c(y0, y1)) %>%
  rename(est = estabs) %>%
  arrange(fips, year) %>%
  mutate(
    est_net = last(est) - first(est),
    est_denom = 0.5 * last(est) + 0.5 * first(est),
    est_gr = est_net / est_denom,
    emp_net = last(emp) - first(emp),
    emp_denom = 0.5 * last(emp) + 0.5 * first(emp),
    emp_gr = emp_net / emp_denom,
    churn = estabs_entry_rate + estabs_exit_rate,
    .by = fips
  ) %>%
  filter(year == y0)

d %>%
  ggplot(aes(est_gr, emp_gr, color = rural_2023)) +
  geom_point(size = 0.1) +
  geom_smooth() +
  lims(x = c(-0.5, 0.5), y = c(-0.5, 0.5))

```


```{r}
lm("emp_gr ~ est_gr * rural_2023", data = d) %>%
  summary()

lm("emp_gr ~ est_gr * rural_2023", data = d, weights = d$emp_denom) %>%
  summary()

```

## Establishment churn in base year


```{r}

d %>%
  ggplot(aes(churn, emp_gr, color = rural_2023)) +
  geom_point(size = 0.1) + geom_smooth() +
  lims(x = c(5, 30), y = c(-0.5, 0.5))


lm("emp_gr ~ churn * rural_2023", data = d) %>%
  summary()

lm("emp_gr ~ churn * rural_2023", data = d, weights = d$emp_denom) %>%
  summary()

```


# Business survival

We can use by age tables to generate survival rates of differet business cohorts over time, including COVID years.
Tables by metro/non-metro include detailed age groups with 1-year age bins up to age 5.
Are survival rates are higher in rural?

