---
title: "Rural in BDS"
format:
  html:
    toc: true
    toc-depth: 2
    embed-resources: true
    code-fold: true
    df-print: paged
---


```{r}
#| output: false

library(tidyverse)
library(patchwork)
library(tigris)
library(tmap)

theme_set(theme_minimal())

param <- list(
  rural_ruc = 4:9
)
```

# Data

```{r}
#| output: false

df_bds_met <- pubdata::get("bds", "2022_met") %>%
  mutate(rural = case_match(metro, "M" ~ "metro", "N" ~ "non-metro"))

df_bds_met_eac <- pubdata::get("bds", "2022_met_eac") %>%
  mutate(rural = case_match(metro, "M" ~ "metro", "N" ~ "non-metro"))

df_bds_met_ezc <- pubdata::get("bds", "2022_met_ezc") %>%
  mutate(rural = case_match(metro, "M" ~ "metro", "N" ~ "non-metro"))

df_bds_met_sec <- pubdata::get("bds", "2022_met_sec") %>%
  mutate(rural = case_match(metro, "M" ~ "metro", "N" ~ "non-metro"))

df_ruc_2003 <- pubdata::get("ers_rural", "ruc_2003")

df_ruc_2023 <- pubdata::get("ers_rural", "ruc_2023") %>%
  mutate(name = case_match(
    attribute,
    "Population_2020" ~ "population_2020",
    "RUCC_2023" ~ "ruc_code_2023",
    "Description" ~ "ruc_desc_2023"
    )) %>%
  pivot_wider(id_cols = c("fips", "state", "county_name")) %>%
  mutate(across(c(population_2020, ruc_code_2023), as.numeric))

df_bds_st_cty <- pubdata::get("bds", "2022_st_cty")

df_st_cty <- df_bds_st_cty %>%
  mutate(fips = paste0(st, cty), .before = "st") %>%
  left_join(df_ruc_2003 %>% select(fips, ruc_code_2003), by = "fips") %>%
  left_join(df_ruc_2023 %>% select(fips, ruc_code_2023), by = "fips") %>%
  mutate(
    rural_2003 = (ruc_code_2003 %in% param$rural_ruc),
    rural_2023 = (ruc_code_2023 %in% param$rural_ruc),
    rural_tr = case_when(
      rural_2003 & !rural_2023 ~ "fr",
      !rural_2023 ~ "urban",
      .default = "rural"
    )
  )

df_geo = tigris::counties(cb = TRUE, resolution = "20m") %>%
  rename_with(tolower) %>%
  rename(fips = geoid)

```

# --- Introduction ---

# Rurality definition

```{r}
df_ruc_2003 %>%
  summarize(counties = n(), population_2000 = sum(population_2000, na.rm = TRUE), .by = c(ruc_code_2003, ruc_desc_2003)) %>%
  arrange(ruc_code_2003)

df_ruc_2023 %>%
  summarize(counties = n(), population_2020 = sum(population_2020), .by = c(ruc_code_2023, ruc_desc_2023)) %>%
  arrange(ruc_code_2023)
```

Defining rural as "non-metro".

```{r}
param["rural_ruc"]

df_st_cty %>%
  filter(year == 2022) %>%
  count(rural_2003, rural_2023)

```


# Establishment and employment growth

From 2000 to 2019, urban greatly outperforms rural.
Rural looks worse under more recent definition.

```{r}

comp_gr <- function(rur_col, y0, y1) {
df_st_cty %>%
  filter(between(year, y0, y1)) %>%
  summarize(estabs = sum(estabs, na.rm = TRUE), emp = sum(emp, na.rm = TRUE), .by = c(year, .data[[rur_col]])) %>%
  mutate(
    est_gr = estabs / first(if_else(year == y0, estabs, NA)),
    emp_gr = emp / first(if_else(year == y0, emp, NA)),
    .by = .data[[rur_col]]
  ) %>%
  pivot_longer({{rur_col}}, names_to = "rurality", values_to = "rural")
}

bind_rows(
  comp_gr("rural_2003", 2000, 2019),
  comp_gr("rural_2023", 2000, 2019)
) %>%
  pivot_longer(c(est_gr, emp_gr)) %>%
  ggplot() + 
  geom_line(aes(year, value, color = rurality, linetype = rural)) +
  facet_wrap("name") +
  theme(legend.position = "bottom")

```

From Great Recession to Covid, rural employment growth lagged, and number of establishments remained stagnant.

```{r}

comp_gr("rural_2023", 2010, 2019) %>%
  pivot_longer(c(est_gr, emp_gr)) %>%
  ggplot() + 
  geom_line(aes(year, value, linetype = rural)) +
  facet_wrap("name") +
  theme(legend.position = "bottom")

```

# --- Results ---

# 1. Establishment and Employment Dynamism

Establishment and Employment Dynamism [2000-2022] for Metro and Non-metro Counties 

```{r}
t1 = df_bds_met %>%
  filter(year >= 2000, !is.na(rural)) %>%
  select(year, rural, estabs_entry_rate, estabs_exit_rate, job_creation_rate, job_destruction_rate) %>%
  mutate(
    estabs_churn_rate = estabs_entry_rate + estabs_exit_rate,
    job_churn_rate = job_creation_rate + job_destruction_rate
  )

t1
```

Establishment: entry, exit, and churn rates

- Establishment dynamism is declining over time.
- Rural has less dynamism, both in entry and exit.
- Dynamism increases during Covid - not just exit, but also entry go up.

```{r}
t1 %>%
  select(year, rural, starts_with("estabs")) %>%
  pivot_longer(!c(year, rural)) %>%
  ggplot() +
  geom_line(aes(year, value, color = name, linetype = rural)) +
  ylim(c(0, NA))
```


Employment: Job creation, destruction, and churn rates

- Employment dynamism shows trends similar to establishment dynamism.
- Job destruction rate was similar between urban and rural in 2011-2020.


```{r}
t1 %>%
  select(year, rural, starts_with("job")) %>%
  pivot_longer(!c(year, rural)) %>%
  ggplot() +
  geom_line(aes(year, value, color = name, linetype = rural)) +
  ylim(c(0, NA))
```

# 2. Job creation and destruction

Job creation and destruction by entrants, continuers, and exiters for Metro and Non-metro counties.

```{r}
t2 = df_bds_met %>%
  filter(year >= 2000, !is.na(rural)) %>%
  mutate(
    job_creation_rate_continuers = 100 * job_creation_continuers / denom,
    job_destruction_rate_continuers = 100 * job_destruction_continuers / denom
  )

t2 %>%
  select(year, rural, job_creation_rate, job_creation_rate_births, job_creation_rate_continuers, job_destruction_rate, job_destruction_rate_deaths, job_destruction_rate_continuers)
```

- Entry and exit contribute about 1/3 of job dynamism.


```{r}
t2 %>%
  select(year, rural, job_creation_rate_births, job_creation_rate_continuers, job_destruction_rate_deaths, job_destruction_rate_continuers) %>%
  pivot_longer(!c(year, rural)) %>%
  mutate(continuer = str_detect(name, "continuers"), creation = str_detect(name, "creation")) %>%
  mutate(creation = case_match(creation, TRUE ~ "creation", FALSE ~ "destruction"),
         continuer = case_match(continuer, TRUE ~ "continuer", FALSE ~ "entry/exit")) %>%
  ggplot() +
  geom_line(aes(year, value, color = creation, linetype = rural)) +
  ylim(c(0, NA)) +
  facet_wrap("continuer") +
  theme(legend.position = "bottom")
```


# 3. Age Distribution and Dynamism for Metro and Non-metro counties 

```{r}
t3 = df_bds_met_eac %>%
  filter(year >= 2000, !is.na(rural)) %>%
  mutate(
    job_creation_rate_continuers = 100 * job_creation_continuers / denom,
    job_destruction_rate_continuers = 100 * job_destruction_continuers / denom
  )

t3 %>% 
  relocate(
    year, rural, eagecoarse, estabs, emp,
    estabs_entry,
    estabs_entry_rate,
    estabs_exit,
    estabs_exit_rate,
    contains("job")
  )
```

## establishments

- Most establishments are 11+ years old.
- Age distribution is shifting towards higher age over time.
- Especially in rural areas.
- "Left Censored" can be combined with 11+ group.

```{r}
x = t3 %>%
  select(year, rural, eagecoarse, estabs) %>%
  pivot_longer(!c(year, rural, eagecoarse)) %>%
  filter(name == "estabs", year %in% c(2000, 2010, 2020, 2022)) %>%
  mutate(year = as.factor(year))

x %>%
  ggplot() +
  geom_col(aes(eagecoarse, value, fill = year), position = "dodge") +
  facet_wrap("rural", nrow = 1, scales = "free_y") +
  labs(title = "Number of establishments") +
  theme(legend.position = "bottom")

x %>%
  mutate(eagecoarse = if_else(eagecoarse == "e) Left Censored", "d) 11+", eagecoarse)) %>%
  summarize(value = sum(value, na.rm = TRUE), .by = c(year, rural, eagecoarse)) %>%
  ggplot() +
  geom_col(aes(eagecoarse, value, fill = year), position = "dodge") +
  facet_wrap("rural", nrow = 1, scales = "free_y") +
  labs(title = "Number of establishments") +
  theme(legend.position = "bottom")

```

- Entry happens even at 1+ age (re-entry).

```{r}
x = t3 %>%
  select(year, rural, eagecoarse, estabs_entry, estabs_exit) %>%
  pivot_longer(!c(year, rural, eagecoarse)) %>%
  mutate(eagecoarse = if_else(eagecoarse == "e) Left Censored", "d) 11+", eagecoarse)) %>%
  summarize(value = sum(value, na.rm = TRUE), .by = c(year, rural, eagecoarse, name))

p1 = x %>%
  filter(rural == "metro") %>%
  ggplot() +
  geom_line(aes(year, value, color = name)) +
  facet_wrap("eagecoarse", nrow = 1) +
  labs(title = "Number of entering and exiting establishments by age group") +
  theme(legend.position = "none") +
  ylab("metro")

p2 = x %>%
  filter(rural == "non-metro") %>%
  ggplot() +
  geom_line(aes(year, value, color = name)) +
  facet_wrap("eagecoarse", nrow = 1) +
  theme(legend.position = "bottom") +
  ylab("non-metro")

p1 / p2
```

- Dynamism declines with establishment age.
- Although overall entry rate is lower in rural, re-entry of age 1+ establishments is slightly higher.

```{r, fig.width=10, fig.height=5}
t3 %>%
  select(year, rural, eagecoarse, estabs_entry_rate, estabs_exit_rate) %>%
  pivot_longer(c(estabs_entry_rate, estabs_exit_rate)) %>%
  filter(value < 200) %>%
  ggplot() +
  geom_line(aes(year, value, linetype = rural, color = name)) +
  ylim(c(0, NA)) +
  facet_wrap("eagecoarse", nrow = 1) +
  theme(legend.position = "bottom") +
  labs(title = "Entry and exit rate by establishment age group")


```


## jobs

- Most jobs are in the 11+ age group, and start-ups have smallest number of jobs.
- All net job creation is happening among startups.
- All age groups were net creators of jobs in urban in 2022, but not in rural.

```{r}
x = t3 %>%
  select(year, rural, eagecoarse, emp, net_job_creation) %>%
  pivot_longer(!c(year, rural, eagecoarse)) %>%
  filter(year %in% c(2000, 2010, 2020, 2022)) %>%
  mutate(eagecoarse = if_else(eagecoarse == "e) Left Censored", "d) 11+", eagecoarse)) %>%
  summarize(value = sum(value, na.rm = TRUE), .by = c(year, rural, eagecoarse, name)) %>%
  mutate(year = as.factor(year))

x %>%
  ggplot() +
  geom_col(aes(eagecoarse, value, fill = year), position = "dodge") +
  facet_wrap(c("name", "rural"), nrow = 2, scales = "free_y") +
  labs(title = "Number of jobs") +
  theme(legend.position = "bottom")
```


```{r}
x = t3 %>%
  select(year, rural, eagecoarse, job_creation, job_destruction) %>%
  pivot_longer(!c(year, rural, eagecoarse)) %>%
  mutate(eagecoarse = if_else(eagecoarse == "e) Left Censored", "d) 11+", eagecoarse)) %>%
  summarize(value = sum(value, na.rm = TRUE), .by = c(year, rural, eagecoarse, name))

p1 = x %>%
  filter(rural == "metro") %>%
  ggplot() +
  geom_line(aes(year, value, color = name)) +
  facet_wrap("eagecoarse", nrow = 1) +
  labs(title = "Jon creation and destruction by age group") +
  theme(legend.position = "none") +
  ylab("metro")

p2 = x %>%
  filter(rural == "non-metro") %>%
  ggplot() +
  geom_line(aes(year, value, color = name)) +
  facet_wrap("eagecoarse", nrow = 1) +
  theme(legend.position = "bottom") +
  ylab("non-metro")

p1 / p2
```

- Job creation and destruction rates are declining over time.

```{r, fig.width=10, fig.heigth=8}
t3 %>%
  select(year, rural, eagecoarse, matches("job.*rate_")) %>%
  pivot_longer(!c(year, rural, eagecoarse)) %>%
  mutate(continuer = str_detect(name, "continuers"), creation = str_detect(name, "creation")) %>%
  mutate(creation = case_match(creation, TRUE ~ "creation", FALSE ~ "destruction"),
         continuer = case_match(continuer, TRUE ~ "continuer", FALSE ~ "entry/exit")) %>%
  filter(eagecoarse != "a) 0") %>%
  ggplot() +
  geom_line(aes(year, value, linetype = rural, color = creation)) +
  ylim(c(0, NA)) +
  facet_wrap(c("continuer", "eagecoarse"), nrow = 2) +
  theme(legend.position = "bottom")
```

# 4. Size Distribution and Dynamism for Metro and Non-metro counties 

```{r}
t4 = df_bds_met_ezc %>%
  filter(year >= 2000, !is.na(rural))

t4 %>% 
  relocate(
    year, rural, esizecoarse, estabs, emp,
    estabs_entry,
    estabs_entry_rate,
    estabs_exit,
    estabs_exit_rate,
    contains("job")
  )
```

## establishments

```{r}
x = t4 %>%
  select(year, rural, esizecoarse, estabs) %>%
  pivot_longer(!c(year, rural, esizecoarse)) %>%
  filter(name == "estabs", year %in% c(2000, 2010, 2020, 2022)) %>%
  mutate(year = as.factor(year))

x %>%
  ggplot() +
  geom_col(aes(esizecoarse, value, fill = year), position = "dodge") +
  facet_wrap("rural", nrow = 1, scales = "free_y") +
  labs(title = "Number of establishments") +
  theme(legend.position = "bottom")


```



## jobs

```{r}
x = t4 %>%
  select(year, rural, esizecoarse, emp, net_job_creation) %>%
  pivot_longer(!c(year, rural, esizecoarse)) %>%
  filter(year %in% c(2000, 2010, 2020, 2022)) %>%
  mutate(year = as.factor(year))

x %>%
  ggplot() +
  geom_col(aes(esizecoarse, value, fill = year), position = "dodge") +
  facet_wrap(c("name", "rural"), nrow = 2, scales = "free_y") +
  labs(title = "Number of jobs") +
  theme(legend.position = "bottom")
```



# 5. Sectoral Distribution and Dynamism for Metro and Non-metro 

2-digit NAICS.
Too many sectors to report everything?
Which should we merge together?

```{r}
df_bds_met_sec %>%
  filter(year == 2022, !is.na(rural)) %>%
  select(year, sector, rural, firms, estabs, emp, estabs_entry_rate, estabs_exit_rate, net_job_creation_rate) %>%
  arrange(year, sector, rural)
```


# 6. Maps: County-level Dynamism for Metro and Non-metro

Which variables are most interesting in the spatial context?

```{r}
y = 2022

t6 = df_geo %>%
  inner_join(df_st_cty %>% filter(year == y), by = "fips") %>%
  shift_geometry()

f = function(x, title) {
  tm_shape(x) +
  tm_polygons(
    "net_job_creation_rate", 
    midpoint = 0, 
    breaks = c(-100, -25, -10, 0, 10, 25, 100),
    palette = "PRGn",
    border.alpha = 0.3,
    title = title)
}

t6 %>%
  f("all counties")

t6 %>%
  filter(!rural_2023) %>%
  f("metro")

t6 %>%
  filter(rural_2023) %>%
  f("non-metro")

  
```



# --- Extras ---


# Job creation and destruction

Rural employment is less dynamic: both job creation and destruction rates are below urban.

```{r}
y0 = 2010
y1 = 2019

d <- df_st_cty %>%
  filter(between(year, y0, y1)) %>%
  summarize(across(
    c(denom, job_creation, job_creation_births, job_creation_continuers, job_destruction, job_destruction_deaths, job_destruction_continuers),
    \(x) sum(x, na.rm = TRUE)), 
    .by = c(year, rural_2023)) %>%
  mutate(across(
    c(job_creation, job_creation_births, job_creation_continuers, job_destruction, job_destruction_deaths, job_destruction_continuers),
    \(x) x / denom))

d %>%
  pivot_longer(c(job_creation, job_destruction)) %>%
  ggplot() +
  geom_line(aes(year, value, color = name, linetype = rural_2023)) +
  ylim(c(0, NA))

```


Job destruction rate by continuing establishments in rural is similar to urban, but creation is much lower.
Net creation by continuers is thus much higher in urban than in rural.

```{r}
d %>%
  pivot_longer(c(job_creation_births, job_creation_continuers, job_destruction_deaths, job_destruction_continuers)) %>%
  mutate(name = ordered(name, c("job_creation_births", "job_destruction_deaths", "job_creation_continuers", "job_destruction_continuers"))) %>%
  ggplot() +
  geom_line(aes(year, value, linetype = rural_2023)) +
  ylim(c(0, NA)) +
  facet_wrap("name", ncol = 2)


```

"Creative destruction" (birth and death) accounts for about a third of all job dynamism in both rural and urban.
In other words, about half as many jobs are created and destroyed by entry and exit.

If we combine young establishments with birth, the role of start-ups in job creation will be even higher. 
*No figure here to support this claim, but it is almost certainly true.*


```{r}
d %>%
  pivot_longer(c(job_creation_births, job_creation_continuers, job_destruction_deaths, job_destruction_continuers)) %>%
  mutate(continuer = str_detect(name, "continuers"), creation = str_detect(name, "creation")) %>%
  mutate(creation = case_match(creation, TRUE ~ "creation", FALSE ~ "destruction"),
         continuer = case_match(continuer, TRUE ~ "continuer", FALSE ~ "death birth"),
         rural_2023 = case_match(rural_2023, TRUE ~ "rural", FALSE ~ "urban")) %>%
  ggplot() +
  geom_line(aes(year, value, color = creation, linetype = continuer)) +
  ylim(c(0, NA)) +
  facet_wrap("rural_2023", ncol = 2)

```

# Creative destruction across counties

Looking at un-weighted correlation between county establishment and employment growth rate from 2010 to 2019, employment dynamism plays smaller role in job creation in rural areas.
But if we weight counties by employment, rural effect becomes positive.

```{r}
d <- df_st_cty %>%
  filter(year %in% c(y0, y1)) %>%
  rename(est = estabs) %>%
  arrange(fips, year) %>%
  mutate(
    est_net = last(est) - first(est),
    est_denom = 0.5 * last(est) + 0.5 * first(est),
    est_gr = est_net / est_denom,
    emp_net = last(emp) - first(emp),
    emp_denom = 0.5 * last(emp) + 0.5 * first(emp),
    emp_gr = emp_net / emp_denom,
    churn = estabs_entry_rate + estabs_exit_rate,
    .by = fips
  ) %>%
  filter(year == y0)

d %>%
  ggplot(aes(est_gr, emp_gr, color = rural_2023)) +
  geom_point(size = 0.1) +
  geom_smooth() +
  lims(x = c(-0.5, 0.5), y = c(-0.5, 0.5))

```


```{r}
lm("emp_gr ~ est_gr * rural_2023", data = d) %>%
  summary()

lm("emp_gr ~ est_gr * rural_2023", data = d, weights = d$emp_denom) %>%
  summary()

```

## Establishment churn in base year


```{r}

d %>%
  ggplot(aes(churn, emp_gr, color = rural_2023)) +
  geom_point(size = 0.1) + geom_smooth() +
  lims(x = c(5, 30), y = c(-0.5, 0.5))


lm("emp_gr ~ churn * rural_2023", data = d) %>%
  summary()

lm("emp_gr ~ churn * rural_2023", data = d, weights = d$emp_denom) %>%
  summary()

```


# Business survival

We can use by age tables to generate survival rates of differet business cohorts over time, including COVID years.
Tables by metro/non-metro include detailed age groups with 1-year age bins up to age 5.
Are survival rates are higher in rural?

```{r}
x = pubdata::get("bds", "2022_met_ea")
x %>%
  filter(metro == "M") %>%
  pivot_wider(id_cols = year, names_from = eage, values_from = estabs)
```

