
---
title: "&nbsp;"
format:
  html:
    self-contained: true
    page-layout: full
    code-fold: true
    code-tools: true
    code_download: yes
    latex_engine: pdflatex
---

```{r packages and libraries, include = FALSE}

############ Load and attach necessary packages
library(rprojroot)

############ Load custom functions
source(file.path(find_rstudio_root_file(), "nbs", "rural_typology_r_functions.R"))

```

```{r}
cbp_year = "2012"
ilevel = "det"

impedance_mat = gaus_impedance_mat(tiger_year = year2tiger(cbp_year), 
                                    rms_width = miles2meters(200))
impedance_call = "gaus200"

  fd <- industry_factor_demand_tidy_matrix(cbp_year = cbp_year, ilevel = ilevel)
  go <- industry_output_tidy_matrix(cbp_year = cbp_year, ilevel = ilevel)
  fs <- industry_factor_supply_tidy_matrix(cbp_year = cbp_year, ilevel = ilevel)
  
  y <- names(which(!is.na(rowSums(fs)) & !rowSums(fs) == 0 ))

  if (file.exists(file.path(find_rstudio_root_file(), "data", "robjs", "df_a"))){ 
    df_a <- readRDS(file.path(find_rstudio_root_file(), "data", "robjs", "df_a"))
  } else {
    df_a <- normalized_absorption_share(stacked_absorption_share(net_input_supply(go, fd), net_input_demand(go, fd)), net_input_supply(go, fd)) 
    saveRDS(df_a, file = file.path(find_rstudio_root_file(), "data", "robjs", "df_a"))
  }
  
    if (file.exists(file.path(find_rstudio_root_file(), "data", "robjs", "df_b"))){ 
    df_b <- readRDS(file.path(find_rstudio_root_file(), "data", "robjs", "df_b"))
  } else {
    df_b <- normalized_absorption_share(stacked_absorption_share(net_input_supply(fs[y, , drop=F], fd[y, , drop=F]), net_input_demand(fs[y, , drop=F], fd[y, , drop=F])), net_input_supply(fs[y, , drop=F], fd[y, , drop=F])) 
    saveRDS(df_b, file = file.path(find_rstudio_root_file(), "data", "robjs", "df_b"))
  }
  
    if (file.exists(file.path(find_rstudio_root_file(), "data", "robjs", "df_c"))){ 
    df_c <- readRDS(file.path(find_rstudio_root_file(), "data", "robjs", "df_c"))
  } else {
    df_c <- Reduce('+', load_tradeflows(cbp_year = cbp_year, ilevel = ilevel, impedance_call = impedance_call, impedance_mat = impedance_mat) ) %>% sweep( 1, rowSums(.), FUN="/")
    saveRDS(df_c, file = file.path(find_rstudio_root_file(), "data", "robjs", "df_c"))
  }
  
  

```

#Density plot of a and b (row) max absorption (normalized) values 
```{r}
  df <- rbind(noimpedance_absorption_maximum(df_a, impedance_mat = impedance_mat), 
              noimpedance_absorption_maximum(df_b, impedance_mat = impedance_mat))
ggplot(df, aes(x = ab_max, color = id)) + geom_density()

```
#Scatter plot of (log) industry gross output to a and b (row) max absorption (normalized) values
```{r}

 outp <- colSums(industry_output_tidy_matrix(cbp_year = cbp_year, ilevel = ilevel))
  df <- rbind(cbind(noimpedance_absorption_maximum(df_a, impedance_mat = impedance_mat), 
                    outp),
              cbind(noimpedance_absorption_maximum(df_b, impedance_mat = impedance_mat), 
                    outp))
ggplot(df, aes(x = ab_max, y = log(outp), color = id)) + geom_point()

```
#Scatter plot of b to a (row) max absorption (normalized) values by county
```{r}
  df <- cbind(noimpedance_absorption_maximum(df_a, impedance_mat = impedance_mat), 
              noimpedance_absorption_maximum(df_b, impedance_mat = impedance_mat), 
              outp)

colnames(df) <- c("a_abs", "id", "b_abs", "id2", "outp")
g <- ggplot(df, aes(x = a_abs, y = b_abs, tooltip = rownames(df), data_id = rownames(df))) + geom_point_interactive()

girafe(ggobj = g)

```



#Density plot of a, b and c (row) max absorption (normalized) values 
```{r}

  df <- noimpedance_absorption_maximum(df_a, impedance_mat = impedance_mat)
  df <- rbind(df, noimpedance_absorption_maximum(df_b, impedance_mat = impedance_mat))
  df <- rbind(df, noimpedance_absorption_maximum(df_c, impedance_mat = matrix(data = 1, nrow = dim(df_c)[1], ncol = dim(df_c)[2], dimnames = list(rownames(df_c), colnames(df_c))) ))
ggplot(df, aes(x = ab_max, color = id)) + geom_density()

```
#Scatter plot of (log) industry gross output to a, b and c (row) max absorption (normalized) values
```{r}

  df <- rbind(cbind(noimpedance_absorption_maximum(df_a, impedance_mat = impedance_mat), 
                    outp),
              cbind(noimpedance_absorption_maximum(df_b, impedance_mat = impedance_mat), 
                    outp),
              cbind(noimpedance_absorption_maximum(df_c, impedance_mat = matrix(data = 1, nrow = dim(df_c)[1], ncol = dim(df_c)[2], dimnames = list(rownames(df_c), colnames(df_c)))), 
                    outp))
ggplot(df, aes(x = ab_max, y = log(outp), color = id)) + geom_point(alpha = 1/5)

```

#Scatter plot of c to a (row) max absorption (normalized) values by county
```{r}
  df <- cbind(noimpedance_absorption_maximum(df_a, impedance_mat = impedance_mat), 
              noimpedance_absorption_maximum(df_c, impedance_mat = matrix(data = 1, nrow = dim(df_c)[1], ncol = dim(df_c)[2], dimnames = list(rownames(df_c), colnames(df_c)))), 
              outp)

colnames(df) <- c("a_abs", "id", "c_abs", "id2", "outp")
g <- ggplot(df, aes(x = a_abs, y = c_abs, tooltip = rownames(df), data_id = rownames(df))) + geom_point_interactive()

girafe(ggobj = g)

```

#Scatter plot of c to b (row) max absorption (normalized) values by county
```{r}
  df <- cbind(noimpedance_absorption_maximum(df_b, impedance_mat = impedance_mat), 
              noimpedance_absorption_maximum(df_c, impedance_mat = matrix(data = 1, nrow = dim(df_c)[1], ncol = dim(df_c)[2], dimnames = list(rownames(df_c), colnames(df_c)))), 
              outp)

colnames(df) <- c("b_abs", "id", "c_abs", "id2", "outp")
g <- ggplot(df, aes(x = b_abs, y = c_abs, tooltip = rownames(df), data_id = rownames(df))) + geom_point_interactive()

girafe(ggobj = g)

```

```{r}

if (file.exists(file.path(find_rstudio_root_file(), "data", "robjs", "abc_top"))){ 
    df <- readRDS(file.path(find_rstudio_root_file(), "data", "robjs", "abc_top"))
  } else {

  df_a <- df_a * impedance_mat[rownames(df_a), colnames(df_a)]
  df_b <- df_b * impedance_mat[rownames(df_b), colnames(df_b)]
  
  tempr <- list()
for(i in 1:10){
  tempr[[i]] <- as.matrix(apply(df_a, 1, function(x){sort(x, decreasing = T)[i] }))
}

tempc <- list()
for(i in 1:10){
  tempc[[i]] <- as.matrix(apply(df_a, 2, function(x){sort(x, decreasing = T)[i] }))
}

      abs_a <- cbind(place = rownames(df_a), 
                     match_out_1 = colnames(df_a)[apply(df_a, 1, function(x){which(x %in% tempr[[1]])})],
                      match_out_2 = colnames(df_a)[apply(df_a, 1, function(x){which(x %in% tempr[[2]])})],
                      match_out_3 = colnames(df_a)[apply(df_a, 1, function(x){which(x %in% tempr[[3]])})],
                      match_out_4 = colnames(df_a)[apply(df_a, 1, function(x){which(x %in% tempr[[4]])})],
                      match_out_5 = colnames(df_a)[apply(df_a, 1, function(x){which(x %in% tempr[[5]])})],
                      match_out_6 = colnames(df_a)[apply(df_a, 1, function(x){which(x %in% tempr[[6]])})],
                      match_out_7 = colnames(df_a)[apply(df_a, 1, function(x){which(x %in% tempr[[7]])})],
                      match_out_8 = colnames(df_a)[apply(df_a, 1, function(x){which(x %in% tempr[[8]])})],
                      match_out_9 = colnames(df_a)[apply(df_a, 1, function(x){which(x %in% tempr[[9]])})],
                      match_out_10 = colnames(df_a)[apply(df_a, 1, function(x){which(x %in% tempr[[10]])})],
                     match_in_1 = rownames(df_a)[apply(df_a, 2, function(x){which(x %in% tempc[[1]])})],
                      match_in_2 = rownames(df_a)[apply(df_a, 2, function(x){which(x %in% tempc[[2]])})],
                      match_in_3 = rownames(df_a)[apply(df_a, 2, function(x){which(x %in% tempc[[3]])})],
                      match_in_4 = rownames(df_a)[apply(df_a, 2, function(x){which(x %in% tempc[[4]])})],
                      match_in_5 = rownames(df_a)[apply(df_a, 2, function(x){which(x %in% tempc[[5]])})],
                      match_in_6 = rownames(df_a)[apply(df_a, 2, function(x){which(x %in% tempc[[6]])})],
                      match_in_7 = rownames(df_a)[apply(df_a, 2, function(x){which(x %in% tempc[[7]])})],
                      match_in_8 = rownames(df_a)[apply(df_a, 2, function(x){which(x %in% tempc[[8]])})],
                      match_in_9 = rownames(df_a)[apply(df_a, 2, function(x){which(x %in% tempc[[9]])})],
                      match_in_10 = rownames(df_a)[apply(df_a, 2, function(x){which(x %in% tempc[[10]])})]
                     )
      
      
        
  tempr <- list()
for(i in 1:10){
  tempr[[i]] <- as.matrix(apply(df_b, 1, function(x){sort(x, decreasing = T)[i] }))
}

tempc <- list()
for(i in 1:10){
  tempc[[i]] <- as.matrix(apply(df_b, 2, function(x){sort(x, decreasing = T)[i] }))
}

      abs_b <- cbind(place = rownames(df_b), 
                     match_out_1 = colnames(df_b)[apply(df_b, 1, function(x){which(x %in% tempr[[1]])})],
                      match_out_2 = colnames(df_b)[apply(df_b, 1, function(x){which(x %in% tempr[[2]])})],
                      match_out_3 = colnames(df_b)[apply(df_b, 1, function(x){which(x %in% tempr[[3]])})],
                      match_out_4 = colnames(df_b)[apply(df_b, 1, function(x){which(x %in% tempr[[4]])})],
                      match_out_5 = colnames(df_b)[apply(df_b, 1, function(x){which(x %in% tempr[[5]])})],
                      match_out_6 = colnames(df_b)[apply(df_b, 1, function(x){which(x %in% tempr[[6]])})],
                      match_out_7 = colnames(df_b)[apply(df_b, 1, function(x){which(x %in% tempr[[7]])})],
                      match_out_8 = colnames(df_b)[apply(df_b, 1, function(x){which(x %in% tempr[[8]])})],
                      match_out_9 = colnames(df_b)[apply(df_b, 1, function(x){which(x %in% tempr[[9]])})],
                      match_out_10 = colnames(df_b)[apply(df_b, 1, function(x){which(x %in% tempr[[10]])})],
                     match_in_1 = rownames(df_b)[apply(df_b, 2, function(x){which(x %in% tempc[[1]])})],
                      match_in_2 = rownames(df_b)[apply(df_b, 2, function(x){which(x %in% tempc[[2]])})],
                      match_in_3 = rownames(df_b)[apply(df_b, 2, function(x){which(x %in% tempc[[3]])})],
                      match_in_4 = rownames(df_b)[apply(df_b, 2, function(x){which(x %in% tempc[[4]])})],
                      match_in_5 = rownames(df_b)[apply(df_b, 2, function(x){which(x %in% tempc[[5]])})],
                      match_in_6 = rownames(df_b)[apply(df_b, 2, function(x){which(x %in% tempc[[6]])})],
                      match_in_7 = rownames(df_b)[apply(df_b, 2, function(x){which(x %in% tempc[[7]])})],
                      match_in_8 = rownames(df_b)[apply(df_b, 2, function(x){which(x %in% tempc[[8]])})],
                      match_in_9 = rownames(df_b)[apply(df_b, 2, function(x){which(x %in% tempc[[9]])})],
                      match_in_10 = rownames(df_b)[apply(df_b, 2, function(x){which(x %in% tempc[[10]])})]
                     )
      
  tempr <- list()
for(i in 1:10){
  tempr[[i]] <- as.matrix(apply(df_c, 1, function(x){sort(x, decreasing = T)[i] }))
}

tempc <- list()
for(i in 1:10){
  tempc[[i]] <- as.matrix(apply(df_c, 2, function(x){sort(x, decreasing = T)[i] }))
}

      abs_c <- cbind(place = rownames(df_c), 
                     match_out_1 = colnames(df_c)[apply(df_c, 1, function(x){which(x %in% tempr[[1]])})],
                      match_out_2 = colnames(df_c)[apply(df_c, 1, function(x){which(x %in% tempr[[2]])})],
                      match_out_3 = colnames(df_c)[apply(df_c, 1, function(x){which(x %in% tempr[[3]])})],
                      match_out_4 = colnames(df_c)[apply(df_c, 1, function(x){which(x %in% tempr[[4]])})],
                      match_out_5 = colnames(df_c)[apply(df_c, 1, function(x){which(x %in% tempr[[5]])})],
                      match_out_6 = colnames(df_c)[apply(df_c, 1, function(x){which(x %in% tempr[[6]])})],
                      match_out_7 = colnames(df_c)[apply(df_c, 1, function(x){which(x %in% tempr[[7]])})],
                      match_out_8 = colnames(df_c)[apply(df_c, 1, function(x){which(x %in% tempr[[8]])})],
                      match_out_9 = colnames(df_c)[apply(df_c, 1, function(x){which(x %in% tempr[[9]])})],
                      match_out_10 = colnames(df_c)[apply(df_c, 1, function(x){which(x %in% tempr[[10]])})],
                     match_in_1 = rownames(df_c)[apply(df_c, 2, function(x){which(x %in% tempc[[1]])})],
                      match_in_2 = rownames(df_c)[apply(df_c, 2, function(x){which(x %in% tempc[[2]])})],
                      match_in_3 = rownames(df_c)[apply(df_c, 2, function(x){which(x %in% tempc[[3]])})],
                      match_in_4 = rownames(df_c)[apply(df_c, 2, function(x){which(x %in% tempc[[4]])})],
                      match_in_5 = rownames(df_c)[apply(df_c, 2, function(x){which(x %in% tempc[[5]])})],
                      match_in_6 = rownames(df_c)[apply(df_c, 2, function(x){which(x %in% tempc[[6]])})],
                      match_in_7 = rownames(df_c)[apply(df_c, 2, function(x){which(x %in% tempc[[7]])})],
                      match_in_8 = rownames(df_c)[apply(df_c, 2, function(x){which(x %in% tempc[[8]])})],
                      match_in_9 = rownames(df_c)[apply(df_c, 2, function(x){which(x %in% tempc[[9]])})],
                      match_in_10 = rownames(df_c)[apply(df_c, 2, function(x){which(x %in% tempc[[10]])})]
                     )

      colnames(abs_a)[2:21] <- paste(colnames(abs_a)[2:21], "a", sep = '.')
      colnames(abs_b)[2:21] <- paste(colnames(abs_b)[2:21], "b", sep = '.')
      colnames(abs_c)[2:21] <- paste(colnames(abs_c)[2:21], "c", sep = '.')

df <- inner_join(inner_join(as.data.frame(abs_a), as.data.frame(abs_b), by = "place"), as.data.frame(abs_c), by = "place")
saveRDS(df, file = file.path(find_rstudio_root_file(), "data", "robjs", "abc_top"))
}

```

#Bar plot of top 1 max absorption (normalized) location overlap
```{r}

barplot(table(df$match_out_1.a == df$match_out_1.b))
barplot(table(df$match_out_1.a == df$match_out_1.c))
barplot(table(df$match_out_1.b == df$match_out_1.c))

barplot(table(df$match_in_1.a == df$match_in_1.b))
barplot(table(df$match_in_1.a == df$match_in_1.c))
barplot(table(df$match_in_1.b == df$match_in_1.c))

```
#Bar plot of top 4 max absorption (normalized) location overlap
```{r}


a_b <- list()
for(i in 1:dim(df)[1]){
  a_b[[i]] <- intersect(cbind(df$match_out_1.a, df$match_out_2.a, df$match_out_3.a, df$match_out_4.a)[i,] , cbind(df$match_out_1.b, df$match_out_2.b, df$match_out_3.b, df$match_out_4.b)[i,]) %>% length()
}    
a_b <- as.data.frame(t(data.frame(a_b)))
colnames(a_b) <- "a_b"
ggplot(a_b, aes(x=a_b)) + geom_bar()
a_c <- list()
for(i in 1:dim(df)[1]){
  a_c[[i]] <- intersect(cbind(df$match_out_1.a, df$match_out_2.a, df$match_out_3.a, df$match_out_4.a)[i,] , cbind(df$match_out_1.c, df$match_out_2.c, df$match_out_3.c, df$match_out_4.c)[i,]) %>% length()
}    
a_c <- as.data.frame(t(data.frame(a_c)))
colnames(a_c) <- "a_c"
ggplot(a_c, aes(x=a_c)) + geom_bar()
b_c <- list()
for(i in 1:dim(df)[1]){
  b_c[[i]] <- intersect(cbind(df$match_out_1.b, df$match_out_2.b, df$match_out_3.b, df$match_out_4.b)[i,] , cbind(df$match_out_1.c, df$match_out_2.c, df$match_out_3.c, df$match_out_4.c)[i,]) %>% length()
}    
b_c <- as.data.frame(t(data.frame(b_c)))
colnames(b_c) <- "b_c"
ggplot(b_c, aes(x=b_c)) + geom_bar()


a_b <- list()
for(i in 1:dim(df)[1]){
  a_b[[i]] <- intersect(cbind(df$match_in_1.a, df$match_in_2.a, df$match_in_3.a, df$match_in_4.a)[i,] , cbind(df$match_in_1.b, df$match_in_2.b, df$match_in_3.b, df$match_in_4.b)[i,]) %>% length()
}    
a_b <- as.data.frame(t(data.frame(a_b)))
colnames(a_b) <- "a_b"
ggplot(a_b, aes(x=a_b)) + geom_bar()
a_c <- list()
for(i in 1:dim(df)[1]){
  a_c[[i]] <- intersect(cbind(df$match_in_1.a, df$match_in_2.a, df$match_in_3.a, df$match_in_4.a)[i,] , cbind(df$match_in_1.c, df$match_in_2.c, df$match_in_3.c, df$match_in_4.c)[i,]) %>% length()
}    
a_c <- as.data.frame(t(data.frame(a_c)))
colnames(a_c) <- "a_c"
ggplot(a_c, aes(x=a_c)) + geom_bar()
b_c <- list()
for(i in 1:dim(df)[1]){
  b_c[[i]] <- intersect(cbind(df$match_in_1.b, df$match_in_2.b, df$match_in_3.b, df$match_in_4.b)[i,] , cbind(df$match_in_1.c, df$match_in_2.c, df$match_in_3.c, df$match_in_4.c)[i,]) %>% length()
}    
b_c <- as.data.frame(t(data.frame(b_c)))
colnames(b_c) <- "b_c"
ggplot(b_c, aes(x=b_c)) + geom_bar()


```
#Bar plot of top 10 max absorption (normalized) location overlap
```{r}



a_b <- list()
for(i in 1:dim(df)[1]){
  a_b[[i]] <- intersect(cbind(df$match_out_1.a, df$match_out_2.a, df$match_out_3.a, df$match_out_4.a, df$match_out_5.a, df$match_out_6.a, df$match_out_7.a, df$match_out_8.a, df$match_out_9.a, df$match_out_10.a)[i,] , cbind(df$match_out_1.b, df$match_out_2.b, df$match_out_3.b, df$match_out_4.b, df$match_out_5.b, df$match_out_6.b, df$match_out_7.b, df$match_out_8.b, df$match_out_9.b, df$match_out_10.b)[i,]) %>% length()
}    
a_b <- as.data.frame(t(data.frame(a_b)))
colnames(a_b) <- "a_b"
ggplot(a_b, aes(x=a_b)) + geom_bar()
a_c <- list()
for(i in 1:dim(df)[1]){
  a_c[[i]] <- intersect(cbind(df$match_out_1.a, df$match_out_2.a, df$match_out_3.a, df$match_out_4.a, df$match_out_5.a, df$match_out_6.a, df$match_out_7.a, df$match_out_8.a, df$match_out_9.a, df$match_out_10.a)[i,] , cbind(df$match_out_1.c, df$match_out_2.c, df$match_out_3.c, df$match_out_4.c, df$match_out_5.c, df$match_out_6.c, df$match_out_7.c, df$match_out_8.c, df$match_out_9.c, df$match_out_10.c)[i,]) %>% length()
}    
a_c <- as.data.frame(t(data.frame(a_c)))
colnames(a_c) <- "a_c"
ggplot(a_c, aes(x=a_c)) + geom_bar()
b_c <- list()
for(i in 1:dim(df)[1]){
  b_c[[i]] <- intersect(cbind(df$match_out_1.b, df$match_out_2.b, df$match_out_3.b, df$match_out_4.b, df$match_out_5.b, df$match_out_6.b, df$match_out_7.b, df$match_out_8.b, df$match_out_9.b, df$match_out_10b_c)[i,] , cbind(df$match_out_1.c, df$match_out_2.c, df$match_out_3.c, df$match_out_4.c, df$match_out_5.c, df$match_out_6.c, df$match_out_7.c, df$match_out_8.c, df$match_out_9.c, df$match_out_10.c)[i,]) %>% length()
}    
b_c <- as.data.frame(t(data.frame(b_c)))
colnames(b_c) <- "b_c"
ggplot(b_c, aes(x=b_c)) + geom_bar()


a_b <- list()
for(i in 1:dim(df)[1]){
  a_b[[i]] <- intersect(cbind(df$match_in_1.a, df$match_in_2.a, df$match_in_3.a, df$match_in_4.a, df$match_in_5.a, df$match_in_6.a, df$match_in_7.a, df$match_in_8.a, df$match_in_9.a, df$match_in_10.a)[i,] , cbind(df$match_in_1.b, df$match_in_2.b, df$match_in_3.b, df$match_in_4.b, df$match_in_5.b, df$match_in_6.b, df$match_in_7.b, df$match_in_8.b, df$match_in_9.b, df$match_in_10.b)[i,]) %>% length()
}    
a_b <- as.data.frame(t(data.frame(a_b)))
colnames(a_b) <- "a_b"
ggplot(a_b, aes(x=a_b)) + geom_bar()
a_c <- list()
for(i in 1:dim(df)[1]){
  a_c[[i]] <- intersect(cbind(df$match_in_1.a, df$match_in_2.a, df$match_in_3.a, df$match_in_4.a, df$match_in_5.a, df$match_in_6.a, df$match_in_7.a, df$match_in_8.a, df$match_in_9.a, df$match_in_10.a)[i,] , cbind(df$match_in_1.c, df$match_in_2.c, df$match_in_3.c, df$match_in_4.c, df$match_in_5.c, df$match_in_6.c, df$match_in_7.c, df$match_in_8.c, df$match_in_9.c, df$match_in_10.c)[i,]) %>% length()
}    
a_c <- as.data.frame(t(data.frame(a_c)))
colnames(a_c) <- "a_c"
ggplot(a_c, aes(x=a_c)) + geom_bar()
b_c <- list()
for(i in 1:dim(df)[1]){
  b_c[[i]] <- intersect(cbind(df$match_in_1.b, df$match_in_2.b, df$match_in_3.b, df$match_in_4.b, df$match_in_5.b, df$match_in_6.b, df$match_in_7.b, df$match_in_8.b, df$match_in_9.b, df$match_in_10b_c)[i,] , cbind(df$match_in_1.c, df$match_in_2.c, df$match_in_3.c, df$match_in_4.c, df$match_in_5.c, df$match_in_6.c, df$match_in_7.c, df$match_in_8.c, df$match_in_9.c, df$match_in_10.c)[i,]) %>% length()
}    
b_c <- as.data.frame(t(data.frame(b_c)))
colnames(b_c) <- "b_c"
ggplot(b_c, aes(x=b_c)) + geom_bar()



```
