---
title: "BEA I-O tables"
format: html
---


```{r}
source("R/bea_io.R", local = (bea_io <- new.env()))
```


# Combining industries

## sector

Sector tables are more aggregated than sectors in NAICS concordance.
To align concordance with sector-level tables, we combine codes in the concordance table to match aggregated codes.

Commodity only: Scrap, used and secondhand goods; Noncomparable imports and rest-of-the-world adjustment.

```{r}
tab_com <- pubdata::bea_io_get("2022_sup_sec_2012") %>%
  filter(core_matrix) %>%
  select(code = row_code, commodity = row_name) %>%
  distinct()
tab_ind <- pubdata::bea_io_get("2022_sup_sec_2012") %>%
  filter(core_matrix) %>%
  select(code = col_code, industry = col_name) %>%
  distinct()
conc <- pubdata::bea_io_get("2022_naics") %>%
  filter(is.na(summary)) %>%
  select(code = sector, concordance = title)
x <- full_join(tab_com, tab_ind, "code") %>%
  full_join(conc, "code")

# before combination: some mis-alignment between commodity and concordance codes
x %>% filter(is.na(commodity) | is.na(industry) | is.na(concordance))

conc1 <- left_join(
  conc, 
  bea_io$ind_code_comb %>%
    filter(ilevel == "sector") %>%
    select(code, new_code, new_title),
  by = "code") %>%
  mutate(
    code = if_else(is.na(new_code), code, new_code),
    concordance = if_else(is.na(new_title), concordance, new_title)
  ) %>%
  distinct(code, concordance)
x1 <- full_join(tab_com, tab_ind, "code") %>%
  full_join(conc1, "code")

# # after combination: everything matches
x1 %>% filter(is.na(commodity) | is.na(industry) | is.na(concordance))

```


## summary

Nothing needs to be combined at summary level.
Only discrepancy is in "Used" and "Other" commodities that are not in industry.


```{r}
tab_com <- pubdata::bea_io_get("2022_sup_sum_2012") %>%
  filter(core_matrix) %>%
  select(code = row_code, commodity = row_name) %>%
  distinct()
tab_ind <- pubdata::bea_io_get("2022_sup_sum_2012") %>%
  filter(core_matrix) %>%
  select(code = col_code, industry = col_name) %>%
  distinct()
conc <- pubdata::bea_io_get("2022_naics") %>%
  filter(!is.na(summary), is.na(u_summary)) %>%
  select(code = summary, concordance = title)
x <- full_join(tab_com, tab_ind, "code") %>%
  full_join(conc, "code")

x %>% filter(is.na(commodity) | is.na(industry) | is.na(concordance))

```


## detail 

There is one-to-one match between each concordance code and commodity or industry, nothing needs to be combined.
Commodity only: Scrap; Used and secondhand goods; Noncomparable imports; Rest of the world adjustment.
Industry only: Secondary smelting and alloying of aluminum; Federal electric utilities; State and local government passenger transit; State and local government electric utilities.


```{r}
tab_com <- pubdata::bea_io_get("2022_sup_det_2012") %>%
  filter(core_matrix) %>%
  select(code = row_code, commodity = row_name) %>%
  distinct()
tab_ind <- pubdata::bea_io_get("2022_sup_det_2012") %>%
  filter(core_matrix) %>%
  select(code = col_code, industry = col_name) %>%
  distinct()
conc <- pubdata::bea_io_get("2022_naics") %>%
  filter(!is.na(detail)) %>%
  select(code = detail, concordance = title)
x <- full_join(tab_com, tab_ind, "code") %>%
  full_join(conc, "code")

x %>% filter(is.na(commodity) | is.na(industry) | is.na(concordance))

```


## NAICS duplicates

Some BEA industries refer to the same NAICS code.
Construction is the primary case, where all codes at every level map to NAICS "23".
Another case is summary-level Housing ("HS") and Other real estate ("ORE") - both map into NAICS "531".
Notably, detail "531HSO" and "531HST" are both in summary "HS", but "531HSO" has no NAICS code
We combine all into "531".

We combine these codes both in concordances and in I-O tables.

```{r}
# show NAICS duplicates
pubdata::bea_io_get("2022_naics") %>%
  filter(!is.na(naics), n() > 1, .by = naics)

```

```{r}
bea_io$concordance()
```



# Import

```{r}
pubdata::bea_io_get("2022_imp_det_2012") %>% head()

```



```{r}
import_table <- pubdata::bea_io_get("2022_imp_det_2012")
use_table <- pubdata::bea_io_get("2022_use_det_2012")

df <- full_join(
  import_table %>%
    filter(core_matrix) %>%
    select(comcode = row_code, comname = row_name, indcode = col_code, indname = col_name, import = value) %>%
    mutate(import_sh = import / sum(import, na.rm = TRUE), .by = "comcode"),
  use_table %>%
    filter(core_matrix) %>%
    select(comcode = row_code, indcode = col_code, use = value) %>%
    mutate(use_sh = use / sum(use, na.rm = TRUE), .by = "comcode"),
  by = c("comcode", "indcode")
)

df %>% head()
```




```{r}
x <- df %>%
  summarize(import_tot = sum(import, na.rm = TRUE), .by = "comcode") %>%
  mutate(com_imp_rank = row_number(desc(import_tot)))
dfgg <- df %>%
  left_join(x, by = "comcode") %>%
  arrange(com_imp_rank, desc(use)) %>%
  slice_head(n = 10, by = "comcode")

theme_set(theme_minimal())
ps <- list()
for (r in 1:30) {
  xtitle <- dfgg %>%
    filter(com_imp_rank == r) %>%
    distinct(comcode, comname) %>%
    mutate(label = glue("{comcode} ({comname})")) %>%
    pull(label)
    
  ps[[r]] <- dfgg %>%
    filter(com_imp_rank == r) %>%
    select(indcode, use_sh, import_sh) %>%
    pivot_longer(!indcode) %>%
    mutate(indcode = ordered(indcode, unique(indcode)),
           name = ordered(name, c("use_sh", "import_sh"))) %>%
    ggplot() +
    geom_col(aes(indcode, value, fill = name), position = "dodge") +
    theme(legend.position = "inside", legend.position.inside = c(0.9, 0.9)) +
    labs(title = xtitle, x = NULL, y = NULL)
}

p <- patchwork::wrap_plots(ps[1:10], ncol = 2)
ps[[1]]
```

```{r}
for (p in ps) {
  print(p)
}
```


## Intermediate vs consumer use.

334300	Audio and video equipment manufacturing

```{r}
import_table <- pubdata::bea_io_get("2022_imp_det_2012")
use_table <- pubdata::bea_io_get("2022_use_det_2012")

df <- full_join(
  use_table %>% 
    filter(col_name %in% c("Total Intermediate", "Total use of products")) %>%
    pivot_wider(id_cols = c("row_code", "row_name"), names_from = col_name) %>%
    rename(comcode = row_code, comname = row_name, use_int = `Total Intermediate`, use_tot = `Total use of products`) %>%
    mutate(use_int_sh = use_int / abs(use_tot)),
  import_table %>%
    filter(col_name %in% c("Imports of goods and services", "Total Intermediate")) %>%
    pivot_wider(id_cols = c("row_code", "row_name"), names_from = col_name) %>%
    rename(comcode = row_code, comname = row_name, imp_int = `Total Intermediate`, imp_tot = `Imports of goods and services`) %>%
    mutate(imp_tot = abs(imp_tot), imp_int_sh = imp_int / imp_tot),
  by = c("comcode", "comname")
)

# fishing for a good example
df %>%
  mutate(across(c(imp_int_sh, use_int_sh), \(x) round(x, 4))) %>%
  filter((imp_int_sh < 0.3 | imp_int_sh > 0.7), between(use_int_sh, 0.4, 0.6)) %>%
  View()

df %>%
  mutate(across(c(imp_int_sh, use_int_sh), \(x) round(x, 4))) %>%
  filter(comcode %in% c("337215", "311700"))

```


# Tests

Verify that detail matrix aggregated to summary level via concordance is the same as the original summary matrix.

```{r}

local({
  conc <- bea_io$concordance() %>%
    distinct(summary, detail) %>%
    drop_na()
  for (tab_fn_name in c("supply_table", "use_table")) {
    tab_fn <- get(tab_fn_name, envir = bea_io)
    tab_sum <- tab_fn(2012, "sum") %>% filter(core_matrix)
    tab_det <- tab_fn(2012, "det") %>% filter(core_matrix)
    
    tab_det_agg <- tab_det %>%
      left_join(conc, join_by(row_code == detail)) %>%
      mutate(row_code = summary) %>%
      select(!summary) %>%
      left_join(conc, join_by(col_code == detail)) %>%
      mutate(col_code = summary) %>%
      summarize(value = sum(value), .by = c("row_code", "col_code"))
      
    tab_compare <- tab_sum %>%
      select(row_code, col_code, value) %>%
      full_join(tab_det_agg, by = c("row_code", "col_code"), suffix = c("_sum", "_det")) %>%
      mutate(adif = abs(value_sum - value_det))
    
    # within rounding tolerance
    stopifnot(all(tab_compare$adif < 14))
  }
})

```




