---
title: "Variables codebook"
format: 
  html:
    toc: true
    code-fold: true
    df-print: paged
    embed-resources: true
---

```{r}
#| output: false

library(tidyverse)

root_dir <- rprojroot::find_rstudio_root_file()
```


# Connectedness and growth

Created in `nbs/connectedness_and_growth.qmd`.


```{r}
# load dataframe
df <- read.csv(file.path(root_dir, "data/conn_growth_analysis.csv"))
```

**year**  
Data year.

```{r}
df %>% count(year)
```



**est**  
Number of establishments in a county, combined QCEW and hiring farms from AgCensus.

```{r}
df$est %>% summary()
```


**emp**  
County employment from QCEW, including hired labor from AgCensus, number of workers.

```{r}
df$emp %>% summary()
```


**output**  
Total county output, in thousands of 2020 dollars.

```{r}
df$output %>% summary()
```


**emp_gr, est_gr, output_gr**  
Annualized growth rate of county employment, number of establishments and output between years `t` and `t-5`.

```{r}
df$est_gr %>% summary()
df$emp_gr %>% summary()
df$output_gr %>% summary()
```


**agmine_output**  
Total county output in agriculture and mining, in thousands of 2020 dollars.

```{r}
df$agmine_output %>% summary()
```


**intermed_supply**  
Total county supply of intermediate commodities, in thousands of 2020 dollars.

```{r}
df$intermed_supply %>% summary()
```


**intermed_demand**  
Total county demand for intermediate commodities, in thousands of 2020 dollars.

```{r}
df$intermed_demand %>% summary()
```


**excess_intermed_supply**  
Total county supply of intermediate commodities in excess of local demand, in thousands of 2020 dollars.

```{r}
df$excess_intermed_supply %>% summary()
```


**excess_intermed_demand**  
Total county demand for intermediate commodities in excess of local supply, in thousands of 2020 dollars.

```{r}
df$excess_intermed_demand %>% summary()
```


**agmine_excess_intermed_supply**  
Total county supply of intermediate commodities in excess of local demand in agriculture and mining, in thousands of 2020 dollars.

```{r}
df$agmine_excess_intermed_supply %>% summary()
```


**agmine_excess_intermed_demand**  
Total county demand for intermediate commodities in excess of local supply in agriculture and mining, in thousands of 2020 dollars.

```{r}
df$agmine_excess_intermed_demand %>% summary()
```


**conn_maxabs_output**  
County connectedness metric: share of output absorbable by the most compatible receiving county.
Trade impedance: no trade beyond 300-mile distance from the county.

```{r}
df$conn_maxabs_output %>% summary()
```



