---
title: "Variables codebook"
format: 
  html:
    toc: true
    code-fold: true
    df-print: paged
    embed-resources: true
---

```{r}
library(tidyverse)

root_dir <- rprojroot::find_rstudio_root_file()
```


# Connectedness and growth


```{r}
# load dataframe saved from "connectedness_and_growth.qmd"
df <- read.csv(file.path(root_dir, "data/conn_growth_analysis.csv"))
```

**emp**  
County employment from QCEW, including hired labor from AgCensus.

```{r}
df$emp %>% summary()
```


**emp_gr**  
Annualized growth rate of county employment from t-5.

```{r}
df$emp_gr %>% summary()
```

**conn_maxabs_noimp**  
County connectedness metric, calculated as maximum absorption, normalized by county's total output.
No spatial impedance.

```{r}
df$conn_maxabs_noimp %>% summary()
```


**conn_maxabs_dprox300**  
County connectedness metric, calculated as maximum absorption, normalized by county's total output.
Trade impedance: no trade beyond 300-mile distance from the county.

```{r}
df$conn_maxabs_dprox300 %>% summary()
```

