---
title: "MapsIO"
author: "Austin Sandler"
date: "3/8/2022"
format:
  html:
    self-contained: false
    page-layout: full
    code-fold: true
    code-tools: true
    code_download: yes
    latex_engine: pdflatex
---


```{r preamble, include = FALSE}
# When you click the **Render** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document.

options(scipen=999)

```
::: {.panel-tabset}


### RStartup

<details><summary>Possible Warnings</summary>
```{r packages and libraries}
# Download, Install, and Add R packages as necessary.

# List packages needed for this exercise
packages <- c("dash",
              "dlm",
              "DT",
              "geosphere",
              "ggiraph",
              "knitr",
              "Matrix",
              "reshape2",
              "spdep",
              "scales",
              "tidyverse",
              "tmap",
              "tmaptools")

# Install packages not yet installed
installed_packages <- packages %in% rownames(installed.packages())
if (any(installed_packages == FALSE)) {
  install.packages(packages[!installed_packages], dependencies = TRUE)
}

# Load packages
invisible(lapply(packages, library, character.only = TRUE))

```
</details>



`r paste0("Date: ",  Sys.Date())`

`r kable((.packages()), col.names = "R Packages")`

`r paste0("Project directory: ",  getwd())`

```{r}
## Load previously generated datafiles

load(file = "Data.RData")

## Choose Sector level (1) or Summary level (2) or Detail level  (3)
lvl = 1

```

### WI Test

```{r}
WItest_rural <- grep("^55", rownames(TIGER_RUCC), value=TRUE) %>%
                grep("55009", ., value=TRUE, invert = TRUE) %>% 
                grep("55025", ., value=TRUE, invert = TRUE) %>% 
                grep("55079", ., value=TRUE, invert = TRUE)

WItest_primary <- c("55009", "55025", "55079", "17031", "27053")
names(WItest_primary) <- WItest_primary

WItest_all <- c(WItest_rural, WItest_primary) %>% as.vector()

```


"Simple Similarity Index": $\Vert \mathbf{Ax^{r}} - \mathbf{x^{s}} \Vert_{2} ~ \forall ~ r,s \in l$

```{r}
##  Simple Similarity Index
WISim <- Sim_mat[[lvl]][WItest_rural, WItest_primary]

RowMin <- rbind( cbind(place = rownames(WISim), match = colnames(WISim)[apply(WISim, 1, which.min)]), cbind(WItest_primary, WItest_primary))

table(RowMin[,"match"])

TIGER_RUCC_h1m <- inner_join(TIGER_RUCC, RowMin, by = "place", copy = TRUE)

ggplot( TIGER_RUCC_h1m ) +
  geom_sf_interactive(aes(fill = match), color = NA) +
  coord_sf() +
  theme_void() +
  labs(title = "              First Order Similarity Matching",
       subtitle = "              Simple Similarity Index",
       fill = "Regional Cluster") +
  scale_fill_brewer(palette = "Set1",
                    labels = (TIGER_RUCC_h1m  %>% filter(place %in% unique(TIGER_RUCC_h1m$match) ) %>% pull(County_Name)) )

```

"New Regional-level Total Output Table":  $\mathbf{X}$

```{r}
WI_match <- sparseMatrix(
  i = match(rownames(WISim), WItest_rural), 
  j = match(c(colnames(WISim)[apply(WISim, 1, which.min)]), WItest_primary),  
  x = 1L,
  dims = c(nrow(WISim), ncol(WISim)),
  dimnames = list(WItest_rural, WItest_primary)
) %>% as.matrix()

Xpay_mat_H1 <- (Xpay_mat[[lvl]][, WItest_rural] %*% WI_match) + Xpay_mat[[lvl]][, colnames(WI_match)]
Xpay_mat_H1


```


```{r}
# 
# 
# primary_c <- Xpay_mat[[lvl]][, colnames(WI_match)] %>% as.data.frame()  %>% select(order(colnames(.)))
# primary_c <- cbind(industry = seq_along(primary_c[,1]), primary_c) 
# primary_c <- melt(primary_c, id = "industry")
# 
# x=1
# Xpay_mat[[lvl]][, WItest_rural][, x]
# test <- Xpay_mat[[lvl]][, WItest_rural][, x] %>% as.data.frame()
# colnames(test) <- colnames(Xpay_mat[[lvl]])[x]
# test <- cbind(industry = seq_along(test[,1]), test) 
# 
# 
# ggbase <- ggplot(test) +
#       geom_col(aes(x = industry, 
#                  y = log10(`01001`))) +
#   geom_point(data = primary_c, 
#              aes(x = industry,
#                  y = log10(value),
#                  color = variable)) +   
#   geom_line(data = primary_c, 
#             aes(x = industry,
#                 y = log10(value),
#                 color = variable)) +
#   labs(x = "Industry", y = "log(Payroll)") +
#   scale_x_discrete(limits=factor(1:dim(Xpay_mat[[lvl]])[1]), labels = c(rownames(Xpay_mat[[lvl]]))) +
#   #scale_y_continuous(labels = label_number(suffix = " M", scale = 1e-6)) +
#   scale_y_continuous(trans='log10') +
#    labs(title = "Payroll Distribution by Industry",
#        color = "Counties") +
#   scale_color_discrete(labels = (TIGER_RUCC_h1m  %>% filter(place %in% unique(TIGER_RUCC_h1m$match) ) %>% pull(County_Name)) ) 
# ggbase 
# 
# ggplotly(ggbase)

```




"Relative Similarity Index": $\Vert \mathbf{Ax^{r}}(\mathbf{iAx^{r}})^{-1} - \mathbf{x^{s}}(\mathbf{ix^{s}})^{-1} \Vert_{2} ~ \forall ~ r,s \in l$

```{r}
##  Relative Similarity Index
WISim <- Sim_mat_rel[[lvl]][WItest_rural, WItest_primary]

RowMin <- rbind( cbind(place = rownames(WISim), match = colnames(WISim)[apply(WISim, 1, which.min)]), cbind(WItest_primary, WItest_primary))

table(RowMin[,"match"])

TIGER_RUCC_h1m <- inner_join(TIGER_RUCC, RowMin, by = "place", copy = TRUE)

ggplot( TIGER_RUCC_h1m ) +
  geom_sf(aes(fill = match), color = NA) +
  coord_sf() +
  theme_void() +
  labs(title = "              First Order Similarity Matching",
       subtitle = "              Relative Similarity Index",
       fill = "Regional Cluster") +
  scale_fill_brewer(palette = "Set1",
                    labels = (TIGER_RUCC_h1m  %>% filter(place %in% unique(TIGER_RUCC_h1m$match) ) %>% pull(County_Name)) )
```

"New Regional-level Total Output Table":  $\mathbf{X}$

```{r}
WI_match <- sparseMatrix(
  i = match(rownames(WISim), WItest_rural), 
  j = match(c(colnames(WISim)[apply(WISim, 1, which.min)]), WItest_primary),  
  x = 1L,
  dims = c(nrow(WISim), ncol(WISim)),
  dimnames = list(WItest_rural, WItest_primary)
) %>% as.matrix()

Xpay_mat_H1 <- (Xpay_mat[[lvl]][, WItest_rural] %*% WI_match) + Xpay_mat[[lvl]][, colnames(WI_match)]
Xpay_mat_H1
```

"Import Similarity Index": $\Vert \max\{\mathbf{Ax^{r}} - \mathbf{x^{r}}, 0\} - \mathbf{x^{s}} \Vert_{2} ~ \forall ~ r,s \in l$

```{r}
##  Import Similarity Index
WISim <- Sim_mat_imp[[lvl]][WItest_rural, WItest_primary]

RowMin <- rbind( cbind(place = rownames(WISim), match = colnames(WISim)[apply(WISim, 1, which.min)]), cbind(WItest_primary, WItest_primary))

table(RowMin[,"match"])

TIGER_RUCC_h1m <- inner_join(TIGER_RUCC, RowMin, by = "place", copy = TRUE)

ggplot( TIGER_RUCC_h1m ) +
  geom_sf(aes(fill = match), color = NA) +
  coord_sf() +
  theme_void() +
  labs(title = "              First Order Similarity Matching",
       subtitle = "              Import Similarity Index",
       fill = "Regional Cluster") +
  scale_fill_brewer(palette = "Set1",
                    labels = (TIGER_RUCC_h1m  %>% filter(place %in% unique(TIGER_RUCC_h1m$match) ) %>% pull(County_Name)) )
```

"New Regional-level Total Output Table":  $\mathbf{X}$

```{r}
WI_match <- sparseMatrix(
  i = match(rownames(WISim), WItest_rural), 
  j = match(c(colnames(WISim)[apply(WISim, 1, which.min)]), WItest_primary),  
  x = 1L,
  dims = c(nrow(WISim), ncol(WISim)),
  dimnames = list(WItest_rural, WItest_primary)
) %>% as.matrix()

Xpay_mat_H1 <- (Xpay_mat[[lvl]][, WItest_rural] %*% WI_match) + Xpay_mat[[lvl]][, colnames(WI_match)]
Xpay_mat_H1
```


"*Import Similarity Index - Net Exports": $\Vert \max\{\mathbf{Ax^{r}} - \mathbf{x^{r}}, 0\} - \mathbf{\tilde{x}^{s}} \Vert_{2} ~ \forall ~ r,s \in l$

```{r}
##  *Import Similarity Index - Net Exports
WISim <- Sim_mat_exp[[lvl]][WItest_rural, WItest_primary]

RowMin <- rbind( cbind(place = rownames(WISim), match = colnames(WISim)[apply(WISim, 1, which.min)]), cbind(WItest_primary, WItest_primary))

table(RowMin[,"match"])

TIGER_RUCC_h1m <- inner_join(TIGER_RUCC, RowMin, by = "place", copy = TRUE)

ggplot( TIGER_RUCC_h1m ) +
  geom_sf(aes(fill = match), color = NA) +
  coord_sf() +
  theme_void() +
  labs(title = "              First Order Similarity Matching",
       subtitle = "              *Import Similarity Index - Net Exports",
       fill = "Regional Cluster") +
  scale_fill_brewer(palette = "Set1",
                    labels = (TIGER_RUCC_h1m  %>% filter(place %in% unique(TIGER_RUCC_h1m$match) ) %>% pull(County_Name))  )
```

"New Regional-level Total Output Table":  $\mathbf{X}$

```{r}
WI_match <- sparseMatrix(
  i = match(rownames(WISim), WItest_rural), 
  j = match(c(colnames(WISim)[apply(WISim, 1, which.min)]), WItest_primary),  
  x = 1L,
  dims = c(nrow(WISim), ncol(WISim)),
  dimnames = list(WItest_rural, WItest_primary)
) %>% as.matrix()

Xpay_mat_H1 <- (Xpay_mat[[lvl]][, WItest_rural] %*% WI_match) + Xpay_mat[[lvl]][, colnames(WI_match)]
Xpay_mat_H1
```


"Relative Import Similarity Index": $\Vert \max\{\mathbf{Ax^{r}}(\mathbf{iAx^{r}})^{-1} - \mathbf{x^{r}}(\mathbf{ix^{r}})^{-1}, 0\} - \mathbf{x^{s}}(\mathbf{ix^{s}})^{-1} \Vert_{2} ~ \forall ~ r,s \in l$

```{r}
##  Relative Import Similarity Index
WISim <- Sim_mat_imp_rel[[lvl]][WItest_rural, WItest_primary]

RowMin <- rbind( cbind(place = rownames(WISim), match = colnames(WISim)[apply(WISim, 1, which.min)]), cbind(WItest_primary, WItest_primary))

table(RowMin[,"match"])

TIGER_RUCC_h1m <- inner_join(TIGER_RUCC, RowMin, by = "place", copy = TRUE)

ggplot( TIGER_RUCC_h1m ) +
  geom_sf(aes(fill = match), color = NA) +
  coord_sf() +
  theme_void() +
  labs(title = "              First Order Similarity Matching",
       subtitle = "              Relative Import Similarity Index",
       fill = "Regional Cluster") +
  scale_fill_brewer(palette = "Set1",
                    labels = (TIGER_RUCC_h1m  %>% filter(place %in% unique(TIGER_RUCC_h1m$match) ) %>% pull(County_Name)) )
```

"New Regional-level Total Output Table":  $\mathbf{X}$

```{r}
WI_match <- sparseMatrix(
  i = match(rownames(WISim), WItest_rural), 
  j = match(c(colnames(WISim)[apply(WISim, 1, which.min)]), WItest_primary),  
  x = 1L,
  dims = c(nrow(WISim), ncol(WISim)),
  dimnames = list(WItest_rural, WItest_primary)
) %>% as.matrix()

Xpay_mat_H1 <- (Xpay_mat[[lvl]][, WItest_rural] %*% WI_match) + Xpay_mat[[lvl]][, colnames(WI_match)]
Xpay_mat_H1
```


"*Relative Import Similarity Index - Net Exports": $\Vert \max\{\mathbf{Ax^{r}}(\mathbf{iAx^{r}})^{-1} - \mathbf{x^{r}}(\mathbf{ix^{r}})^{-1}, 0\} - \mathbf{\tilde{x}^{s}}(\mathbf{i\tilde{x}^{s}})^{-1} \Vert_{2} ~ \forall ~ r,s \in l$

```{r}
##  *Relative Import Similarity Index - Net Exports
WISim <- Sim_mat_exp_rel[[lvl]][WItest_rural, WItest_primary]

RowMin <- rbind( cbind(place = rownames(WISim), match = colnames(WISim)[apply(WISim, 1, which.min)]), cbind(WItest_primary, WItest_primary))

table(RowMin[,"match"])

TIGER_RUCC_h1m <- inner_join(TIGER_RUCC, RowMin, by = "place", copy = TRUE)

ggplot( TIGER_RUCC_h1m ) +
  geom_sf(aes(fill = match), color = NA) +
  coord_sf() +
  theme_void() +
  labs(title = "              First Order Similarity Matching",
       subtitle = "              *Relative Import Similarity Index - Net Exports",
       fill = "Regional Cluster") +
  scale_fill_brewer(palette = "Set1",
                    labels = (TIGER_RUCC_h1m  %>% filter(place %in% unique(TIGER_RUCC_h1m$match) ) %>% pull(County_Name)) )
```

"New Regional-level Total Output Table":  $\mathbf{X}$

```{r}
WI_match <- sparseMatrix(
  i = match(rownames(WISim), WItest_rural), 
  j = match(c(colnames(WISim)[apply(WISim, 1, which.min)]), WItest_primary),  
  x = 1L,
  dims = c(nrow(WISim), ncol(WISim)),
  dimnames = list(WItest_rural, WItest_primary)
) %>% as.matrix()

Xpay_mat_H1 <- (Xpay_mat[[lvl]][, WItest_rural] %*% WI_match) + Xpay_mat[[lvl]][, colnames(WI_match)]
Xpay_mat_H1
```



### WI Test - Spatial

"Geographical Impedance": $Q^{rs} = e^{-d^{rs}} ~ \forall ~ r,s \in l$


```{r}

WItest_rural <- grep("^55", rownames(TIGER_RUCC), value=TRUE) %>%
                grep("55009", ., value=TRUE, invert = TRUE) %>% 
                grep("55025", ., value=TRUE, invert = TRUE) %>% 
                grep("55079", ., value=TRUE, invert = TRUE)

WItest_primary <- c("55009", "55025", "55079", "17031", "27053")
names(WItest_primary) <- WItest_primary

WItest_all <- c(WItest_rural, WItest_primary) %>% as.vector()

```


"Spatial Simple Similarity Index": $\Vert \mathbf{Ax^{r}} - \mathbf{x^{s}} \Vert_{2} / Q^{rs} ~ \forall ~ r,s \in l$

```{r}
##  Spatial Simple Similarity Index
WISim <- Sim_mat[[lvl]][WItest_rural, WItest_primary] / Q_mat[[lvl]][WItest_rural, WItest_primary] 

RowMin <- rbind( cbind(place = rownames(WISim), match = colnames(WISim)[apply(WISim, 1, which.min)]), cbind(WItest_primary, WItest_primary))

table(RowMin[,"match"])

TIGER_RUCC_h1m <- inner_join(TIGER_RUCC, RowMin, by = "place", copy = TRUE)

ggplot( TIGER_RUCC_h1m ) +
  geom_sf(aes(fill = match), color = NA) +
  coord_sf() +
  theme_void() +
  labs(title = "              First Order Similarity Matching",
       subtitle = "              Spatial Simple Similarity Index",
       fill = "Regional Cluster") +
  scale_fill_brewer(palette = "Set1",
                    labels = (TIGER_RUCC_h1m  %>% filter(place %in% unique(TIGER_RUCC_h1m$match) ) %>% pull(County_Name)) )
```

"New Regional-level Total Output Table":  $\mathbf{X}$

```{r}
WI_match <- sparseMatrix(
  i = match(rownames(WISim), WItest_rural), 
  j = match(c(colnames(WISim)[apply(WISim, 1, which.min)]), WItest_primary),  
  x = 1L,
  dims = c(nrow(WISim), ncol(WISim)),
  dimnames = list(WItest_rural, WItest_primary)
) %>% as.matrix()

Xpay_mat_H1 <- (Xpay_mat[[lvl]][, WItest_rural] %*% WI_match) + Xpay_mat[[lvl]][, colnames(WI_match)]
Xpay_mat_H1
```

"Spatial Relative Similarity Index": $\Vert \mathbf{Ax^{r}}(\mathbf{iAx^{r}})^{-1} - \mathbf{x^{s}}(\mathbf{ix^{s}})^{-1} \Vert_{2} / Q^{rs} ~ \forall ~ r,s \in l$

```{r}
##  Spatial Relative Similarity Index
WISim <- Sim_mat_rel[[lvl]][WItest_rural, WItest_primary] / Q_mat[[lvl]][WItest_rural, WItest_primary] 

RowMin <- rbind( cbind(place = rownames(WISim), match = colnames(WISim)[apply(WISim, 1, which.min)]), cbind(WItest_primary, WItest_primary))

table(RowMin[,"match"])

TIGER_RUCC_h1m <- inner_join(TIGER_RUCC, RowMin, by = "place", copy = TRUE)

ggplot( TIGER_RUCC_h1m ) +
  geom_sf(aes(fill = match), color = NA) +
  coord_sf() +
  theme_void() +
  labs(title = "              First Order Similarity Matching",
       subtitle = "              Spatial Relative Similarity Index",
       fill = "Regional Cluster") +
  scale_fill_brewer(palette = "Set1",
                    labels = (TIGER_RUCC_h1m  %>% filter(place %in% unique(TIGER_RUCC_h1m$match) ) %>% pull(County_Name)) )
```

"New Regional-level Total Output Table":  $\mathbf{X}$

```{r}
WI_match <- sparseMatrix(
  i = match(rownames(WISim), WItest_rural), 
  j = match(c(colnames(WISim)[apply(WISim, 1, which.min)]), WItest_primary),  
  x = 1L,
  dims = c(nrow(WISim), ncol(WISim)),
  dimnames = list(WItest_rural, WItest_primary)
) %>% as.matrix()

Xpay_mat_H1 <- (Xpay_mat[[lvl]][, WItest_rural] %*% WI_match) + Xpay_mat[[lvl]][, colnames(WI_match)]
Xpay_mat_H1
```

"Spatial Import Similarity Index": $\Vert \max\{\mathbf{Ax^{r}} - \mathbf{x^{r}}, 0\} - \mathbf{x^{s}} \Vert_{2} / Q^{rs} ~ \forall ~ r,s \in l$

```{r}
##  Spatial Import Similarity Index
WISim <- Sim_mat_imp[[lvl]][WItest_rural, WItest_primary]  / Q_mat[[lvl]][WItest_rural, WItest_primary] 

RowMin <- rbind( cbind(place = rownames(WISim), match = colnames(WISim)[apply(WISim, 1, which.min)]), cbind(WItest_primary, WItest_primary))

table(RowMin[,"match"])

TIGER_RUCC_h1m <- inner_join(TIGER_RUCC, RowMin, by = "place", copy = TRUE)

ggplot( TIGER_RUCC_h1m ) +
  geom_sf(aes(fill = match), color = NA) +
  coord_sf() +
  theme_void() +
  labs(title = "              First Order Similarity Matching",
       subtitle = "              Spatial Import Similarity Index",
       fill = "Regional Cluster") +
  scale_fill_brewer(palette = "Set1",
                    labels = (TIGER_RUCC_h1m  %>% filter(place %in% unique(TIGER_RUCC_h1m$match) ) %>% pull(County_Name)) )
```

"New Regional-level Total Output Table":  $\mathbf{X}$

```{r}
WI_match <- sparseMatrix(
  i = match(rownames(WISim), WItest_rural), 
  j = match(c(colnames(WISim)[apply(WISim, 1, which.min)]), WItest_primary),  
  x = 1L,
  dims = c(nrow(WISim), ncol(WISim)),
  dimnames = list(WItest_rural, WItest_primary)
) %>% as.matrix()

Xpay_mat_H1 <- (Xpay_mat[[lvl]][, WItest_rural] %*% WI_match) + Xpay_mat[[lvl]][, colnames(WI_match)]
Xpay_mat_H1
```

"Spatial *Import Similarity Index - Net Exports": $\Vert \max\{\mathbf{Ax^{r}} - \mathbf{x^{r}}, 0\} - \mathbf{\tilde{x}^{s}} \Vert_{2}  / Q^{rs} ~ \forall ~ r,s \in l$

```{r}
##  Spatial *Import Similarity Index - Net Exports
WISim <- Sim_mat_exp[[lvl]][WItest_rural, WItest_primary] / Q_mat[[lvl]][WItest_rural, WItest_primary] 

RowMin <- rbind( cbind(place = rownames(WISim), match = colnames(WISim)[apply(WISim, 1, which.min)]), cbind(WItest_primary, WItest_primary))

table(RowMin[,"match"])

TIGER_RUCC_h1m <- inner_join(TIGER_RUCC, RowMin, by = "place", copy = TRUE)

ggplot( TIGER_RUCC_h1m ) +
  geom_sf(aes(fill = match), color = NA) +
  coord_sf() +
  theme_void() +
  labs(title = "              First Order Similarity Matching",
       subtitle = "              Spatial *Import Similarity Index - Net Exports",
       fill = "Regional Cluster") +
  scale_fill_brewer(palette = "Set1",
                    labels = (TIGER_RUCC_h1m  %>% filter(place %in% unique(TIGER_RUCC_h1m$match) ) %>% pull(County_Name))  )
```

"New Regional-level Total Output Table":  $\mathbf{X}$

```{r}
WI_match <- sparseMatrix(
  i = match(rownames(WISim), WItest_rural), 
  j = match(c(colnames(WISim)[apply(WISim, 1, which.min)]), WItest_primary),  
  x = 1L,
  dims = c(nrow(WISim), ncol(WISim)),
  dimnames = list(WItest_rural, WItest_primary)
) %>% as.matrix()

Xpay_mat_H1 <- (Xpay_mat[[lvl]][, WItest_rural] %*% WI_match) + Xpay_mat[[lvl]][, colnames(WI_match)]
Xpay_mat_H1
```

"Spatial Relative Import Similarity Index": $\Vert \max\{\mathbf{Ax^{r}}(\mathbf{iAx^{r}})^{-1} - \mathbf{x^{r}}(\mathbf{ix^{r}})^{-1}, 0\} - \mathbf{x^{s}}(\mathbf{ix^{s}})^{-1} \Vert_{2} / Q^{rs} ~ \forall ~ r,s \in l$

```{r}
##  Spatial Relative Import Similarity Index
WISim <- Sim_mat_imp_rel[[lvl]][WItest_rural, WItest_primary] / Q_mat[[lvl]][WItest_rural, WItest_primary] 

RowMin <- rbind( cbind(place = rownames(WISim), match = colnames(WISim)[apply(WISim, 1, which.min)]), cbind(WItest_primary, WItest_primary))

table(RowMin[,"match"])

TIGER_RUCC_h1m <- inner_join(TIGER_RUCC, RowMin, by = "place", copy = TRUE)

ggplot( TIGER_RUCC_h1m ) +
  geom_sf(aes(fill = match), color = NA) +
  coord_sf() +
  theme_void() +
  labs(title = "              First Order Similarity Matching",
       subtitle = "              Spatial Relative Import Similarity Index",
       fill = "Regional Cluster") +
  scale_fill_brewer(palette = "Set1",
                    labels = (TIGER_RUCC_h1m  %>% filter(place %in% unique(TIGER_RUCC_h1m$match) ) %>% pull(County_Name)) )
```

"New Regional-level Total Output Table":  $\mathbf{X}$

```{r}
WI_match <- sparseMatrix(
  i = match(rownames(WISim), WItest_rural), 
  j = match(c(colnames(WISim)[apply(WISim, 1, which.min)]), WItest_primary),  
  x = 1L,
  dims = c(nrow(WISim), ncol(WISim)),
  dimnames = list(WItest_rural, WItest_primary)
) %>% as.matrix()

Xpay_mat_H1 <- (Xpay_mat[[lvl]][, WItest_rural] %*% WI_match) + Xpay_mat[[lvl]][, colnames(WI_match)]
Xpay_mat_H1
```

"Spatial *Relative Import Similarity Index - Net Exports": $\Vert \max\{\mathbf{Ax^{r}}(\mathbf{iAx^{r}})^{-1} - \mathbf{x^{r}}(\mathbf{ix^{r}})^{-1}, 0\} - \mathbf{\tilde{x}^{s}}(\mathbf{i\tilde{x}^{s}})^{-1} \Vert_{2} / Q^{rs} ~ \forall ~ r,s \in l$

```{r}
##  Spatial  *Relative Import Similarity Index - Net Exports

WISim <- Sim_mat_exp_rel[[lvl]][WItest_rural, WItest_primary] / Q_mat[[lvl]][WItest_rural, WItest_primary] 

RowMin <- rbind( cbind(place = rownames(WISim), match = colnames(WISim)[apply(WISim, 1, which.min)]), cbind(WItest_primary, WItest_primary))

table(RowMin[,"match"])

TIGER_RUCC_h1m <- inner_join(TIGER_RUCC, RowMin, by = "place", copy = TRUE)

ggplot( TIGER_RUCC_h1m ) +
  geom_sf(aes(fill = match), color = NA) +
  coord_sf() +
  theme_void() +
  labs(title = "              First Order Similarity Matching",
       subtitle = "              Spatial *Relative Import Similarity Index - Net Exports",
       fill = "Regional Cluster") +
  scale_fill_brewer(palette = "Set1",
                    labels = (TIGER_RUCC_h1m  %>% filter(place %in% unique(TIGER_RUCC_h1m$match) ) %>% pull(County_Name)) )
```

"New Regional-level Total Output Table":  $\mathbf{X}$

```{r}

WI_match <- sparseMatrix(
  i = match(rownames(WISim), WItest_rural), 
  j = match(c(colnames(WISim)[apply(WISim, 1, which.min)]), WItest_primary),  
  x = 1L,
  dims = c(nrow(WISim), ncol(WISim)),
  dimnames = list(WItest_rural, WItest_primary)
) %>% as.matrix()

Xpay_mat_H1 <- (Xpay_mat[[lvl]][, WItest_rural] %*% WI_match) + Xpay_mat[[lvl]][, colnames(WI_match)]
Xpay_mat_H1
```





### WI Test - Natural

```{r}
WItest_rural <- grep("^55", rownames(TIGER_RUCC), value=TRUE) %>%
                grep("55009", ., value=TRUE, invert = TRUE) %>% 
                grep("55025", ., value=TRUE, invert = TRUE) %>% 
                grep("55079", ., value=TRUE, invert = TRUE)

WItest_primary <- c(grep("^55", rownames(TIGER_RUCC), value=TRUE) , "17031", "27053")
names(WItest_primary) <- WItest_primary

WItest_all <- c(WItest_primary) %>% as.vector()

```


"Simple Similarity Index": $\Vert \mathbf{Ax^{r}} - \mathbf{x^{s}} \Vert_{2} ~ \forall ~ r,s \in l$

```{r}
##  Simple Similarity Index
WISim <- Sim_mat[[lvl]][WItest_rural, WItest_primary]

RowMin <- rbind( cbind(place = rownames(WISim), match = colnames(WISim)[apply(WISim, 1, which.min)]), 
                 cbind( c("55009", "55025", "55079", "17031", "27053"),  c("55009", "55025", "55079", "17031", "27053") ) )

table(RowMin[,"match"])

TIGER_RUCC_h1m <- inner_join(TIGER_RUCC, RowMin, by = "place", copy = TRUE)

ggplot( TIGER_RUCC_h1m ) +
  geom_sf(aes(fill = match), color = NA) +
  coord_sf() +
  theme_void() +
  labs(title = "              First Order Similarity Matching",
       subtitle = "              Simple Similarity Index",
       fill = "Regional Cluster")  +
       scale_fill_discrete(labels = (TIGER_RUCC_h1m  %>% filter(place %in% unique(TIGER_RUCC_h1m$match) ) %>% pull(County_Name)) )

```


"Relative Similarity Index": $\Vert \mathbf{Ax^{r}}(\mathbf{iAx^{r}})^{-1} - \mathbf{x^{s}}(\mathbf{ix^{s}})^{-1} \Vert_{2} ~ \forall ~ r,s \in l$

```{r}
##  Relative Similarity Index
WISim <- Sim_mat_rel[[lvl]][WItest_rural, WItest_primary]


RowMin <- rbind( cbind(place = rownames(WISim), match = colnames(WISim)[apply(WISim, 1, which.min)]), 
                 cbind( c("55009", "55025", "55079", "17031", "27053"),  c("55009", "55025", "55079", "17031", "27053") ) )

table(RowMin[,"match"])

TIGER_RUCC_h1m <- inner_join(TIGER_RUCC, RowMin, by = "place", copy = TRUE)

ggplot( TIGER_RUCC_h1m ) +
  geom_sf(aes(fill = match), color = NA) +
  coord_sf() +
  theme_void() +
  labs(title = "              First Order Similarity Matching",
       subtitle = "              Relative Similarity Index",
       fill = "Regional Cluster") +
       scale_fill_discrete(labels = (TIGER_RUCC_h1m  %>% filter(place %in% unique(TIGER_RUCC_h1m$match) ) %>% pull(County_Name)) )
```



"Import Similarity Index": $\Vert \max\{\mathbf{Ax^{r}} - \mathbf{x^{r}}, 0\} - \mathbf{x^{s}} \Vert_{2} ~ \forall ~ r,s \in l$

```{r}
##  Import Similarity Index
WISim <- Sim_mat_imp[[lvl]][WItest_rural, WItest_primary]

RowMin <- rbind( cbind(place = rownames(WISim), match = colnames(WISim)[apply(WISim, 1, which.min)]), 
                 cbind( c("55009", "55025", "55079", "17031", "27053"),  c("55009", "55025", "55079", "17031", "27053") ) )

table(RowMin[,"match"])

TIGER_RUCC_h1m <- inner_join(TIGER_RUCC, RowMin, by = "place", copy = TRUE)

ggplot( TIGER_RUCC_h1m ) +
  geom_sf(aes(fill = match), color = NA) +
  coord_sf() +
  theme_void() +
  labs(title = "              First Order Similarity Matching",
       subtitle = "              Import Similarity Index",
       fill = "Regional Cluster") +
       scale_fill_discrete(labels = (TIGER_RUCC_h1m  %>% filter(place %in% unique(TIGER_RUCC_h1m$match) ) %>% pull(County_Name)) )
```

"*Import Similarity Index - Net Exports": $\Vert \max\{\mathbf{Ax^{r}} - \mathbf{x^{r}}, 0\} - \mathbf{\tilde{x}^{s}} \Vert_{2} ~ \forall ~ r,s \in l$

```{r}
##  *Import Similarity Index - Net Exports
WISim <- Sim_mat_exp[[lvl]][WItest_rural, WItest_primary]

RowMin <- rbind( cbind(place = rownames(WISim), match = colnames(WISim)[apply(WISim, 1, which.min)]), 
                 cbind( c("55009", "55025", "55079", "17031", "27053"),  c("55009", "55025", "55079", "17031", "27053") ) )

table(RowMin[,"match"])

TIGER_RUCC_h1m <- inner_join(TIGER_RUCC, RowMin, by = "place", copy = TRUE)

ggplot( TIGER_RUCC_h1m ) +
  geom_sf(aes(fill = match), color = NA) +
  coord_sf() +
  theme_void() +
  labs(title = "              First Order Similarity Matching",
       subtitle = "              *Import Similarity Index - Net Exports",
       fill = "Regional Cluster") +
       scale_fill_discrete(labels = (TIGER_RUCC_h1m  %>% filter(place %in% unique(TIGER_RUCC_h1m$match) ) %>% pull(County_Name)) )
```

"Relative Import Similarity Index": $\Vert \max\{\mathbf{Ax^{r}}(\mathbf{iAx^{r}})^{-1} - \mathbf{x^{r}}(\mathbf{ix^{r}})^{-1}, 0\} - \mathbf{x^{s}}(\mathbf{ix^{s}})^{-1} \Vert_{2} ~ \forall ~ r,s \in l$

```{r}
##  Relative Import Similarity Index
WISim <- Sim_mat_imp_rel[[lvl]][WItest_rural, WItest_primary]

RowMin <- rbind( cbind(place = rownames(WISim), match = colnames(WISim)[apply(WISim, 1, which.min)]), 
                 cbind( c("55009", "55025", "55079", "17031", "27053"),  c("55009", "55025", "55079", "17031", "27053") ) )

table(RowMin[,"match"])

TIGER_RUCC_h1m <- inner_join(TIGER_RUCC, RowMin, by = "place", copy = TRUE)

ggplot( TIGER_RUCC_h1m ) +
  geom_sf(aes(fill = match), color = NA) +
  coord_sf() +
  theme_void() +
  labs(title = "              First Order Similarity Matching",
       subtitle = "              Relative Import Similarity Index",
       fill = "Regional Cluster") +
       scale_fill_discrete(labels = (TIGER_RUCC_h1m  %>% filter(place %in% unique(TIGER_RUCC_h1m$match) ) %>% pull(County_Name)) )
```

"*Relative Import Similarity Index - Net Exports": $\Vert \max\{\mathbf{Ax^{r}}(\mathbf{iAx^{r}})^{-1} - \mathbf{x^{r}}(\mathbf{ix^{r}})^{-1}, 0\} - \mathbf{\tilde{x}^{s}}(\mathbf{i\tilde{x}^{s}})^{-1} \Vert_{2} ~ \forall ~ r,s \in l$

```{r}
##  *Relative Import Similarity Index - Net Exports
WISim <- Sim_mat_exp_rel[[lvl]][WItest_rural, WItest_primary]

RowMin <- rbind( cbind(place = rownames(WISim), match = colnames(WISim)[apply(WISim, 1, which.min)]), 
                 cbind( c("55009", "55025", "55079", "17031", "27053"),  c("55009", "55025", "55079", "17031", "27053") ) )

table(RowMin[,"match"])

TIGER_RUCC_h1m <- inner_join(TIGER_RUCC, RowMin, by = "place", copy = TRUE)

ggplot( TIGER_RUCC_h1m ) +
  geom_sf(aes(fill = match), color = NA) +
  coord_sf() +
  theme_void() +
  labs(title = "              First Order Similarity Matching",
       subtitle = "              *Relative Import Similarity Index - Net Exports",
       fill = "Regional Cluster") +
       scale_fill_discrete(labels = (TIGER_RUCC_h1m  %>% filter(place %in% unique(TIGER_RUCC_h1m$match) ) %>% pull(County_Name)) )
```











### WI Test - Natural and Spatial 

```{r}
WItest_rural <- grep("^55", rownames(TIGER_RUCC), value=TRUE) %>%
                grep("55009", ., value=TRUE, invert = TRUE) %>% 
                grep("55025", ., value=TRUE, invert = TRUE) %>% 
                grep("55079", ., value=TRUE, invert = TRUE)

WItest_primary <- c(grep("^55", rownames(TIGER_RUCC), value=TRUE) , "17031", "27053")
names(WItest_primary) <- WItest_primary

WItest_all <- c(WItest_primary) %>% as.vector()

```


"Spatial Simple Similarity Index": $\Vert \mathbf{Ax^{r}} - \mathbf{x^{s}} \Vert_{2} ~ \forall ~ r,s \in l$

```{r}
## Spatial Simple Similarity Index
WISim <- Sim_mat[[lvl]][WItest_rural, WItest_primary] / Q_mat[[lvl]][WItest_rural, WItest_primary] 

RowMin <- rbind( cbind(place = rownames(WISim), match = colnames(WISim)[apply(WISim, 1, which.min)]), 
                 cbind( c("55009", "55025", "55079", "17031", "27053"),  c("55009", "55025", "55079", "17031", "27053") ) )

table(RowMin[,"match"])

TIGER_RUCC_h1m <- inner_join(TIGER_RUCC, RowMin, by = "place", copy = TRUE)

ggplot( TIGER_RUCC_h1m ) +
  geom_sf(aes(fill = match), color = NA) +
  coord_sf() +
  theme_void() +
  labs(title = "              First Order Similarity Matching",
       subtitle = "             Spatial Simple Similarity Index",
       fill = "Regional Cluster")  +
       scale_fill_discrete(labels = (TIGER_RUCC_h1m  %>% filter(place %in% unique(TIGER_RUCC_h1m$match) ) %>% pull(County_Name)) )

```


"Spatial Relative Similarity Index": $\Vert \mathbf{Ax^{r}}(\mathbf{iAx^{r}})^{-1} - \mathbf{x^{s}}(\mathbf{ix^{s}})^{-1} \Vert_{2} ~ \forall ~ r,s \in l$

```{r}
## Spatial Relative Similarity Index
WISim <- Sim_mat_rel[[lvl]][WItest_rural, WItest_primary] / Q_mat[[lvl]][WItest_rural, WItest_primary] 


RowMin <- rbind( cbind(place = rownames(WISim), match = colnames(WISim)[apply(WISim, 1, which.min)]), 
                 cbind( c("55009", "55025", "55079", "17031", "27053"),  c("55009", "55025", "55079", "17031", "27053") ) )

table(RowMin[,"match"])

TIGER_RUCC_h1m <- inner_join(TIGER_RUCC, RowMin, by = "place", copy = TRUE)

ggplot( TIGER_RUCC_h1m ) +
  geom_sf(aes(fill = match), color = NA) +
  coord_sf() +
  theme_void() +
  labs(title = "              First Order Similarity Matching",
       subtitle = "             Spatial Relative Similarity Index",
       fill = "Regional Cluster") +
       scale_fill_discrete(labels = (TIGER_RUCC_h1m  %>% filter(place %in% unique(TIGER_RUCC_h1m$match) ) %>% pull(County_Name)) )
```



"Spatial Import Similarity Index": $\Vert \max\{\mathbf{Ax^{r}} - \mathbf{x^{r}}, 0\} - \mathbf{x^{s}} \Vert_{2} ~ \forall ~ r,s \in l$

```{r}
## Spatial Import Similarity Index
WISim <- Sim_mat_imp[[lvl]][WItest_rural, WItest_primary] / Q_mat[[lvl]][WItest_rural, WItest_primary] 

RowMin <- rbind( cbind(place = rownames(WISim), match = colnames(WISim)[apply(WISim, 1, which.min)]), 
                 cbind( c("55009", "55025", "55079", "17031", "27053"),  c("55009", "55025", "55079", "17031", "27053") ) )

table(RowMin[,"match"])

TIGER_RUCC_h1m <- inner_join(TIGER_RUCC, RowMin, by = "place", copy = TRUE)

ggplot( TIGER_RUCC_h1m ) +
  geom_sf(aes(fill = match), color = NA) +
  coord_sf() +
  theme_void() +
  labs(title = "              First Order Similarity Matching",
       subtitle = "             Spatial Import Similarity Index",
       fill = "Regional Cluster") +
       scale_fill_discrete(labels = (TIGER_RUCC_h1m  %>% filter(place %in% unique(TIGER_RUCC_h1m$match) ) %>% pull(County_Name)) )
```

"Spatial *Import Similarity Index - Net Exports": $\Vert \max\{\mathbf{Ax^{r}} - \mathbf{x^{r}}, 0\} - \mathbf{\tilde{x}^{s}} \Vert_{2} ~ \forall ~ r,s \in l$

```{r}
## Spatial *Import Similarity Index - Net Exports
WISim <- Sim_mat_exp[[lvl]][WItest_rural, WItest_primary] / Q_mat[[lvl]][WItest_rural, WItest_primary] 

RowMin <- rbind( cbind(place = rownames(WISim), match = colnames(WISim)[apply(WISim, 1, which.min)]), 
                 cbind( c("55009", "55025", "55079", "17031", "27053"),  c("55009", "55025", "55079", "17031", "27053") ) )

table(RowMin[,"match"])

TIGER_RUCC_h1m <- inner_join(TIGER_RUCC, RowMin, by = "place", copy = TRUE)

ggplot( TIGER_RUCC_h1m ) +
  geom_sf(aes(fill = match), color = NA) +
  coord_sf() +
  theme_void() +
  labs(title = "              First Order Similarity Matching",
       subtitle = "             Spatial *Import Similarity Index - Net Exports",
       fill = "Regional Cluster") +
       scale_fill_discrete(labels = (TIGER_RUCC_h1m  %>% filter(place %in% unique(TIGER_RUCC_h1m$match) ) %>% pull(County_Name)) )
```

"Spatial Relative Import Similarity Index": $\Vert \max\{\mathbf{Ax^{r}}(\mathbf{iAx^{r}})^{-1} - \mathbf{x^{r}}(\mathbf{ix^{r}})^{-1}, 0\} - \mathbf{x^{s}}(\mathbf{ix^{s}})^{-1} \Vert_{2} ~ \forall ~ r,s \in l$

```{r}
## Spatial Relative Import Similarity Index
WISim <- Sim_mat_imp_rel[[lvl]][WItest_rural, WItest_primary] / Q_mat[[lvl]][WItest_rural, WItest_primary] 

RowMin <- rbind( cbind(place = rownames(WISim), match = colnames(WISim)[apply(WISim, 1, which.min)]), 
                 cbind( c("55009", "55025", "55079", "17031", "27053"),  c("55009", "55025", "55079", "17031", "27053") ) )

table(RowMin[,"match"])

TIGER_RUCC_h1m <- inner_join(TIGER_RUCC, RowMin, by = "place", copy = TRUE)

ggplot( TIGER_RUCC_h1m ) +
  geom_sf(aes(fill = match), color = NA) +
  coord_sf() +
  theme_void() +
  labs(title = "              First Order Similarity Matching",
       subtitle = "             Spatial Relative Import Similarity Index",
       fill = "Regional Cluster") +
       scale_fill_discrete(labels = (TIGER_RUCC_h1m  %>% filter(place %in% unique(TIGER_RUCC_h1m$match) ) %>% pull(County_Name)) )
```

"Spatial *Relative Import Similarity Index - Net Exports": $\Vert \max\{\mathbf{Ax^{r}}(\mathbf{iAx^{r}})^{-1} - \mathbf{x^{r}}(\mathbf{ix^{r}})^{-1}, 0\} - \mathbf{\tilde{x}^{s}}(\mathbf{i\tilde{x}^{s}})^{-1} \Vert_{2} ~ \forall ~ r,s \in l$

```{r}
##  Spatial *Relative Import Similarity Index - Net Exports
WISim <- Sim_mat_exp_rel[[lvl]][WItest_rural, WItest_primary] / Q_mat[[lvl]][WItest_rural, WItest_primary] 

RowMin <- rbind( cbind(place = rownames(WISim), match = colnames(WISim)[apply(WISim, 1, which.min)]), 
                 cbind( c("55009", "55025", "55079", "17031", "27053"),  c("55009", "55025", "55079", "17031", "27053") ) )

table(RowMin[,"match"])

TIGER_RUCC_h1m <- inner_join(TIGER_RUCC, RowMin, by = "place", copy = TRUE) %>% mutate(match_name = with(TIGER_RUCC_h1m, County_Name[match(match, place)]), .after = match)

test <- TIGER_RUCC_h1m

ggplot( TIGER_RUCC_h1m ) +
  geom_sf(aes(fill = match), color = NA) +
  coord_sf() +
  theme_void() +
  labs(title = "              First Order Similarity Matching",
       subtitle = "             Spatial *Relative Import Similarity Index - Net Exports",
       fill = "Regional Cluster") +
       scale_fill_discrete(labels = (TIGER_RUCC_h1m  %>% filter(place %in% unique(TIGER_RUCC_h1m$match) ) %>% pull(County_Name)) )
```





```{r}
  tmap_mode("view")
clust_name <- unique(TIGER_RUCC_h1m$match_name)
TIGER_RUCC_h1m %>% 
  tm_shape() +
  tm_fill(col = "match",
          title = "Regional Cluster", 
          id = "NAMELSAD", 
          palette = "Dark2" ) +
  tm_borders(lwd = .5)        
  

```


















### Maps

```{r}
##Match rural to mid

rural <- rownames(Q_mat[[lvl]])[rownames(Q_mat[[lvl]]) %in% filter(TIGER_RUCC, H3==3)$place ]
primary <- rownames(Q_mat[[lvl]])[rownames(Q_mat[[lvl]]) %in% filter(TIGER_RUCC, H3==2)$place ]

Sim <- Sim_mat[[lvl]][rural, primary]

RowMin <- cbind(place = rownames(Sim), match = colnames(Sim)[apply(Sim, 1, which.min)])

table(RowMin[,"match"])

TIGER_RUCC_h1m <- inner_join(TIGER_RUCC, RowMin, by = "place", copy = TRUE)

ggplot(  filter(TIGER_RUCC_h1m, STATEFP=="55")) +
  geom_sf(aes(fill = match), color = NA) +
  coord_sf() +
  theme_void() +
  labs(title = "              First Order Similarity Matching",
       subtitle = "              Simple Similarity Index",
       fill = "Regional Cluster") +
       scale_fill_discrete(labels = (TIGER_RUCC_h1m  %>% filter(place %in% unique(TIGER_RUCC_h1m$match) ) %>% pull(County_Name)) )


H_match <- sparseMatrix(
  i = match(rownames(Sim), rural), 
  j = match(c(colnames(Sim)[apply(Sim, 1, which.min)]), primary),  
  x = 1L,
  dims = c(nrow(Sim), ncol(Sim)),
  dimnames = list(rural, primary)
) %>% as.matrix()


  
TIGER_RUCC %>% dplyr::filter(STATEFP!="15" & STATEFP!="02" ) %>% 
  tm_shape() +
  tm_fill(col = "H3",  
          title = "RUCC", 
          id = "NAMELSAD", 
          palette = "inferno", 
          n = 4, 
          contrast = c(.35, 1), 
          style = "fixed", 
          breaks = c(1, 2, 3, 4), 
          labels = c("1", "2", "3")
          )

```






```{r}
# Display tools
#display.brewer.all(n=NULL, type="all", select=NULL, exact.n=TRUE,  colorblindFriendly=FALSE)
#display.brewer.pal()
#palette_explorer()

  
TIGER_RUCC %>% dplyr::filter(STATEFP!="15" & STATEFP!="02" ) %>% 
  tm_shape() +
  tm_fill(col = "RUCC_2013",  
          title = "RUCC", 
          id = "NAMELSAD", 
          palette = "inferno", 
          n = 10, 
          contrast = c(.35, 1), 
          style = "fixed", 
          breaks = c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10), 
          labels = c("1", "2", "3", "4", "5", "6", "7", "8", "9")
          )
  
  

tmap_last()

```


```{r}
ggplot(filter(TIGER_RUCC, STATEFP=="55")) +
  geom_sf(aes(fill = RUCC_2013), color = NA) +
  scale_fill_viridis_c(option = "C",
                       limits = c(0, 10),
                       breaks  = c(1, 3, 5, 7, 9),
                       labels = c("1", "3", "5", "7", 9)) + 
  coord_sf() +
  theme_void() +
  labs(title = "              Wisconsin Rural-urban Continuum",
       fill = "RUCC") 
```





FIN
::: 
















