---
title: "&nbsp;"
format:
  html:
    self-contained: true
    page-layout: full
    code-fold: true
    code-tools: true
    code_download: yes
    latex_engine: pdflatex
---

```{r preamble, include = FALSE}

############ When you click the **Render** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document.
############ The params:county_fips:"..." option in YAML header allows the user to select a county for the notebook to analyze  


############ Set chunk behavior 
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 

############ For nonscientific notation
options(scipen=999)

```


```{r packages and libraries, include = FALSE}

############ Load and attach necessary packages
library(rprojroot)
library(tidyr)
library(Matrix)
library(dplyr)

############ To add in future for more capabilities
 library(ggplot2)
 library(ggiraph)
 library(glue)
 library(RColorBrewer)
 library(viridis)
 library(reshape2)
 library(scales)
 library(cowplot)
 library(ggnewscale)

############ Load custom functions
source(file.path(find_rstudio_root_file(), "nbs", "rural_typology_r_functions.R"))

```

```{r parameters, include = FALSE}
#Initialize parameter values
#params

```

::: panel-tabset

### National Absorption Maps

```{r national connectedness}

#####Specify the parameter options for national scope maps

#Year of CBP industry output (annual payroll), from ("1986" through "2020") 
cbp_year = "2012" 

#Minimum matching absorption cutoff, typically only specified if absorption is normalized, range [0,1], default == 0.05
threshold = .05 
#NULL, otherwise user specified impedance structure applied to absorption matrix, default == NULL
#e.g. "power_impedance_mat", "expo_impedance_mat",  "hyper_impedance_mat", "dprox_mat", "bprox_mat", "dist_mat"
impedance = NULL
#TRUE or FALSE, is county-by-county absorption matrix normalized by column-sums of the NIS matrix, default == TRUE
normalized = TRUE 
#TRUE or FALSE, row-match for forward supply linkages column-match for backward demand linkages in matching calculation of county-by-county absorption matrix, default == TRUE
row_max_match = TRUE 
#TRUE or FALSE, to aggregate shapefiles and industry output data to start with cbsa clusters or counties
cbsa_clust = FALSE
#BEA industry level specification, one of ("det", "sum", "sec"), default == "det"
ilevel = "det"
#TRUE or FALSE, to add html image data of industry distribution characteristics to dataframe, if TRUE can significantly increase processing time if done for the first time, default == FALSE 
add_html = FALSE

df <- spatial_connectedness(cbp_year = cbp_year,
                            threshold = threshold,
                            impedance = impedance,
                            normalized = normalized,
                            row_max_match = row_max_match,
                            cbsa_clust = cbsa_clust,
                            ilevel = ilevel,
                            add_html = add_html)

```


```{r cluster map}

#variable name for alpha value in map hover tooltip, default == "max_absorption_alpha"
#e.g., "max_absorption_alpha" "absorption_alpha_gini" "absorption_alpha_total" "absorption_alpha_mean" "absorption_alpha_sd" or a user specified transformation
alpha = "max_absorption_alpha"

#TRUE or FALSE, to add html image data of industry {output} distribution to map tooltip hover, if TRUE may prevent successful html render of whole notebook, default == FALSE
add_html = FALSE 

clustmember_map(df = df,
                alpha = alpha,
                add_html = add_html)

```

National cluster membership. \
`r cbp_year` CBP data. \
`r year2agcensus(cbp_year)` Farm sales data. \
`r year2bea(cbp_year, ilevel)` BEA requirements tables. \
`r year2tiger(cbp_year)` county boundary year. \
`r if(isTRUE(normalized)){"Normalized"}else{"Nominal"}` Absorption. \
`r if(ilevel == "det"){"Detail"}else if(ilevel == "sum"){"Summary"}else if(ilevel == "sec"){"Sector"}` level industry structure. \
`r if(isTRUE(row_max_match)){"Export Excess (row-wise)"}else{"Import Needs (column-wise)"}` matching. \
No spatial impedance. \

```{r absorption map 1}

#Variable name for alpha value in map hover tooltip
#e.g., one of "max_absorption_alpha" "absorption_alpha_gini" "absorption_alpha_total" "absorption_alpha_mean" "absorption_alpha_sd" or a user specified transformation
fill = "max_absorption_alpha"

#Display label in map
fill_lab = "Max Absorption"

#TRUE or FALSE, to add html image data of industry {output} distribution to map tooltip hover, if TRUE may prevent successful html render of whole notebook, default == FALSE
add_html = FALSE 

#TRUE or FALSE, set color scale min and max to (0,1) or set color scale min and max to fill variable min and max, default == TRUE
unit_scale = TRUE

absorption_map(df = df,
               add_html = add_html,
               fill = fill, 
               fill_lab = fill_lab,
               unit_scale = unit_scale)


```

`r fill_lab` connectedness. \
`r cbp_year` CBP data. \
`r year2agcensus(cbp_year)` Farm sales data. \
`r year2bea(cbp_year, ilevel)` BEA requirements tables. \
`r year2tiger(cbp_year)` county boundary year. \
`r if(isTRUE(normalized)){"Normalized"}else{"Nominal"}` Absorption. \
`r if(ilevel == "det"){"Detail"}else if(ilevel == "sum"){"Summary"}else if(ilevel == "sec"){"Sector"}` level industry structure. \
`r if(isTRUE(row_max_match)){"Export Excess (row-wise)"}else{"Import Needs (column-wise)"}` matching. \
No spatial impedance.


```{r absorption map 2}

#Variable name for alpha value in map hover tooltip
#e.g., one of "max_absorption_alpha" "absorption_alpha_gini" "absorption_alpha_total" "absorption_alpha_mean" "absorption_alpha_sd" or a user specified transformation
fill = "absorption_alpha_gini"

#Display label in map
fill_lab = "Absorption Gini"

#TRUE or FALSE, to add html image data of industry {output} distribution to map tooltip hover, if TRUE may prevent successful html render of whole notebook, default == FALSE
add_html = FALSE 

#TRUE or FALSE, set color scale min and max to (0,1) or set color scale min and max to fill variable min and max, default == TRUE
unit_scale = TRUE

absorption_map(df = df,
               add_html = add_html,
               fill = fill, 
               fill_lab = fill_lab,
               unit_scale = unit_scale)

```

`r fill_lab` connectedness. \
`r cbp_year` CBP data. \
`r year2agcensus(cbp_year)` Farm sales data. \
`r year2bea(cbp_year, ilevel)` BEA requirements tables. \
`r year2tiger(cbp_year)` county boundary year. \
`r if(isTRUE(normalized)){"Normalized"}else{"Nominal"}` Absorption. \
`r if(ilevel == "det"){"Detail"}else if(ilevel == "sum"){"Summary"}else if(ilevel == "sec"){"Sector"}` level industry structure. \
`r if(isTRUE(row_max_match)){"Export Excess (row-wise)"}else{"Import Needs (column-wise)"}` matching. \
No spatial impedance. \
A Gini of 0 is no concentration and Gini of 1 is complete concentration. 



### Hierarchical Absorption Maps

```{r hierarchial maps}
df <- one_hierarchical_connectedness(cbp_year = "2012", ilevel = "det")
```

```{r hierarchial impedance clust maps}

folder = "national_hierarcy_clust"
map_function = clustmember_map
national_hierarchy_gif(df = df$Hierarchical_Connectedness_table, 
                      folder = folder,
                      map_function = map_function)
file.path(find_rstudio_root_file(), "data", folder) %>% 
  list.files(full.names = T, pattern = "\\.gif$") %>% 
  image_read()

```

```{r hierarchial impedance abs maps}

folder = "national_hierarcy_abs"
map_function = absorption_map
national_hierarchy_gif(df = df$Hierarchical_Connectedness_table, 
                      folder = folder,
                      map_function = map_function)
file.path(find_rstudio_root_file(), "data", folder) %>% 
  list.files(full.names = T, pattern = "\\.gif$") %>% 
  image_read()

```

```{r hierarchial clustering L1}
clustmember_map(df = df$Hierarchical_Connectedness_table$level_1,
                caption = glue("Level 1 Hierarchy\n {threshold*100}% Isolation Threshold"))
```

```{r hierarchial absorption L1}
absorption_map(df = df$Hierarchical_Connectedness_table$level_1,
               caption = glue("Level 1 Hierarchy\n {threshold*100}% Isolation Threshold"))
```


### Progressive Distance Impedance Maps

```{r progressive distance impedance clust maps}
folder = "national_progressive_impedance_cluster_bdist"
if(!file.exists(file.path(find_rstudio_root_file(), "data", folder, paste0(folder, ".gif")))){
national_progressiveimpedance_gif(cbp_year,
                                   dist,
                                   impd = "B",
                                   folder = folder,
                                   map_function = clustmember_map)
}
file.path(find_rstudio_root_file(), "data", folder) %>%
  list.files(full.names = T, pattern = "\\.gif$") %>%
  image_read()
```


National cluster membership. \
`r cbp_year` CBP data. \
`r year2agcensus(cbp_year)` Farm sales data. \
`r year2bea(cbp_year, ilevel)` BEA requirements tables. \
`r year2tiger(cbp_year)` county boundary year. \
`r if(isTRUE(normalized)){"Normalized"}else{"Nominal"}` Absorption. \
`r if(ilevel == "det"){"Detail"}else if(ilevel == "sum"){"Summary"}else if(ilevel == "sec"){"Sector"}` level industry structure. \
`r if(isTRUE(row_max_match)){"Export Excess (row-wise)"}else{"Import Needs (column-wise)"}` matching. \


```{r progressive distance impedance abs maps}
folder = "national_progressive_impedance_absorption_bdist"
if(!file.exists(file.path(find_rstudio_root_file(), "data", folder, paste0(folder, ".gif")))){
national_progressiveimpedance_gif(cbp_year,
                                   dist,
                                   impd = "B",
                                   folder = folder,
                                   map_function = absorption_map)
}
file.path(find_rstudio_root_file(), "data", folder) %>%
  list.files(full.names = T, pattern = "\\.gif$") %>%
  image_read()
```

`r fill_lab` connectedness. \
`r cbp_year` CBP data. \
`r year2agcensus(cbp_year)` Farm sales data. \
`r year2bea(cbp_year, ilevel)` BEA requirements tables. \
`r year2tiger(cbp_year)` county boundary year. \
`r if(isTRUE(normalized)){"Normalized"}else{"Nominal"}` Absorption. \
`r if(ilevel == "det"){"Detail"}else if(ilevel == "sum"){"Summary"}else if(ilevel == "sec"){"Sector"}` level industry structure. \
`r if(isTRUE(row_max_match)){"Export Excess (row-wise)"}else{"Import Needs (column-wise)"}` matching. \


Distance from Border Qualitative ECAs
```{r }
cbp_year = "2012"
dist = c("5", "10", "25", "50", "75", "100", "150", "200", "250", "300", "400", "500", "750", "1000", "1250", "1500", "2000", "2500", "3000")

sasf <- paste0("spacecon", "_", ilevel, "class", "_", cbp_year, "cbp", "_", "PDI")
  if (file.exists(file.path(find_rstudio_root_file(), "data", "robjs", sasf) ) ){ 
    df <- readRDS(file.path(find_rstudio_root_file(), "data", "robjs", sasf))
  } else {
    df <- list()
    tiger_year = year2tiger(cbp_year)
    for(d in dist){
      impedance = dist_matb(dist = miles2meters(d), 
                            tiger_year = tiger_year)
      df[[d]] <- spatial_connectedness(cbp_year = cbp_year,
                                        impedance = impedance)
    }
    saveRDS(df,  file = file.path(find_rstudio_root_file(), "data", "robjs", sasf) )
  }
 
 
```


```{r}
aqual_map(df = df[["5"]], caption = paste0("5", " Mile Impedance"))
aqual_map(df = df[["10"]], caption = paste0("10", " Mile Impedance"))
aqual_map(df = df[["25"]], caption = paste0("25", " Mile Impedance"))
aqual_map(df = df[["50"]], caption = paste0("50", " Mile Impedance"))
aqual_map(df = df[["75"]], caption = paste0("75", " Mile Impedance"))
aqual_map(df = df[["100"]], caption = paste0("100", " Mile Impedance"))
aqual_map(df = df[["150"]], caption = paste0("150", " Mile Impedance"))
aqual_map(df = df[["200"]], caption = paste0("200", " Mile Impedance"))
aqual_map(df = df[["250"]], caption = paste0("250", " Mile Impedance"))
aqual_map(df = df[["300"]], caption = paste0("300", " Mile Impedance"))
aqual_map(df = df[["400"]], caption = paste0("400", " Mile Impedance"))
aqual_map(df = df[["500"]], caption = paste0("500", " Mile Impedance"))
aqual_map(df = df[["750"]], caption = paste0("750", " Mile Impedance"))
aqual_map(df = df[["1000"]], caption = paste0("1000", " Mile Impedance"))
aqual_map(df = df[["1250"]], caption = paste0("1250", " Mile Impedance"))
aqual_map(df = df[["1500"]], caption = paste0("1500", " Mile Impedance"))
aqual_map(df = df[["2000"]], caption = paste0("2000", " Mile Impedance"))
aqual_map(df = df[["2500"]], caption = paste0("2500", " Mile Impedance"))
aqual_map(df = df[["3000"]], caption = paste0("3000", " Mile Impedance"))
```


### Hierarchical with Impedance Maps

All Matching Restricted to 5 Mile Border Buffer

```{r hierarchial distance impedance maps}
df <- one_hierarchical_connectedness(cbp_year = "2012", ilevel = "det",  impedance = 5)

#Minimum matching absorption cutoff, typically only specified if absorption is normalized, range [0,1], default == 0.05
threshold = .05 
#Display label in map
fill_lab = "Max Absorption"
```


```{r distance 5 impedance cluster map}
clustmember_map(df = df$Hierarchical_Connectedness_table$level_1,
                caption = glue("5 Mile Border Impedance \n {threshold*100}% Isolation Threshold"))
```

National cluster membership. \
`r cbp_year` CBP data. \
`r year2agcensus(cbp_year)` Farm sales data. \
`r year2bea(cbp_year, ilevel)` BEA requirements tables. \
`r year2tiger(cbp_year)` county boundary year. \
`r if(isTRUE(normalized)){"Normalized"}else{"Nominal"}` Absorption. \
`r if(ilevel == "det"){"Detail"}else if(ilevel == "sum"){"Summary"}else if(ilevel == "sec"){"Sector"}` level industry structure. \
`r if(isTRUE(row_max_match)){"Export Excess (row-wise)"}else{"Import Needs (column-wise)"}` matching. \
5 Mile Border Impedance. \

```{r distance 5 impedance absorption map}
absorption_map(df = df$Hierarchical_Connectedness_table$level_1,
               caption = glue("5 Mile Border Impedance \n {threshold*100}% Isolation Threshold"))
```

`r fill_lab` connectedness. \
`r cbp_year` CBP data. \
`r year2agcensus(cbp_year)` Farm sales data. \
`r year2bea(cbp_year, ilevel)` BEA requirements tables. \
`r year2tiger(cbp_year)` county boundary year. \
`r if(isTRUE(normalized)){"Normalized"}else{"Nominal"}` Absorption. \
`r if(ilevel == "det"){"Detail"}else if(ilevel == "sum"){"Summary"}else if(ilevel == "sec"){"Sector"}` level industry structure. \
`r if(isTRUE(row_max_match)){"Export Excess (row-wise)"}else{"Import Needs (column-wise)"}` matching. \
5 Mile Border Impedance. \


```{r hierarchial distance 5 impedance clust maps}
folder = "national_hierarcy_dist5_clust"
map_function = clustmember_map
national_hierarchy_gif(df = df$Hierarchical_Connectedness_table, 
                      folder = folder,
                      map_function = map_function)
file.path(find_rstudio_root_file(), "data", folder) %>% 
  list.files(full.names = T, pattern = "\\.gif$") %>% 
  image_read()

```

```{r hierarchial distance 5 impedance abs maps}
folder = "national_hierarcy_dist5_abs"
map_function = absorption_map
national_hierarchy_gif(df = df$Hierarchical_Connectedness_table, 
                      folder = folder,
                      map_function = map_function)
file.path(find_rstudio_root_file(), "data", folder) %>% 
  list.files(full.names = T, pattern = "\\.gif$") %>% 
  image_read()

```

All Matching Restricted to Shared Queen Borders

```{r hierarchial queen border impedance maps}
df <- one_hierarchical_connectedness(cbp_year = "2012", ilevel = "det",  impedance = "B", queen = TRUE)
```

```{r queen border impedance cluster map}
clustmember_map(df = df$Hierarchical_Connectedness_table$level_1,
                caption = glue("Queen Border Impedance \n {threshold*100}% Isolation Threshold"))
```

National cluster membership. \
`r cbp_year` CBP data. \
`r year2agcensus(cbp_year)` Farm sales data. \
`r year2bea(cbp_year, ilevel)` BEA requirements tables. \
`r year2tiger(cbp_year)` county boundary year. \
`r if(isTRUE(normalized)){"Normalized"}else{"Nominal"}` Absorption. \
`r if(ilevel == "det"){"Detail"}else if(ilevel == "sum"){"Summary"}else if(ilevel == "sec"){"Sector"}` level industry structure. \
`r if(isTRUE(row_max_match)){"Export Excess (row-wise)"}else{"Import Needs (column-wise)"}` matching. \
Queen Border spatial impedance. \

```{r queen border impedance absorption map}
absorption_map(df = df$Hierarchical_Connectedness_table$level_1,
               caption = glue("Queen Border Impedance \n {threshold*100}% Isolation Threshold"))
```

`r fill_lab` connectedness. \
`r cbp_year` CBP data. \
`r year2agcensus(cbp_year)` Farm sales data. \
`r year2bea(cbp_year, ilevel)` BEA requirements tables. \
`r year2tiger(cbp_year)` county boundary year. \
`r if(isTRUE(normalized)){"Normalized"}else{"Nominal"}` Absorption. \
`r if(ilevel == "det"){"Detail"}else if(ilevel == "sum"){"Summary"}else if(ilevel == "sec"){"Sector"}` level industry structure. \
`r if(isTRUE(row_max_match)){"Export Excess (row-wise)"}else{"Import Needs (column-wise)"}` matching. \
Queen Border spatial impedance. \

```{r hierarchial queen border impedance clust maps}
folder = "national_hierarcy_queenboundry_clust"
map_function = clustmember_map
national_hierarchy_gif(df = df$Hierarchical_Connectedness_table, 
                      folder = folder,
                      map_function = map_function)

file.path(find_rstudio_root_file(), "data", folder) %>% 
  list.files(full.names = T, pattern = "\\.gif$") %>% 
  image_read()

```

```{r hierarchial queen border impedance abs maps}
folder = "national_hierarcy_queenboundry_abs"
map_function = absorption_map

national_hierarchy_gif(df = df$Hierarchical_Connectedness_table, 
                      folder = folder,
                      map_function = map_function)

file.path(find_rstudio_root_file(), "data", folder) %>% 
  list.files(full.names = T, pattern = "\\.gif$") %>% 
  image_read()

```


All Matching Restricted to Shared Rook Borders

```{r hierarchial rook border impedance maps}
df <- one_hierarchical_connectedness(cbp_year = "2012", ilevel = "det",  impedance = "B", queen = FALSE)
```

```{r rook border impedance cluster map}
clustmember_map(df = df$Hierarchical_Connectedness_table$level_1,
                caption = glue("Rook Border Impedance \n {threshold*100}% Isolation Threshold"))
```

National cluster membership. \
`r cbp_year` CBP data. \
`r year2agcensus(cbp_year)` Farm sales data. \
`r year2bea(cbp_year, ilevel)` BEA requirements tables. \
`r year2tiger(cbp_year)` county boundary year. \
`r if(isTRUE(normalized)){"Normalized"}else{"Nominal"}` Absorption. \
`r if(ilevel == "det"){"Detail"}else if(ilevel == "sum"){"Summary"}else if(ilevel == "sec"){"Sector"}` level industry structure. \
`r if(isTRUE(row_max_match)){"Export Excess (row-wise)"}else{"Import Needs (column-wise)"}` matching. \
Rook Border spatial impedance. \

```{r rook border impedance absorption map}
absorption_map(df = df$Hierarchical_Connectedness_table$level_1,
               caption = glue("Rook Border Impedance \n {threshold*100}% Isolation Threshold"))
```

`r fill_lab` connectedness. \
`r cbp_year` CBP data. \
`r year2agcensus(cbp_year)` Farm sales data. \
`r year2bea(cbp_year, ilevel)` BEA requirements tables. \
`r year2tiger(cbp_year)` county boundary year. \
`r if(isTRUE(normalized)){"Normalized"}else{"Nominal"}` Absorption. \
`r if(ilevel == "det"){"Detail"}else if(ilevel == "sum"){"Summary"}else if(ilevel == "sec"){"Sector"}` level industry structure. \
`r if(isTRUE(row_max_match)){"Export Excess (row-wise)"}else{"Import Needs (column-wise)"}` matching. \
Rook Border spatial impedance. \

```{r hierarchial rook border impedance clust maps}
folder = "national_hierarcy_rookboundry_clust"
map_function = clustmember_map

national_hierarchy_gif(df = df$Hierarchical_Connectedness_table, 
                      folder = folder,
                      map_function = map_function)

file.path(find_rstudio_root_file(), "data", folder) %>% 
  list.files(full.names = T, pattern = "\\.gif$") %>% 
  image_read()

```

```{r hierarchial rook border impedance abs maps}
folder = "national_hierarcy_rookboundry_abs"
map_function = absorption_map

national_hierarchy_gif(df = df$Hierarchical_Connectedness_table, 
                      folder = folder,
                      map_function = map_function)

file.path(find_rstudio_root_file(), "data", folder) %>% 
  list.files(full.names = T, pattern = "\\.gif$") %>% 
  image_read()

```





### Top Level Hierarchical Qualitative ECA Maps

```{r hierarchial top qual map}
df <- one_hierarchical_connectedness(cbp_year = "2012", ilevel = "det")
aqual_map(df=df$Hierarchical_Connectedness_table[[names(df$Hierarchical_Connectedness_table)[length(names(df$Hierarchical_Connectedness_table))]]])
```

No Matching Impedance\
Top ECAs of `r length(names(df$Hierarchical_Connectedness_table))` Iterations\
Normalized Absorption Matrix\
Minimum of 0.05 Absorption Cutoff\

```{r hierarchical queen border impedance qual maps}
df <- one_hierarchical_connectedness(cbp_year = "2012", ilevel = "det",  impedance = "B", queen = TRUE)
aqual_map(df=df$Hierarchical_Connectedness_table[[names(df$Hierarchical_Connectedness_table)[length(names(df$Hierarchical_Connectedness_table))]]])
```

Matching Impedance Restricted to Shared (Queens) Borders\
Top ECAs of `r length(names(df$Hierarchical_Connectedness_table))` Iterations\
Normalized Absorption Matrix\
Minimum of 0.05 Absorption Cutoff\

```{r hierarchical rook border impedance qual maps}
df <- one_hierarchical_connectedness(cbp_year = "2012", ilevel = "det",  impedance = "B", queen = FALSE)
aqual_map(df=df$Hierarchical_Connectedness_table[[names(df$Hierarchical_Connectedness_table)[length(names(df$Hierarchical_Connectedness_table))]]])
```

Matching Impedance Restricted to Shared (Rooks) Borders\
Top ECAs of `r length(names(df$Hierarchical_Connectedness_table))` Iterations\
Normalized Absorption Matrix\
Minimum of 0.05 Absorption Cutoff\

```{r hierarchial 5 mile border distance impedance maps}
df <- one_hierarchical_connectedness(cbp_year = "2012", ilevel = "det", impedance = 5)
aqual_map(df=df$Hierarchical_Connectedness_table[[names(df$Hierarchical_Connectedness_table)[length(names(df$Hierarchical_Connectedness_table))]]])
```

Matching Impedance Restricted to 5 Mile Border Distance\
Top ECAs of `r length(names(df$Hierarchical_Connectedness_table))` Iterations\
Normalized Absorption Matrix\
Minimum of 0.05 Absorption Cutoff\

```{r hierarchial 50 mile border distance impedance maps}
df <- one_hierarchical_connectedness(cbp_year = "2012", ilevel = "det", impedance = 50)
aqual_map(df=df$Hierarchical_Connectedness_table[[names(df$Hierarchical_Connectedness_table)[length(names(df$Hierarchical_Connectedness_table))]]])
```

Matching Impedance Restricted to 50 Mile Border Distance\
Top ECAs of `r length(names(df$Hierarchical_Connectedness_table))` Iterations\
Normalized Absorption Matrix\
Minimum of 0.05 Absorption Cutoff\

```{r hierarchial 500 mile border distance impedance maps}
df <- one_hierarchical_connectedness(cbp_year = "2012", ilevel = "det", impedance = 500)
aqual_map(df=df$Hierarchical_Connectedness_table[[names(df$Hierarchical_Connectedness_table)[length(names(df$Hierarchical_Connectedness_table))]]])
```

Matching Impedance Restricted to 500 Mile Border Distance\
Top ECAs of `r length(names(df$Hierarchical_Connectedness_table))` Iterations\
Normalized Absorption Matrix\
Minimum of 0.05 Absorption Cutoff\


### National Absorption Statistics

```{r absorption distribution}

#####Specify year range for normalized and nominal absorption stats
years = 2000:2020

  if (file.exists(file.path(find_rstudio_root_file(), file.path("data", "robjs"), "normal_absorb_ot") ) ){
    normal_absorb_ot <- readRDS(file.path(find_rstudio_root_file(), file.path("data", "robjs"), "normal_absorb_ot"))
  } else {
    normal_absorb_ot <- absorption_match_overtime(years = years)
    saveRDS(normal_absorb_ot,  file = file.path(find_rstudio_root_file(), file.path("data", "robjs"), "normal_absorb_ot") )
  }

  # if (file.exists(file.path(find_rstudio_root_file(), file.path("data", "robjs"), "nominal_absorb_ot") ) ){
  #   nominal_absorb_ot <- readRDS(file.path(find_rstudio_root_file(), file.path("data", "robjs"), "nominal_absorb_ot"))
  # } else {
  #   nominal_absorb_ot <- absorption_match_overtime(years = years,
  #                                                   normalized = FALSE)
  #   saveRDS(nominal_absorb_ot,  file = file.path(find_rstudio_root_file(), file.path("data", "robjs"), "nominal_absorb_ot") )
  # }


```


```{r absorption density}


#Specify density fill and label
fill = "max_absorption_alpha"
fill_lab = "Max Absorption Alpha"

#TRUE or FALSE, label if using normal_absorb_ot or nominal_absorb_ot
normalized = TRUE
#NULL, otherwise a mathematical transformation of x-axis units 
#e.g., "log10"
trans = NULL
#colors for specific year ranges of interest
colorbreaks = c("#440154FF" = 2000:2006, "#1E9C89FF" = 2007:2016 , "#D64B40FF" = 2017, "#FDE725FF" =  2018:2020)

absorption_density_plot(df = normal_absorb_ot,
                        fill = fill,
                        fill_lab = fill_lab,
                        normalized = normalized,
                        trans = trans,
                        colorbreaks = colorbreaks)


```

`r fill_lab` density. \
`r if(isTRUE(normalized)){"Normalized"}else{"Nominal"}` Absorption. \
`r min(years)` to `r max(years)` CBP data. \
`r if(ilevel == "det"){"Detail"}else if(ilevel == "sum"){"Summary"}else if(ilevel == "sec"){"Sector"}` level industry structure. \
`r if(isTRUE(row_max_match)){"Export Excess (row-wise)"}else{"Import Needs (column-wise)"}` matching.


:::



