---
title: "&nbsp;"
format:
  html:
    self-contained: true
    page-layout: full
    code-fold: true
    code-tools: true
    code_download: yes
    latex_engine: pdflatex
params:
  county_fips: "21195"
---

```{r preamble, include = FALSE}

############ When you click the **Render** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document.
############ The params:county_fips:"..." option in YAML header allows the user to select a county for the notebook to analyze  


############ Set chunk behavior 
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 

############ For nonscientific notation
options(scipen=999)

```


```{r packages and libraries, include = FALSE}

############ Load and attach necessary packages
library(rprojroot)
library(tidyr)
library(Matrix)
library(dplyr)

############ To add in future for more capabilities
 library(ggplot2)
 library(ggiraph)
 library(glue)
 library(RColorBrewer)
 library(viridis)
 library(reshape2)
 library(scales)
 library(cowplot)
 library(ggnewscale)
library(qgraph)

############ Load custom functions
source(file.path(find_rstudio_root_file(), "nbs", "rural_typology_r_data.R"))

```

```{r parameters, include = FALSE}
#Initialize parameter values
params

#Year of shapefile boundary data, one of ("2000", "2010", or "2013" through "2020")
tiger_year = "2020" 

#Get county and state names of params:county_fips
t <- call_tiger(tiger_year, 
                geometry = FALSE)

county_name <- t$NAME[t$place == params$county_fips]
state_name <- t$STATE[t$place == params$county_fips]

```

::: panel-tabset

### `r county_name`, `r state_name` Industry Distributions

```{r industry distribution tables}

#####Specify industry output/farms sales data for industry distribution tables

#Year of CBP industry output (annual payroll), from ("1986" through "2020")
cbp_year <- "2017"

#BEA industry level specification, one of ("det", "sum", "sec"), default == "det"
ilevel = "det"
#industry level aggregation class, one of ("det", "sum", "sec"), default == "sec"
industry_aggregate_class = "sec"

```


```{r industry distribution plots}

industry_dist_plots(central_place = params$county_fips,
                    cbp_year = cbp_year,
                    ilevel = ilevel,
                    industry_aggregate_class = industry_aggregate_class,
                    interact = TRUE)

```

`r county_name`, `r state_name` industry activity distribution. \
`r cbp_year` CBP data. \
`r if(cbp_year > "2014"){"2017"}else if(cbp_year %in% 2014:2010){"2012"}else if(cbp_year %in% 2009:2005){"2007"}else if(cbp_year < "2005"){"2002"}` Farm sales data. \
`r if(ilevel != "det"){cbp_year}else{if(isTRUE(cbp_year > "2007")){"2012"}else{"2007"}}` BEA requirements tables. \
`r if(industry_aggregate_class == "det"){"Detail"}else if(industry_aggregate_class == "sum"){"Summary"}else if(industry_aggregate_class == "sec"){"Sector"}`-level industry structure aggregate from `r if(ilevel == "det"){"Detail"}else if(ilevel == "sum"){"Summary"}else if(ilevel == "sec"){"Sector"}`-level structure.



### `r county_name`, `r state_name` Economic Catchment Area


```{r national connectedness}

#####Specify the parameter options for national scope maps

#Year of CBP industry output (annual payroll), from ("1986" through "2020") 
cbp_year = "2012" 

#Year of shapefile boundary data, one of ("2000", "2010", or "2013" through "2020"), default == cbp_year
tiger_year = "2013" 


#Minimum matching absorption cutoff, typically only specified if absorption is normalized, range [0,1], default == 0.05
threshold = .05 
#NULL, otherwise user specified impedance structure applied to absorption matrix, default == NULL
#e.g. "power_impedance_mat", "expo_impedance_mat",  "hyper_impedance_mat", "dprox_mat", "bprox_mat", "dist_mat"
impedance = NULL
#TRUE or FALSE, is county-by-county absorption matrix normalized by column-sums of the NIS matrix, default == TRUE
normalized = TRUE 
#TRUE or FALSE, row-match for forward supply linkages column-match for backward demand linkages in matching calculation of county-by-county absorption matrix, default == TRUE
row_max_match = TRUE 
#NULL for county level aggregation, or Year of cbsa concordance to aggregate shapefiles and industry output data, one of ("2020", "2018", "2017", "2015", "2013", "2009", "2008", "2007", "2006", "2005", "2004", "2003"), default == NULL
cbsa_year = NULL 
#BEA industry level specification, one of ("det", "sum", "sec"), default == "det"
ilevel = "det"
#TRUE or FALSE, to add html image data of industry distribution characteristics to dataframe, if TRUE can significantly increase processing time if done for the first time, default == FALSE 
add_html = TRUE


df <- spatial_connectedness(cbp_year = cbp_year,
                            tiger_year = tiger_year,
                            threshold = threshold,
                            impedance = impedance, 
                            normalized = normalized,
                            row_max_match = row_max_match,
                            cbsa_year = cbsa_year,
                            ilevel = ilevel,
                            add_html = add_html)


```


```{r ECA map}

hier_ab_map(df = df,
            central_place = params$county_fips,
            threshold = threshold)

```


### `r county_name`, `r state_name` Absorption Maps

```{r place connectedness}

#####Specify the parameter options for place specific maps

#Year of CBP industry output (annual payroll), from ("1986" through "2020") 
cbp_year <- "2017"

#Year of shapefile boundary data, one of ("2000", "2010", or "2013" through "2020"), default == cbp_year
tiger_year = "2013" 

#NULL for county level aggregation, or Year of cbsa concordance to aggregate shapefiles and industry output data, one of ("2020", "2018", "2017", "2015", "2013", "2009", "2008", "2007", "2006", "2005", "2004", "2003"), default == NULL
cbsa_year = NULL 
#NULL, otherwise user specified impedance structure applied to absorption matrix, default == NULL
#e.g. "power_impedance_mat", "expo_impedance_mat",  "hyper_impedance_mat", "dprox_mat", "bprox_mat", "dist_mat"
impedance = NULL
#TRUE or FALSE, is county-by-county absorption matrix normalized by column-sums of the NIS matrix, default == TRUE
normalized = TRUE 
#BEA industry level specification, one of ("det", "sum", "sec"), default == "det"
ilevel = "det"
#TRUE or FALSE, to add html image data of industry distribution characteristics to dataframe, if TRUE can significantly increase processing time if done for the first time, default == FALSE 
add_html = TRUE


df <- place_centric_connect(central_place = params$county_fips,
                            cbp_year = cbp_year,
                            tiger_year = tiger_year,
                            impedance = impedance, 
                            normalized = normalized,
                            cbsa_year = cbsa_year,
                            ilevel = ilevel,
                            add_html = add_html)

```

`r county_name`, `r state_name` connectedness to the nation

```{r place absorption map 1}

#####Specify the display options for place specific maps

#row-match forward supply linkages or column-match backward demand linkages in absorption matrix, one of "export_absorption" or "import_absorption"
fill <- "export_absorption"

#TRUE or FALSE, to add html image data of industry {output} distribution to map tooltip hover, if TRUE may prevent successful html render of whole notebook, default == FALSE
add_html = FALSE 
#TRUE or FALSE, set color scale min and max to (0,1) or set color scale min and max to fill variable min and max, default == TRUE
unit_scale = TRUE

place_absorption_map(central_place = params$county_fips,
                     df = df,
                     fill = fill,
                     add_html = add_html,
                     unit_scale = unit_scale)

```

`r if(fill == "export_absorption"){"Export Excess (row-wise)"}else{"Import Needs (column-wise)"}` absorption. \
`r if(isTRUE(normalized)){"Normalized"}else{"Nominal"}` Absorption. \
`r cbp_year` CBP data. \
`r cbp_year` Farm sales data. \
`r if(ilevel != "det"){cbp_year}else{if(isTRUE(cbp_year > "2007")){"2012"}else{"2007"}}` BEA requirements tables. \
`r tiger_year` county boundary year. \
`r if(ilevel == "det"){"Detail"}else if(ilevel == "sum"){"Summary"}else if(ilevel == "sec"){"Sector"}` level industry structure. \
No spatial impedance.


```{r place absorption map 2}

#####Specify the display options for place specific maps

#row-match forward supply linkages or column-match backward demand linkages in absorption matrix, one of "export_absorption" or "import_absorption"
fill <- "import_absorption"

#TRUE or FALSE, to add html image data of industry {output} distribution to map tooltip hover, if TRUE may prevent successful html render of whole notebook, default == FALSE
add_html = FALSE 
#TRUE or FALSE, set color scale min and max to (0,1) or set color scale min and max to fill variable min and max, default == TRUE
unit_scale = TRUE

place_absorption_map(central_place = params$county_fips,
                     df = df,
                     fill = fill,
                     add_html = add_html,
                     unit_scale = unit_scale)

```

`r if(fill == "export_absorption"){"Export Excess (row-wise)"}else{"Import Needs (column-wise)"}` absorption. \
`r if(isTRUE(normalized)){"Normalized"}else{"Nominal"}` Absorption. \
`r cbp_year` CBP data. \
`r cbp_year` Farm sales data. \
`r if(ilevel != "det"){cbp_year}else{if(isTRUE(cbp_year > "2007")){"2012"}else{"2007"}}` BEA requirements tables. \
`r tiger_year` county boundary year. \
`r if(ilevel == "det"){"Detail"}else if(ilevel == "sum"){"Summary"}else if(ilevel == "sec"){"Sector"}` level industry structure. \
No spatial impedance.


### `r county_name`, `r state_name` Absorption Change Maps

```{r place connectedness delta}

#####Specify the parameter options for place specific maps - with changes over time

#Sample year of industry output data
sample_years = c("2017", "2012", "2007", "2002")
#combn(sample_years, 2) #table of combinations 

#Year of shapefile boundary data, one of ("2000", "2010", or "2013" through "2020"), default == "2013"
tiger_year = "2013" 
#NULL, otherwise user specified impedance structure applied to absorption matrix, default == NULL
#e.g. "power_impedance_mat", "expo_impedance_mat",  "hyper_impedance_mat", "dprox_mat", "bprox_mat", "dist_mat"
impedance = NULL
#TRUE or FALSE, is county-by-county absorption matrix normalized by column-sums of the NIS matrix, default == TRUE
normalized = TRUE 
#BEA industry level specification, one of ("det", "sum", "sec"), default == "det"
ilevel = "det"
#TRUE or FALSE, to add html image data of industry distribution characteristics to dataframe, if TRUE can significantly increase processing time if done for the first time, default == FALSE
add_html = FALSE

df <- place_connect_delta(central_place = params$county_fips,
                          sample_years = sample_years,
                          tiger_year = tiger_year,
                          impedance = impedance, 
                          normalized = normalized,
                          ilevel = ilevel,
                          add_html = add_html)

#Set scale min and max to include extremes across all years 
delta_min <- df %>% select( grep("delta", names(df), value = TRUE)) %>% st_set_geometry(., NULL) %>% min()
delta_max <- df %>% select( grep("delta", names(df), value = TRUE)) %>% st_set_geometry(., NULL) %>% max()

```


```{r place absorption delta map 1}

#####Specify the display options for place specific maps over time

#Starting year of place specific absorption, one of "sample_years" specified above
y1 <- "2012"

#Ending year of place specific absorption, one of "sample_years" specified above
y2 <- "2017"

#row-match forward supply linkages or column-match backward demand linkages in absorption matrix, one of "export_absorption_delta" or "import_absorption_delta"
fill <- "export_absorption_delta" 


absorption_delta_map(central_place = params$county_fips,
                     df = df,
                     fill = paste0(fill, "_", substr(y1, 3, 4), substr(y2, 3, 4)),
                     delta_min = delta_min,
                     delta_max = delta_max)

```

`r y1` to `r y2`. \
`r if(fill == "export_absorption_delta"){"Export Excess (row-wise)"}else{"Import Needs (column-wise)"}` absorption. \
`r if(isTRUE(normalized)){"Normalized"}else{"Nominal"}` Absorption. \
`r tiger_year` county boundary year. \
`r if(ilevel == "det"){"Detail"}else if(ilevel == "sum"){"Summary"}else if(ilevel == "sec"){"Sector"}` level industry structure. \
No spatial impedance. \



```{r place absorption delta map 2}

#####Specify the display options for place specific maps over time

#Starting year of place specific absorption, one of "sample_years" specified above
y1 <- "2002"

#Ending year of place specific absorption, one of "sample_years" specified above
y2 <- "2017"

#row-match forward supply linkages or column-match backward demand linkages in absorption matrix, one of "export_absorption_delta" or "import_absorption_delta"
fill <- "export_absorption_delta" 


absorption_delta_map(central_place = params$county_fips,
                     df = df,
                     fill = paste0(fill, "_", substr(y1, 3, 4), substr(y2, 3, 4)),
                     delta_min = delta_min,
                     delta_max = delta_max)

```

`r y1` to `r y2`. \
`r if(fill == "export_absorption_delta"){"Export Excess (row-wise)"}else{"Import Needs (column-wise)"}` absorption. \
`r if(isTRUE(normalized)){"Normalized"}else{"Nominal"}` Absorption. \
`r tiger_year` county boundary year. \
`r if(ilevel == "det"){"Detail"}else if(ilevel == "sum"){"Summary"}else if(ilevel == "sec"){"Sector"}` level industry structure. \
No spatial impedance. \


:::



