---
title: "&nbsp;"
format:
  html:
    self-contained: true
    page-layout: full
    code-fold: true
    code-tools: true
    code_download: yes
    latex_engine: pdflatex
params:
  sna_state: "24"
  sna_place: "24510"
---

```{r preamble, include = FALSE}

############ When you click the **Render** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document.
############ The params:county_fips:"..." option in YAML header allows the user to select a county for the notebook to analyze  

############ Set chunk behavior 
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 

############ For nonscientific notation
options(scipen=999)

```


```{r packages and libraries, include = FALSE}

############ Load and attach necessary packages
library(rprojroot)
library(tidyr)
library(Matrix)
library(dplyr)

############ To add in future for more capabilities
 library(ggplot2)
 library(ggiraph)
 library(glue)
 library(RColorBrewer)
 library(viridis)
 library(reshape2)
 library(scales)
 library(cowplot)
 library(ggnewscale)
library(qgraph)

############ Load custom functions
source(file.path(find_rstudio_root_file(), "nbs", "rural_typology_r_functions.R"))

```

```{r parameters, include = FALSE}
#Initialize parameter values
params

```

::: panel-tabset

### Social Network Analysis

Social Network Analysis (SNA):

In network parlance the social actors (e.g. counties) are nodes or vertices and the ties between them are edges. The data for a social network can be organized as an adjacency matrix, with non-zero entries in the $i,j^{th}$  entry of the $i^{th}$ node shares an edge with the $j^{th}$ node.

If the edges have weights, the entries in the non-zero entries are the weights for the edges. If the edges are not weighted, the non-zero entries are simply one.
Directed graphs have edges originating from one node and ending in another. If the network is undirected this matrix is symmetric.


```{r sna}

#####Specify

#Year of CBP industry output (annual payroll), from ("1986" through "2020") 
cbp_year <- "2017"

#TRUE or FALSE, is county-by-county absorption matrix normalized by column-sums of the NIS matrix, default == TRUE
normalized = TRUE 
#NULL, otherwise user specified impedance structure applied to absorption matrix, default == NULL
#e.g. "power_impedance_mat", "expo_impedance_mat",  "hyper_impedance_mat", "dprox_mat", "bprox_mat", "dist_mat"
impedance = NULL
#BEA industry level specification, one of ("det", "sum", "sec"), default == "det"
ilevel = "det"
#TRUE or FALSE, to aggregate shapefiles and industry output data to start with cbsa clusters or counties
cbsa_clust = FALSE 

conmat <- absorption_matrix(cbp_year = cbp_year,
                           normalized = normalized,
                           impedance = impedance,
                           ilevel = ilevel,
                           cbsa_clust = cbsa_clust)

```

```{r sample sna}

#Specify state sub sample
df <-  grep(paste0("^", params$sna_state), rownames(conmat)) %>% conmat[., .] 
  
```


```{r qgraph}

#Edges with absolute weight that are not above this value are REMOVED from the network.
threshold = 0.05

Q <- qgraph(df, 
            threshold = threshold, 
            layout = 'spring', 
            vsize = 4, 
            labels = colnames(df))

```

Network plot of directed and weighted state sub-sample absorption matrix with minimium `r threshold` edge weight threshold.

```{r flow graph}

Q <- qgraph(df,
            threshold = 0,
            layout = 'spring',
            vsize = 4,
            labels = colnames(df)) %>%
      flow(from = params$sna_place)


#centrality_auto(Q, weighted = TRUE, signed = TRUE)
#centrality(Q)
```

Flow to and from `r params$sna_place`

```{r centrality plot 1}

centralityPlot(Q, include = c("Betweenness","Closeness"), decreasing = TRUE)
```

Betweeness:  measure of the extent to which the node connects disparate group. A node’s betweenness is the number of shortest paths the node is included in divided by the total number of shortest paths.

Closeness: measures how well connected a node is to every other node in the network. A node’s closeness is the average number of hops required to reach every other node in the network. 


```{r centrality plot 2}
centralityPlot(Q, include = c("InStrength", "OutStrength"), decreasing = TRUE)
```

Degree is the number of edges for a node. In a directed network, edges have a starting and end node, thus there are both the in-degree (number of edges the node is an end node of), out-degree (number of edges a node is a starting node of).

InStrength (weighted in-degree): Strength of incoming edges (pointing to the node). 

OutStrength (weighted out-degree): Strength of outgoing edges (pointing away from the node).


```{r centrality plot 3}
centralityPlot(Q, include = c("OutExpectedInfluence", "InExpectedInfluence"), decreasing = TRUE)

```

InExpectedInfluence: Expected incoming influence - sum of incoming edge weights connected to a node (not made absolute).

OutExpectedInfluence: Expected outgoing influence - sum of outgoing edge weights connected to a node (not made absolute).


[SNA INTRO](http://www.stat.yale.edu/~blc3/SNA2016/Intro_SNA_Presentation.pdf) 
[SNA BEGIN](https://builtin.com/data-science/social-network-analysis)


:::



