---
title: "paa spatial regress"
format:
  html:
    toc: true
    toc-depth: 1
    self-contained: true
    page-layout: full
    code-fold: true
    code-tools: true
    code_download: yes
    latex_engine: pdflatex
params:
  year: 2012
  panel_data: FALSE
  year_range: !expr c(2012, 2017) # or c(2012, 2015, 2017) or c(2010:2020)
  normalize_w: TRUE
  reg_factors: !expr c("establishments_gr", "wage_gr", "jobs_gr", "payroll_gr", "employment_gr", "poverty_rate_gr", "income_rate_gr", "population_gr", "entry_rate", "exit_rate") 
  reg_factors_names: !expr c("Establishment Growth", "Wage Growth", "Job Growth", "Payroll Growth", "Employment Growth", "Poverty Growth", "Income Growth", "Population Growth", "Entry Rate", "Exit Rate")
  reg_factor_funs: !expr c("call_establishments", "call_wage", "call_jobs", "call_payroll", "call_employment", "call_poverty_rate", "call_income_rate", "call_population")
  spec_names: !expr c("ols", "ols_weight", "ols_naive", "slm_ml", "slm_iv", "sem_ml", "sem_weight", "sarar", "slm_di", "sem_di", "regime_ols", "regime_slm", "regime_sem", "rural_ols", "rural_ols_naive", "rural_slm", "rural_sem", "nonrural_ols", "nonrural_ols_naive", "nonrural_slm", "nonrural_sem")
  w_names: !expr c("inverse", "square", "exponential", "gaussian", "qnn", "knn", "proximity", "bi-square" )
---


```{r preamble, include = FALSE}

# additional library list
library(logger)
library(tidyverse)

library(moments)
library(PerformanceAnalytics)
library(correlation)

library(estimatr)
library(margins)

library(modelsummary)

library(spatialreg)

library(hspm)
library(sphet)

library(kableExtra)


# scripts
source("projects/eca_paa/dataprep.R", local = (dp <- new.env()))

# chunk behavior 
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 

# nonscientific notation
options(scipen=999)

# log output 
log_threshold(DEBUG)

# initialize parameters
params

```

```{r, include = FALSE}

# uncomment and pack/unpack as necessary
# dp$zip_pack("tmp/eca_paa_data_v240324.zip", dp$opath)
# dp$zip_unpack("tmp/eca_paa_data_v240324.zip", overwrite = TRUE)

proj_path <- "data/projects/eca_paa/space_regress/"



```


```{r plot vis functions}


twg <- "trade_weighted_growth"

spec_f <- function(spec_list, mn = "ols", wm = "none"){
  df <- data.frame()
  for (i in rf){
    estimate <-  spec_list[[i]][["coefficients"]][[2]]
    lower <- confint(spec_list[[i]])[twg, 1]
    upper <- confint(spec_list[[i]])[twg, 2]
    wmatrix <- wm
    model <- mn
    indicator <- coef_names[match(i, rf)]
    df = rbind(df, data.frame(model, wmatrix, indicator, estimate, lower, upper))
  }
  return(df)
}


spec_df <- function(spec_list, mn = "ols", wm = NULL){
  if(is.null(wm)){
    df <- data.frame()
    for (j in params$w_names){
      x <- spec_list[[match(j, params$w_names)]] %>% 
        {spec_f(spec_list = ., mn = mn, wm = j)}
      df <- rbind(df, x)
      }
  } else {
    df <- spec_f(spec_list = spec_list, mn = mn, wm = wm)
  }
return(df)
}

point_plot_factor <- function(df){
  df$model <- factor(df$model)
  df$wmatrix <- factor(df$wmatrix, levels=c("none", params$w_names))
  df$indicator <- factor(df$indicator, levels=c(coef_names), labels = gsub(" Growth*|Rate*", '', coef_names))
  return(df)
}


# c("none"="black", 
#   "inverse"="#a6cee3", 
#   "square"="#1f78b4",
#   "exponential"="#b2df8a", 
#   "gaussian"="#33a02c", 
#   "qnn"="#fb9a99", 
#   "knn"-"#e31a1c", 
#   "proximity"="#fdbf6f", 
#   "bi-square"="#ff7f00")

point_plot <- function(df, error = TRUE){
  g <- df %>% 
  ggplot(aes(x=indicator, y=estimate, colour=wmatrix, shape=model,
             tooltip = wmatrix,
             data_id = wmatrix,
             `data-id` = wmatrix)) +
    geom_point_interactive(position = position_dodge(width=.3)) +
    scale_color_manual_interactive(values=c(
      "none"="black", 
      "inverse"="#a6cee3", 
      "square"="#1f78b4",
      "exponential"="#b2df8a", 
      "gaussian"="#33a02c", 
      "qnn"="#fb9a99", 
      "knn"="#e31a1c", 
      "proximity"="#fdbf6f", 
      "bi-square"="#ff7f00"),
                                            extra_interactive_params = "data-id",
                                            `data-id` = function(x) x, 
                                            data_id = function(x) x) +
    labs(x = "Growth Indicator", y = "Coeffecient Estimate", color = "W Matrix", shape = "Model") + 
    geom_hline(yintercept = 0) +
    theme_minimal() +
      theme(axis.text.x = element_text(size = rel(.75), angle = 300, hjust = 0),
            axis.text.y = element_text(size = rel(.75)), legend.margin = margin(r=10, l=5, t=1, b=1) )
  if (error){g <- g + geom_errorbar_interactive(aes(x=indicator, y=estimate, ymin=lower, ymax=upper), width=.2, position = position_dodge(width=.3), data = df)}
  return(g)
}

coef_point_plot <- function(df, error = TRUE){
  g <- point_plot_factor(df) %>% 
    point_plot(error = error) %>% 
    dp$girafe_plot()
  g[["x"]][["settings"]][["hover"]][["css"]] <- ".hover_data_SVGID_ { stroke:orange;r:4pt; }\ntext.hover_data_SVGID_ { stroke:none;fill:black;fill-opacity:1; }"
  return(g)
}

```

# data

## dataframe
```{r}

# optional panel data format 
if (params$panel_data){
  df <- dp$call_temporal_range_df(params$year_range)
} else{
  df <- dp$call_proj_df(params$year)
}

# scale variables for better summary and coefficient interpretation
  df <- df %>% 
  mutate(eca_center_distance = dp$meters2miles(eca_center_distance)/100) %>% 
  mutate(wage = wage/1000) %>% 
  mutate(income_rate = income_rate/1000) %>% 
  mutate(ypll75 = ypll75/100000) %>% 
  mutate(eca_cluster_category = ifelse(eca_cluster_category == "Cluster Source", "Source", "Destination"))
  
# adjust names to "source" and "destination" vocabulary
df$metro_micro_destination <- df$metro_micro_sink
  df$metro_micro_destination[df$metro_micro_destination == "Sink and Metro "] <-  "Destination and Metro"
    df$metro_micro_destination[df$metro_micro_destination == "Sink and Micro"] <-  "Destination and Micro"
      df$metro_micro_destination[df$metro_micro_destination == "Sink only"] <-  "Destination Only"

```


## growth rates 

```{r}

if (!params$panel_data){
  x <- dp$temporal_permutations(params$year_range)
  tmp <- df$place %>% as.data.frame() %>% `colnames<-`(c("place"))
  for (y in 1:nrow(x)){
    paste(x[y,1], x[y,2]) %>% print()
    for (i in params$reg_factor_funs){
       tmp <- dp$growth_rate(x[y,1], x[y,2], dp[[i]]) %>%
         mutate(grow_rate = replace(grow_rate, which(abs(grow_rate) == 200), NA)) %>%
         rename_at('grow_rate', ~paste0(substring(i, 6), "_gr")) %>%
         # TODO: develop for more than two year growth delta
         # rename_at('grow_rate', ~paste0(substring(i, 6), "_gr_", x[y,1], "_", x[y,2])) %>% 
         left_join(tmp, ., by = "place")
    }
  }
  df <- left_join(df, tmp, by = "place")
  rm(x, tmp)
}

```


## circularity

```{r}

df <- dp$call_circularity_metrics(
  year = params$year, 
  paradigm = "domestic", 
  class_system = "commodity", 
  bus_data = "infogroup", 
  spatial = FALSE) %>% 
  select(!(gross_output:extract)) %>% 
  left_join(df, ., by = "place")

```


## variables

```{r}

# list out all numeric variables
df %>% 
  st_drop_geometry() %>% 
  select(where(is.numeric)) %>% 
  names() 

# set of economic indicator growth variables 
rf <- params$reg_factors

# specify display names of growth variables 
coef_names <- params$reg_factors_names

# dummy check for user variable and label length consistency
stopifnot("variable selection and names do not match" = length(rf) == length(coef_names))

```


## trade flow matrix

```{r}

# row normalized trade value matrix of potential excess trade flows between all places
tm <- dp$call_trade_flows("all_industries") %>% 
  dp$row_normalize()

```


## location sample 

```{r}

# all place names common between trade flow data and economic indicator growth data
pn <- tm %>% 
  {intersect(colnames(.), df$place)}

# subset of places with complete data across all growth variables 
# ensures constant N across all regressions
# allows for efficient construction of sptaialreg listw attribute (necessary)
pn <- df[, c("place", rf)] %>% 
  na.omit() %>% 
  {intersect(pn, .$place)} 

# sub-subset of places where not all neighbors get dropped from previous subset step 
pn <- pn %>% 
  {dp$call_bprox_mat(year = params$year, queen = TRUE)[., .]} %>% 
  {rownames(.)[rowSums(.) > 1]}

# rurality sub-sets
rural_places <-  df %>% 
  {.[.$place %in% pn, ]} %>% 
  {.[.$cbsa_rural_category == "rural", ]} %>% 
  {.[["place"]]}

nonrural_places <-  df %>% 
  {.[.$place %in% pn, ]} %>% 
  {.[.$cbsa_rural_category != "rural", ]} %>% 
  {.[["place"]]}

```


## regression dataset

```{r}

rdf <- df[df$place %in% pn, c("place", "cbsa_rural_category", "trade_capacity", rf)] %>% 
  st_drop_geometry() %>% 
  as.data.frame() 

for(i in rf){
  rdf <- pn %>% 
    {tm[., .] %*% as.matrix(rdf[i])} %>% 
    `colnames<-`(c(paste0(i, "_", twg))) %>% 
    {cbind(rdf, .)}
  }

```


## W matrix

```{r}

## Does the model need to be extended with spatial interaction effects? 
## Is there considerable spatial dependence in these data such that a standard OLS regression that assumes independent observations will be misleading?

W_list <- dp$imped_list(
  pn_dim = pn, 
  dmat = dp$call_dist_mat(year = params$year) )

```


## W edge list

```{r}

cache_path = glue(proj_path, "W_list_alt.rds")
  if (file.exists(cache_path)) {
    W_list_alt <- readRDS(cache_path)
  } else {
    W_list_alt <- list()
    for (j in 1:length(W_list)){
      W_list_alt[[j]] <- intersect(rownames(W_list[[j]]), pn) %>% 
        {W_list[[j]][., .]} %>% 
        mat2listw(style="M", zero.policy=TRUE) 
    }
    saveRDS(W_list_alt, dp$util$mkdir(cache_path))
  }

```



# regression


## OLS - normalized trade

```{r results='asis'}

# Does growth in trade partner export destination(s) correspond with own growth?

# y_growth = b0 + b1_r + e
# s.t. r = M * y_growth

# Row normalized trade values
cache_path = glue(proj_path, "ols.rds")
if (file.exists(cache_path)) {
  specif <- readRDS(cache_path)
} else {
  specif <- list()
  for (i in rf){
    df_reg <- rdf[, c(i, paste0(i, "_", twg))] %>% 
      `colnames<-`(c(i, twg))
    fm <- as.formula(paste(i, "~", twg))
    specif[[i]] <- lm(fm, data = df_reg)
  }
  saveRDS(specif, dp$util$mkdir(cache_path))
}

  specif %>%
  `names<-`(coef_names) %>%  
  {msummary(., 
            statistic = c("conf.int", "{std.error}", "{p.value}"), 
            estimate = "{estimate}{stars}", 
            coef_map = c("(Intercept)" = "(Intercept)",
                         "trade_weighted_growth" = "Trade Weighted Growth") )}  %>% 
    kable_styling(font_size = 11, bootstrap_options = c("hover", "condensed")) %>% print()
  
```
```{r}

glue(proj_path, "ols.rds") %>% 
  readRDS() %>% 
  spec_df(spec_list = ., mn = "ols", wm = "none") %>% 
  coef_point_plot()

```



## OLS - normalized trade and capacity weighted

```{r results='asis'}

cache_path = glue(proj_path, "ols_tradecap_weight.rds")
if (file.exists(cache_path)) {
  specif <- readRDS(cache_path)
} else {
  specif <- list()
  for (i in rf){
    df_reg <- rdf[, c(i, paste0(i, "_", twg))] %>% 
      `colnames<-`(c(i, twg))
    fm <- as.formula(paste(i, "~", twg))
    specif[[i]] <- lm(fm, data = df_reg, weights = rdf$trade_capacity)
  }
  saveRDS(specif, dp$util$mkdir(cache_path))
}
  specif %>%
  `names<-`(coef_names) %>%  
  {msummary(., 
            statistic = c("conf.int", "{std.error}", "{p.value}"), 
            estimate = "{estimate}{stars}", 
            coef_map = c("(Intercept)" = "(Intercept)",
                         "trade_weighted_growth" = "Trade Weighted Growth") )} %>% 
    kable_styling(font_size = 11, bootstrap_options = c("hover", "condensed")) %>% print()

  
```


```{r}

glue(proj_path, "ols_tradecap_weight.rds") %>% 
  readRDS() %>% 
  spec_df(spec_list = ., mn = "ols_weight", wm = "none") %>% 
  coef_point_plot()

```





## OLS - naive spatial lag
::: panel-tabset 
```{r results='asis'}

# Does growth in trade partner export destination(s) correspond with own growth?
# And what is the role of mere spatial diffusion?

# y_growth = b0 + b1_r + b2_d + e
# s.t. r = M * y_growth and  d = W * y_growth

cache_path = glue(proj_path, "ols_naive.rds")
if (file.exists(cache_path)) {
  specif <- readRDS(cache_path)
} else {
  specif <- list()  
  for (j in 1:length(W_list)){
    specif[[j]] <- list()
    for (i in rf){
      w <- {W_list[[j]] %*% as.matrix(rdf[i])}
      df_reg <- rdf[, c(i, paste0(i, "_", twg))] %>% 
        {cbind(., w)} %>% 
        as.data.frame() %>% 
        `colnames<-`(c(i, twg, "spatially_weighted_growth"))
      fm <- as.formula(paste(i, "~", twg, "+", "spatially_weighted_growth"))
      specif[[j]][[i]] <- lm(fm, data = df_reg)
      }
  }
  saveRDS(specif, dp$util$mkdir(cache_path))
}

for (j in 1:length(params$w_names)){
  cat('####', params$w_names[j], ' \n')
   specif[[j]] %>% 
      `names<-`(coef_names) %>%  
      {msummary(., 
                statistic = c("conf.int", "{std.error}", "{p.value}"), 
                estimate = "{estimate}{stars}", 
                coef_map = c("(Intercept)" = "(Intercept)",
                             "trade_weighted_growth" = "Trade Weighted Growth", 
                             "spatially_weighted_growth" = "Space Weighted Growth"))} %>% 
    kable_styling(font_size = 11, bootstrap_options = c("hover", "condensed")) %>% print()
   cat('\n')
}


```
::: 


```{r}

glue(proj_path, "ols_naive.rds") %>% 
  readRDS() %>% 
  spec_df(spec_list = ., mn = "ols_naive") %>% 
  coef_point_plot()

```


## Spatial Lag Model (max likelihood)
::: panel-tabset 
```{r results='asis'}

# Does growth in trade partner export destination(s) correspond with own growth?
# And is there a global diffusion of both observed and unobserved shocks? 


# Spatial Lag Model or Spatial Autoregressive Model
# y_growth = (I - lambda_W)^-1(b0 + b1_r + e)
# s.t. r = M * y_growth 

cache_path = glue(proj_path, "slm_ml.rds")
if (file.exists(cache_path)) {
  specif <- readRDS(cache_path)
} else {
  specif <- list()
  for (j in 1:length(W_list_alt)){
    W <- W_list_alt[[j]]
    specif[[j]] <- list()
    for (i in rf){
      df_reg <- rdf[, c(i, paste0(i, "_", twg))] %>% 
        `colnames<-`(c(i, twg))
      fm <- as.formula(paste(i, "~", twg))
      specif[[j]][[i]] <- lagsarlm(fm, data = df_reg, listw = W, zero.policy = TRUE)
      }
    }
    saveRDS(specif, dp$util$mkdir(cache_path))
  }

for (j in 1:length(params$w_names)){
  cat('####', params$w_names[j], ' \n')
  specif[[j]] %>% 
    `names<-`(coef_names) %>%  
    {msummary(., 
              statistic = c("conf.int", "{std.error}", "{p.value}"), 
              estimate = "{estimate}{stars}", 
              coef_map = c("(Intercept)" = "(Intercept)",
                           "trade_weighted_growth" = "Trade Weighted Growth",
                           "rho" = "lambda") )} %>% 
    kable_styling(font_size = 11, bootstrap_options = c("hover", "condensed")) %>% print()
  cat('\n')
}

```
:::

```{r}

glue(proj_path, "slm_ml.rds") %>% 
  readRDS() %>% 
  spec_df(spec_list = ., mn = "slm_ml") %>% 
  coef_point_plot()

```


## *Spatial Lag Model (GM/IV)
::: panel-tabset 
```{r results='asis'}

if(FALSE){
  cache_path = glue(proj_path, "slm_iv.rds")
  if (file.exists(cache_path)) {
    specif <- readRDS(cache_path)
  } else {
    specif <- list()
    for (j in 1:length(W_list)){
      W <- W_list[[j]]
      specif[[j]] <- list()
      for (i in rf){
        df_reg <- rdf[, c(i, paste0(i, "_", twg))] %>% 
          `colnames<-`(c(i, twg))
        fm <- as.formula(paste(i, "~", twg))
        specif[[j]][[i]] <- spreg(fm, data = df_reg, listw = W,  model = "lag")
      }
    }
    saveRDS(specif, dp$util$mkdir(cache_path))
  }
  
  sum_mod <- function(x){
    df <- summary(x)
    ti <- data.frame(
      term = df$CoefTable@Dimnames[[1]],
      estimate = df$CoefTable[, "Estimate"],
      std.error = df$CoefTable[, "Std. Error"],
      p.value = df$CoefTable[, "Pr(>|t|)"],
      conf.low = df$CoefTable[, "Estimate"] - df$CoefTable[, "Std. Error"] * 1.959964, 
      conf.high = df$CoefTable[, "Estimate"] + df$CoefTable[, "Std. Error"] * 1.959964 )
    gl <- data.frame(
      n = nrow(x$model))
    mod <- list(
      tidy = ti,
      glance = gl)
    class(mod) <- "modelsummary_list"
    return(mod)
  }
  
  for (j in 1:length(params$w_names)){
    cat('####', params$w_names[j], ' \n')
    specif[[j]] %>% 
      `names<-`(coef_names) %>%  
      {lapply(., sum_mod)} %>% 
      {msummary(., 
                statistic = c("conf.int", "{std.error}", "{p.value}"), 
                estimate = "{estimate}{stars}", 
                coef_map = c("(Intercept)" = "(Intercept)",
                             "trade_weighted_growth" = "Trade Weighted Growth",
                             "lambda" = "lambda") )} %>% 
    kable_styling(font_size = 11, bootstrap_options = c("hover", "condensed")) %>% print()
    cat('\n')
  }
}

```
::: 

## *Spatial Lag Model (2sls)

```{r}

if(FALSE){
  cache_path = glue(proj_path, "slm_2sls.rds")
  if (file.exists(cache_path)) {
    specif <- readRDS(cache_path)
  } else {
    specif <- list()
    for (j in 1:length(W_list)){
      W <- W_list_alt[[j]]
      specif[[j]] <- list()
      for (i in rf){
        df_reg <- rdf[, c(i, paste0(i, "_", twg))] %>% 
          `colnames<-`(c(i, twg))
        fm <- as.formula(paste(i, "~", twg))
        specif[[j]][[i]] <- stsls(fm, data = df_reg, listw = W, zero.policy = TRUE)
      }
    }
    saveRDS(specif, dp$util$mkdir(cache_path))
    }
}

```


## *Spatial Lag Model (GMM)

```{r}

if(FALSE){
  cache_path = glue(proj_path, "slm_gmm.rds")
  if (file.exists(cache_path)) {
    specif <- readRDS(cache_path)
  } else {
    specif <- list()
    for (j in 1:length(W_list)){
      W <- W_list_alt[[j]]
      specif[[j]] <- list()
      for (i in rf){
        df_reg <- rdf[, c(i, paste0(i, "_", twg))] %>% 
          `colnames<-`(c(i, twg))
        fm <- as.formula(paste(i, "~", twg))
        specif[[j]][[i]] <- GMerrorsar(fm, data = df_reg, listw = W, zero.policy = TRUE)
      }
    }
    saveRDS(specif, dp$util$mkdir(cache_path))
    }
}

```


## Spatial Error Model (max likelihood)
::: panel-tabset 
```{r results='asis'}

# Does growth in trade partner export destination(s) correspond with own growth?
# And is there a global diffusion of both observed and unobserved shocks? 

#In a spatial error model specification, the observations are related only due to unmeasured factors that, for some unknown reason, are correlated across the distances among the observations

# Spatial Error Model 
# y_growth = (I - rho_W)^-1(b0 + b1_r) + (I - rho_W)^-1e
# s.t. r = M * y_growth 

cache_path = glue(proj_path, "sem_ml.rds")
if (file.exists(cache_path)) {
  specif <- readRDS(cache_path)
} else {
  specif <- list()
  for (j in 1:length(W_list_alt)){
    W <- W_list_alt[[j]]
    specif[[j]] <- list()
    for (i in rf){
      df_reg <- rdf[, c(i, paste0(i, "_", twg))] %>% 
        `colnames<-`(c(i, twg))
      fm <- as.formula(paste(i, "~", twg))
      specif[[j]][[i]] <- errorsarlm(fm, data = df_reg, listw = W, zero.policy = TRUE)
      }
    }
    saveRDS(specif, dp$util$mkdir(cache_path))
  }

for (j in 1:length(params$w_names)){
  cat('####', params$w_names[j], ' \n')
  specif[[j]] %>% 
    `names<-`(coef_names) %>%  
    {msummary(., 
              statistic = c("conf.int", "{std.error}", "{p.value}"), 
              estimate = "{estimate}{stars}", 
              coef_map = c("(Intercept)" = "(Intercept)",
                           "trade_weighted_growth" = "Trade Weighted Growth",
                           "lambda" = "rho") )} %>% 
    kable_styling(font_size = 11, bootstrap_options = c("hover", "condensed")) %>% print()
  cat('\n')
}

```
::: 


```{r}

glue(proj_path, "sem_ml.rds") %>% 
  readRDS() %>% 
  spec_df(spec_list = ., mn = "sem_ml") %>% 
  coef_point_plot()

```


## Spatial Error Model (max likelihood) and capacity weighted
::: panel-tabset 
```{r results='asis'}

cache_path = glue(proj_path, "sem_tradecap_weight.rds")
if (file.exists(cache_path)) {
  specif <- readRDS(cache_path)
} else {
  specif <- list()
  for (j in 1:length(W_list_alt)){
    W <- W_list_alt[[j]]
    specif[[j]] <- list()
    for (i in rf){
      df_reg <- rdf[, c(i, paste0(i, "_", twg))] %>% 
        `colnames<-`(c(i, twg))
      fm <- as.formula(paste(i, "~", twg))
      specif[[j]][[i]] <- errorsarlm(fm, data = df_reg, listw = W, zero.policy = TRUE, weights = rdf$trade_capacity)
      }
    }
    saveRDS(specif, dp$util$mkdir(cache_path))
  }

for (j in 1:length(params$w_names)){
  cat('####', params$w_names[j], ' \n')
  specif[[j]] %>% 
    `names<-`(coef_names) %>%  
    {msummary(., 
              statistic = c("conf.int", "{std.error}", "{p.value}"), 
              estimate = "{estimate}{stars}", 
              coef_map = c("(Intercept)" = "(Intercept)",
                           "trade_weighted_growth" = "Trade Weighted Growth",
                           "lambda" = "rho") )} %>% 
    kable_styling(font_size = 11, bootstrap_options = c("hover", "condensed")) %>% print()
  cat('\n')
}

```
::: 

```{r}

glue(proj_path, "sem_tradecap_weight.rds") %>% 
  readRDS() %>% 
  spec_df(spec_list = ., mn = "sem_tradecap_weight") %>% 
  coef_point_plot()

```

## SARAR(1,1)
::: panel-tabset 
```{r results='asis'}

cache_path = glue(proj_path, "sarar.rds")
if (file.exists(cache_path)) {
  specif <- readRDS(cache_path)
} else {
  specif <- list()
  for (j in 1:length(W_list_alt)){
    W <- W_list_alt[[j]]
    specif[[j]] <- list()
    for (i in rf){
      df_reg <- rdf[, c(i, paste0(i, "_", twg))] %>% 
        `colnames<-`(c(i, twg))
      fm <- as.formula(paste(i, "~", twg))
      specif[[j]][[i]] <- sacsarlm(fm, data = df_reg, listw = W, zero.policy = TRUE)
      }
    }
    saveRDS(specif, dp$util$mkdir(cache_path))
  }

for (j in 1:length(params$w_names)){
  cat('####', params$w_names[j], ' \n')
  specif[[j]] %>% 
    `names<-`(coef_names) %>%  
    {msummary(., 
              statistic = c("conf.int", "{std.error}", "{p.value}"), 
              estimate = "{estimate}{stars}", 
              coef_map = c("(Intercept)" = "(Intercept)",
                           "trade_weighted_growth" = "Trade Weighted Growth",
                           "lambda" = "lambda",
                           "rho" = "rho") )} %>% 
    kable_styling(font_size = 11, bootstrap_options = c("hover", "condensed")) %>% print()
  cat('\n')
}

```
::: 


```{r}

glue(proj_path, "sarar.rds") %>% 
  readRDS() %>% 
  spec_df(spec_list = ., mn = "sarar") %>% 
  coef_point_plot()

```


```{r}

# return dummy factor levels (i.e.rural==1, nonrural==0) 
contrasts(factor(df$cbsa_rural_category))

```


## *Spatial Lag Model - rural dummy
::: panel-tabset 
```{r results='asis'}

# rural dummy

# Spatial Lag Model or Spatial Autoregressive Model
# y_growth = (I - lambda_W)^-1(b0 + b1_r + b2_d + e)
# s.t. r = M * y_growth 

if(FALSE){
  cache_path = glue(proj_path, "slm_rdum.rds")
  if (file.exists(cache_path)) {
    specif <- readRDS(cache_path)
  } else {
    specif <- list()
    for (j in 1:length(W_list_alt)){
      W <- W_list_alt[[j]]
      specif[[j]] <- list()
      for (i in rf){
        df_reg <- rdf[, c(i, paste0(i, "_", twg), "cbsa_rural_category")] %>% 
          `colnames<-`(c(i, twg, "cbsa_rural_category"))
        fm <- as.formula(paste(i, "~", twg, "+", "cbsa_rural_category"))
        specif[[j]][[i]] <- lagsarlm(fm, data = df_reg, listw = W, zero.policy = TRUE)
        }
      }
      saveRDS(specif, dp$util$mkdir(cache_path))
  }
  
  for (j in 1:length(params$w_names)){
    cat('####', params$w_names[j], ' \n')
    specif[[j]] %>% 
      `names<-`(coef_names) %>%  
      {msummary(., 
                statistic = c("conf.int", "{std.error}", "{p.value}"), 
                estimate = "{estimate}{stars}", 
                coef_map = c("(Intercept)" = "(Intercept)",
                             "trade_weighted_growth" = "Trade Weighted Growth",
                             "cbsa_rural_categoryrural" = "Rural Dummy",
                             "rho" = "lambda") )} %>% 
    kable_styling(font_size = 11, bootstrap_options = c("hover", "condensed")) %>% print()
    cat('\n')
  }
}

```
:::


## *Spatial Error Model - rural dummy
::: panel-tabset 
```{r results='asis'}

# rural dummy

# Spatial Error Model 
# y_growth = (I - rho_W)^-1(b0 + b1_r + b2_d) + (I - rho_W)^-1e
# s.t. r = M * y_growth 

if(FALSE){
  cache_path = glue(proj_path, "sem_rdum.rds")
  if (file.exists(cache_path)) {
    specif <- readRDS(cache_path)
  } else {
    specif <- list()
    for (j in 1:length(W_list_alt)){
      W <- W_list_alt[[j]]
      specif[[j]] <- list()
      for (i in rf){
        df_reg <- rdf[, c(i, paste0(i, "_", twg), "cbsa_rural_category")] %>% 
          `colnames<-`(c(i, twg, "cbsa_rural_category"))
        fm <- as.formula(paste(i, "~", twg, "+", "cbsa_rural_category"))
        specif[[j]][[i]] <- errorsarlm(fm, data = df_reg, listw = W, zero.policy = TRUE)
        }
      }
      saveRDS(specif, dp$util$mkdir(cache_path))
  }
  
  
  for (j in 1:length(params$w_names)){
    cat('####', params$w_names[j], ' \n')
    specif[[j]] %>% 
      `names<-`(coef_names) %>%  
      {msummary(., 
                statistic = c("conf.int", "{std.error}", "{p.value}"), 
                estimate = "{estimate}{stars}", 
                coef_map = c("(Intercept)" = "(Intercept)",
                             "trade_weighted_growth" = "Trade Weighted Growth",
                             "cbsa_rural_categoryrural" = "Rural Dummy",
                             "lambda" = "rho") )} %>% 
    kable_styling(font_size = 11, bootstrap_options = c("hover", "condensed")) %>% print()
    cat('\n')
  }
}

```
::: 


## Spatial Lag Model - rural dummy interact
::: panel-tabset 
```{r results='asis'}

# does growth in trade partner export destination(s) correspond with own growth?
# and is there a global diffusion of both observed and unobserved shocks? 

# rural dummy with interaction

# Spatial Lag Model or Spatial Autoregressive Model
# y_growth = (I - lambda_W)^-1(b0 + b1_r + b2_d  + b3_r * d + e)
# s.t. r = M * y_growth 


cache_path = glue(proj_path, "slm_rdumi.rds")
if (file.exists(cache_path)) {
  specif <- readRDS(cache_path)
} else {
  specif <- list()
  for (j in 1:length(W_list_alt)){
    W <- W_list_alt[[j]]
    specif[[j]] <- list()
    for (i in rf){
      df_reg <- rdf[, c(i, paste0(i, "_", twg), "cbsa_rural_category")] %>% 
        `colnames<-`(c(i, twg, "cbsa_rural_category"))
      fm <- as.formula(paste(i, "~", twg, "*", "cbsa_rural_category"))
      specif[[j]][[i]] <- lagsarlm(fm, data = df_reg, listw = W, zero.policy = TRUE)
      }
    }
    saveRDS(specif, dp$util$mkdir(cache_path))
}


for (j in 1:length(params$w_names)){
  cat('####', params$w_names[j], ' \n')
  specif[[j]] %>% 
    `names<-`(coef_names) %>%  
    {msummary(., 
              statistic = c("conf.int", "{std.error}", "{p.value}"), 
              estimate = "{estimate}{stars}", 
              coef_map = c("(Intercept)" = "(Intercept)",
                           "trade_weighted_growth" = "Trade Weighted Growth",
                           "cbsa_rural_categoryrural" = "Rural Dummy",
                           "trade_weighted_growth:cbsa_rural_categoryrural" = "Growth x Dummy",
                           "rho" = "lambda") )} %>% 
    kable_styling(font_size = 11, bootstrap_options = c("hover", "condensed")) %>% print()
  cat('\n')
}

```
::: 


## Spatial Error Model - rural dummy interact
::: panel-tabset 
```{r results='asis'}

# does growth in trade partner export destination(s) correspond with own growth?
# and is there a global diffusion of both observed and unobserved shocks? 

# rural dummy with interaction

# Spatial Error Model 
# y_growth = (I - rho_W)^-1(b0 + b1_r + b2_d + b3_r * _d) + (I - rho_W)^-1e
# s.t. r = M * y_growth 


cache_path = glue(proj_path, "sem_rdumi.rds")
if (file.exists(cache_path)) {
  specif <- readRDS(cache_path)
} else {
  specif <- list()
  for (j in 1:length(W_list_alt)){
    W <- W_list_alt[[j]]
    specif[[j]] <- list()
    for (i in rf){
      df_reg <- rdf[, c(i, paste0(i, "_", twg), "cbsa_rural_category")] %>% 
        `colnames<-`(c(i, twg, "cbsa_rural_category"))
      fm <- as.formula(paste(i, "~", twg, "*", "cbsa_rural_category"))
      specif[[j]][[i]] <- errorsarlm(fm, data = df_reg, listw = W, zero.policy = TRUE)
      }
    }
    saveRDS(specif, dp$util$mkdir(cache_path))
}


for (j in 1:length(params$w_names)){
  cat('####', params$w_names[j], ' \n')
  specif[[j]] %>% 
    `names<-`(coef_names) %>%  
    {msummary(., 
              statistic = c("conf.int", "{std.error}", "{p.value}"), 
              estimate = "{estimate}{stars}", 
              coef_map = c("(Intercept)" = "(Intercept)",
                           "trade_weighted_growth" = "Trade Weighted Growth",
                           "cbsa_rural_categoryrural" = "Rural Dummy",
                           "trade_weighted_growth:cbsa_rural_categoryrural" = "Growth x Dummy",
                           "lambda" = "rho") )} %>% 
    kable_styling(font_size = 11, bootstrap_options = c("hover", "condensed")) %>% print()
  cat('\n')
}

```
::: 


# regression - spatial regimes

```{r}

# specification testing 

if(FALSE){
  j = 5
  i = rf[1]
  
  df_reg <- rdf[, c(i, paste0(i, "_", twg), "cbsa_rural_category")] %>% 
          `colnames<-`(c(i, twg, "cbsa_rural_category")) %>% 
          mutate(cbsa_rural_category = ifelse(cbsa_rural_category == "rural", 1, 0))
  
  fm <- as.formula(paste(i, "~ 0 |", twg, "| 0 | 0 |", twg, "| 0"))
  spregimes(fm, data = df_reg, rgv = ~cbsa_rural_category,  listw = W_list_alt[[j]], wy_rg = F, weps_rg = F, model = "ols", verbose = T) %>% summary()
  spregimes(fm, data = df_reg, rgv = ~cbsa_rural_category,  listw = W_list_alt[[j]], wy_rg = F, weps_rg = F, model = "error", verbose = T) %>% summary()
  spregimes(fm, data = df_reg, rgv = ~cbsa_rural_category,  listw = W_list_alt[[j]], wy_rg = F, weps_rg = T, model = "error", verbose = T) %>% summary()
  
  
  
  fm <- as.formula(paste(i, "~ 0 |", twg, "-1 | 0 | 0 |", twg, "-1 | 0"))
  spregimes(fm, data = df_reg, rgv = ~cbsa_rural_category,  listw = W_list_alt[[j]], wy_rg = F, weps_rg = F, model = "lag", verbose = T) %>% summary()
  spregimes(fm, data = df_reg, rgv = ~cbsa_rural_category,  listw = W_list_alt[[j]], wy_rg = T, weps_rg = F, model = "lag", verbose = T) %>% summary()
  
  # fm <- exit_rate ~ 0 | trade_weighted_growth -1 | 1 | 0 | trade_weighted_growth -1 | 0
  # spregimes(fm, data = df_reg, rgv = ~cbsa_rural_category,  listw = W_list_alt[[j]], wy_rg = F, weps_rg = F, model = "lag", verbose = T) %>% summary()
  # spregimes(fm, data = df_reg, rgv = ~cbsa_rural_category,  listw = W_list_alt[[j]], wy_rg = T, weps_rg = F, model = "lag", verbose = T) %>% summary()
  
  # fm <- exit_rate ~ -1 | trade_weighted_growth -1 | 0 | -1 | trade_weighted_growth -1 | 0
  # spregimes(fm, data = df_reg, rgv = ~cbsa_rural_category,  listw = W_list_alt[[j]], wy_rg = F, weps_rg = F, model = "lag", verbose = T) %>% summary()
  # spregimes(fm, data = df_reg, rgv = ~cbsa_rural_category,  listw = W_list_alt[[j]], wy_rg = T, weps_rg = F, model = "lag", verbose = T) %>% summary()
  
  fm <- as.formula(paste(i, "~ 0 |", twg, "-1 | 0 | 1 |", twg, "-1 | 0"))
  spregimes(fm, data = df_reg, rgv = ~cbsa_rural_category,  listw = W_list_alt[[j]], wy_rg = F, weps_rg = F, model = "lag", verbose = T) %>% summary()
  spregimes(fm, data = df_reg, rgv = ~cbsa_rural_category,  listw = W_list_alt[[j]], wy_rg = T, weps_rg = F, model = "lag", verbose = T) %>% summary()
  
  
  
  fm <- as.formula(paste(i, "~ 1 |", twg, "-1 | 0 | 0 |", twg, "-1 | 0"))
  spregimes(fm, data = df_reg, rgv = ~cbsa_rural_category,  listw = W_list_alt[[j]], wy_rg = F, weps_rg = F, model = "lag", verbose = T) %>% summary()
  spregimes(fm, data = df_reg, rgv = ~cbsa_rural_category,  listw = W_list_alt[[j]], wy_rg = T, weps_rg = F, model = "lag", verbose = T) %>% summary()
  
  # fm <- exit_rate ~ 1 | trade_weighted_growth -1 | 1 | 0 | trade_weighted_growth -1 | 0
  # spregimes(fm, data = df_reg, rgv = ~cbsa_rural_category,  listw = W_list_alt[[j]], wy_rg = F, weps_rg = F, model = "lag", verbose = T) %>% summary()
  # spregimes(fm, data = df_reg, rgv = ~cbsa_rural_category,  listw = W_list_alt[[j]], wy_rg = T, weps_rg = F, model = "lag", verbose = T) %>% summary()
  
  fm <- as.formula(paste(i, "~ 1 |", twg, "-1 | 0 | 1 |", twg, "-1 | 0"))
  spregimes(fm, data = df_reg, rgv = ~cbsa_rural_category,  listw = W_list_alt[[j]], wy_rg = F, weps_rg = F, model = "lag", verbose = T) %>% summary()
  spregimes(fm, data = df_reg, rgv = ~cbsa_rural_category,  listw = W_list_alt[[j]], wy_rg = T, weps_rg = F, model = "lag", verbose = T) %>% summary()
  
  # fm <- exit_rate ~ 1 | trade_weighted_growth -1 | 1 | 1 | trade_weighted_growth -1 | 0
  # spregimes(fm, data = df_reg, rgv = ~cbsa_rural_category,  listw = W_list_alt[[j]], wy_rg = F, weps_rg = F, model = "lag", verbose = T) %>% summary()
  # spregimes(fm, data = df_reg, rgv = ~cbsa_rural_category,  listw = W_list_alt[[j]], wy_rg = T, weps_rg = F, model = "lag", verbose = T) %>% summary()
  
  
  # NFG
  fm <- exit_rate ~ -1 | trade_weighted_growth | 0 | -1 | trade_weighted_growth | 0
  spregimes(fm, data = df_reg, rgv = ~cbsa_rural_category,  listw = W_list_alt[[j]], wy_rg = F, weps_rg = F, model = "lag", verbose = T) %>% summary()
  spregimes(fm, data = df_reg, rgv = ~cbsa_rural_category,  listw = W_list_alt[[j]], wy_rg = T, weps_rg = F, model = "lag", verbose = T) %>% summary()
  
  fm <- exit_rate ~ 1 | trade_weighted_growth | 0 | 1 | trade_weighted_growth | 0
  spregimes(fm, data = df_reg, rgv = ~cbsa_rural_category,  listw = W_list_alt[[j]], wy_rg = F, weps_rg = F, model = "lag", verbose = T) %>% summary()
  spregimes(fm, data = df_reg, rgv = ~cbsa_rural_category,  listw = W_list_alt[[j]], wy_rg = T, weps_rg = F, model = "lag", verbose = T) %>% summary()
  
  fm <- exit_rate ~ -1 | trade_weighted_growth | 1 | -1 | trade_weighted_growth | 0
  spregimes(fm, data = df_reg, rgv = ~cbsa_rural_category,  listw = W_list_alt[[j]], wy_rg = F, weps_rg = F, model = "lag", verbose = T) %>% summary()
  spregimes(fm, data = df_reg, rgv = ~cbsa_rural_category,  listw = W_list_alt[[j]], wy_rg = T, weps_rg = F, model = "lag", verbose = T) %>% summary()
}

```



## OLS
::: panel-tabset 
```{r results='asis'}

cache_path = glue(proj_path, "regime_ols.rds")
if (file.exists(cache_path)) {
  specif <- readRDS(cache_path)
} else {
  specif <- list()
  for (j in 1:length(W_list_alt)){
    W <- W_list_alt[[j]]
    specif[[j]] <- list()
    for (i in rf){
      df_reg <- rdf[, c(i, paste0(i, "_", twg), "cbsa_rural_category")] %>% 
        `colnames<-`(c(i, twg, "cbsa_rural_category")) %>% 
        mutate(cbsa_rural_category = ifelse(cbsa_rural_category == "rural", 1, 0))
      fm <- as.formula(paste(i, "~ 0 |", twg, "| 0 | 0 |", twg, "| 0"))
      specif[[j]][[i]] <- spregimes(fm, data = df_reg, rgv = ~cbsa_rural_category, listw = W_list_alt[[j]], model = "ols")
      }
    }
    saveRDS(specif, dp$util$mkdir(cache_path))
}

sum_mod <- function(x){
  df <- summary(x)
  ti <- data.frame(
    term = names(df$CoefTable[, "Estimate"]),
    estimate = df$CoefTable[, "Estimate"],
    std.error = df$CoefTable[, "Std. Error"],
    p.value = df$CoefTable[, "Pr(>|z|)"],
    conf.low = df$CoefTable[, "Estimate"] - df$CoefTable[, "Std. Error"] * 1.959964, 
    conf.high = df$CoefTable[, "Estimate"] + df$CoefTable[, "Std. Error"] * 1.959964 )
  gl <- data.frame(
    n = length(x[[1]]$y))
  mod <- list(
    tidy = ti,
    glance = gl)
  class(mod) <- "modelsummary_list"
  return(mod)
}

for (j in 1:length(params$w_names)){
  cat('####', params$w_names[j], ' \n')
  specif[[j]] %>% 
    `names<-`(coef_names) %>%  
    {lapply(., sum_mod)} %>% 
    {msummary(., 
              statistic = c("conf.int", "{std.error}", "{p.value}"), 
              estimate = "{estimate}{stars}",
              coef_map = c("(Intercept)_1" = "(Intercept): non-rural",
                           "(Intercept)_2" = "(Intercept): rural",
                           "trade_weighted_growth_1" = "Trade Weighted Growth: non-rural",
                           "trade_weighted_growth_2" = "Trade Weighted Growth: rural") )} %>% 
    kable_styling(font_size = 11, bootstrap_options = c("hover", "condensed")) %>% print()
  cat('\n')
}

```
::: 


## SLM
::: panel-tabset 
```{r results='asis'}

cache_path = glue(proj_path, "regime_slm.rds")
if (file.exists(cache_path)) {
  specif <- readRDS(cache_path)
} else {
  specif <- list()
  for (j in 1:length(W_list_alt)){
    W <- W_list_alt[[j]]
    specif[[j]] <- list()
    for (i in rf){
      df_reg <- rdf[, c(i, paste0(i, "_", twg), "cbsa_rural_category")] %>% 
        `colnames<-`(c(i, twg, "cbsa_rural_category")) %>% 
        mutate(cbsa_rural_category = ifelse(cbsa_rural_category == "rural", 1, 0))
      fm <- as.formula(paste(i, "~ 1 |", twg, "-1 | 0 | 1 |", twg, "-1 | 0"))
      specif[[j]][[i]] <- spregimes(fm, data = df_reg, rgv = ~cbsa_rural_category,  listw = W_list_alt[[j]], wy_rg = T, weps_rg = F, model = "lag")
      }
    }
    saveRDS(specif, dp$util$mkdir(cache_path))
}

sum_mod <- function(x){
  df <- summary(x)
  ti <- data.frame(
    term = c("(Intercept)", "trade_weighted_growth_1", "trade_weighted_growth_2", "lambda1", "lambda2"),
    estimate = df$CoefTable[, "Estimate"],
    std.error = df$CoefTable[, "Std. Error"],
    p.value = df$CoefTable[, "Pr(>|z|)"],
    conf.low = df$CoefTable[, "Estimate"] - df$CoefTable[, "Std. Error"] * 1.959964, 
    conf.high = df$CoefTable[, "Estimate"] + df$CoefTable[, "Std. Error"] * 1.959964 )
  gl <- data.frame(
    n = length(x[[1]]$y))
  mod <- list(
    tidy = ti,
    glance = gl)
  class(mod) <- "modelsummary_list"
  return(mod)
}

for (j in 1:length(params$w_names)){
  cat('####', params$w_names[j], ' \n')
  specif[[j]] %>% 
    `names<-`(coef_names) %>%  
    {lapply(., sum_mod)} %>% 
    {msummary(., 
              statistic = c("conf.int", "{std.error}", "{p.value}"), 
              estimate = "{estimate}{stars}", 
             coef_map = c("(Intercept)" = "(Intercept)",
                          "trade_weighted_growth_1" = "Trade Weighted Growth: non-rural",
                          "trade_weighted_growth_2" = "Trade Weighted Growth: rural",
                          "lambda1" = "lambda: non-rural",
                          "lambda2" = "lambda rural") )} %>% 
    kable_styling(font_size = 11, bootstrap_options = c("hover", "condensed")) %>% print()
  cat('\n')
}

```
::: 


## SEM
::: panel-tabset 
```{r results='asis'}

cache_path = glue(proj_path, "regime_sem.rds")
if (file.exists(cache_path)) {
  specif <- readRDS(cache_path)
} else {
  specif <- list()
  for (j in 1:length(W_list_alt)){
    W <- W_list_alt[[j]]
    specif[[j]] <- list()
    for (i in rf){
      df_reg <- rdf[, c(i, paste0(i, "_", twg), "cbsa_rural_category")] %>% 
        `colnames<-`(c(i, twg, "cbsa_rural_category")) %>% 
        mutate(cbsa_rural_category = ifelse(cbsa_rural_category == "rural", 1, 0))
      fm <- as.formula(paste(i, "~ 0 |", twg, "| 0 | 0 |", twg, "| 0"))
      specif[[j]][[i]] <- spregimes(fm, data = df_reg, rgv = ~cbsa_rural_category, listw = W_list_alt[[j]], wy_rg = F, weps_rg = T, model = "error")
      }
    }
    saveRDS(specif, dp$util$mkdir(cache_path))
}

sum_mod <- function(x){
  df <- summary(x)
  ti <- data.frame(
    term = names(df$CoefTable[, "Estimate"]),
    estimate = df$CoefTable[, "Estimate"],
    std.error = df$CoefTable[, "Std. Error"],
    p.value = df$CoefTable[, "Pr(>|z|)"],
    conf.low = df$CoefTable[, "Estimate"] - df$CoefTable[, "Std. Error"] * 1.959964, 
    conf.high = df$CoefTable[, "Estimate"] + df$CoefTable[, "Std. Error"] * 1.959964 )
  gl <- data.frame(
    n = length(x[[1]]$y))
  mod <- list(
    tidy = ti,
    glance = gl)
  class(mod) <- "modelsummary_list"
  return(mod)
}

for (j in 1:length(params$w_names)){
  cat('####', params$w_names[j], ' \n')
  specif[[j]] %>% 
    `names<-`(coef_names) %>%  
    {lapply(., sum_mod)} %>% 
    {msummary(., 
              statistic = c("conf.int", "{std.error}", "{p.value}"), 
              estimate = "{estimate}{stars}",
              coef_map = c("(Intercept)_1" = "(Intercept): non-rural",
                           "(Intercept)_2" = "(Intercept): rural",
                           "trade_weighted_growth_1" = "Trade Weighted Growth: non-rural",
                           "trade_weighted_growth_2" = "Trade Weighted Growth: rural",
                           "We_1" = "rho: non-rural",
                           "We_2" = "rho: rural") )} %>% 
    kable_styling(font_size = 11, bootstrap_options = c("hover", "condensed")) %>% print()
  cat('\n')
}

```
::: 


# regression - rural only

## W edge list - rural

```{r}

cache_path = glue(proj_path, "W_list_alt_rural.rds")
  if (file.exists(cache_path)) {
    W_list_alt_rural <- readRDS(cache_path)
  } else {
    W_list_alt_rural <- list()
    W_list <- dp$imped_list(
      pn_dim = rural_places, 
      dmat = dp$call_dist_mat(year = params$year) )
    for (j in 1:length(W_list)){
      W_list_alt_rural[[j]] <- W_list[[j]] %>% 
        mat2listw(style="M", zero.policy=TRUE) 
    }
    saveRDS(W_list_alt_rural, dp$util$mkdir(cache_path))
  }

```


## OLS - normalized trade

```{r results='asis'}

# y_growth = b0 + b1_r + e
# s.t. r = M * y_growth

# Row normalized trade values
cache_path = glue(proj_path, "rural_ols.rds")
if (file.exists(cache_path)) {
  specif <- readRDS(cache_path)
} else {
  specif <- list()
  for (i in rf){
    df_reg <- rdf[rural_places, c(i, paste0(i, "_", twg))] %>% 
      `colnames<-`(c(i, twg))
    fm <- as.formula(paste(i, "~", twg))
    specif[[i]] <- lm(fm, data = df_reg)
  }
  saveRDS(specif, dp$util$mkdir(cache_path))
}

  specif %>%
  `names<-`(coef_names) %>%  
  {msummary(., 
            statistic = c("conf.int", "{std.error}", "{p.value}"), 
            estimate = "{estimate}{stars}", 
            coef_map = c("(Intercept)" = "(Intercept)",
                         "trade_weighted_growth" = "Trade Weighted Growth") )} %>% 
    kable_styling(font_size = 11, bootstrap_options = c("hover", "condensed")) %>% print()

  
```


## OLS - naive spatial lag
::: panel-tabset 
```{r results='asis'}

# y_growth = b0 + b1_r + b2_d + e
# s.t. r = M * y_growth and  d = W * y_growth

cache_path = glue(proj_path, "rural_ols_naive.rds")
if (file.exists(cache_path)) {
  specif <- readRDS(cache_path)
} else {
  specif <- list()  
  for (j in 1:length(W_list)){
    specif[[j]] <- list()
    for (i in rf){
      w <- {W_list[[j]][rural_places, rural_places] %*% as.matrix(rdf[rural_places, i])}
      df_reg <- rdf[rural_places, c(i, paste0(i, "_", twg))] %>% 
        {cbind(., w)} %>% 
        as.data.frame() %>% 
        `colnames<-`(c(i, twg, "spatially_weighted_growth"))
      fm <- as.formula(paste(i, "~", twg, "+", "spatially_weighted_growth"))
      specif[[j]][[i]] <- lm(fm, data = df_reg)
      }
  }
  saveRDS(specif, dp$util$mkdir(cache_path))
}

for (j in 1:length(params$w_names)){
  cat('####', params$w_names[j], ' \n')
   specif[[j]] %>% 
      `names<-`(coef_names) %>%  
      {msummary(., 
                statistic = c("conf.int", "{std.error}", "{p.value}"), 
                estimate = "{estimate}{stars}", 
                coef_map = c("(Intercept)" = "(Intercept)",
                             "trade_weighted_growth" = "Trade Weighted Growth", 
                             "spatially_weighted_growth" = "Space Weighted Growth") )} %>% 
    kable_styling(font_size = 11, bootstrap_options = c("hover", "condensed")) %>% print()
  cat('\n')
}

```
::: 


## Spatial Lag Model (max likelihood)
::: panel-tabset 
```{r results='asis'}

# Spatial Lag Model or Spatial Autoregressive Model
# y_growth = (I - lambda_W)^-1(b0 + b1_r + e)
# s.t. r = M * y_growth 

cache_path = glue(proj_path, "rural_slm_ml.rds")
if (file.exists(cache_path)) {
  specif <- readRDS(cache_path)
} else {
  specif <- list()
  for (j in 1:length(W_list_alt_rural)){
    W <- W_list_alt_rural[[j]]
    specif[[j]] <- list()
    for (i in rf){
      df_reg <- rdf[rural_places, c(i, paste0(i, "_", twg))] %>% 
        `colnames<-`(c(i, twg))
      fm <- as.formula(paste(i, "~", twg))
      specif[[j]][[i]] <- lagsarlm(fm, data = df_reg, listw = W, zero.policy = TRUE)
      }
    }
    saveRDS(specif, dp$util$mkdir(cache_path))
  }

for (j in 1:length(params$w_names)){
  cat('####', params$w_names[j], ' \n')
  specif[[j]] %>% 
    `names<-`(coef_names) %>%  
    {msummary(., 
              statistic = c("conf.int", "{std.error}", "{p.value}"), 
              estimate = "{estimate}{stars}", 
              coef_map = c("(Intercept)" = "(Intercept)",
                           "trade_weighted_growth" = "Trade Weighted Growth",
                           "rho" = "lambda") )} %>% 
    kable_styling(font_size = 11, bootstrap_options = c("hover", "condensed")) %>% print()
  cat('\n')
}

```
::: 


## Spatial Error Model (max likelihood)
::: panel-tabset 
```{r results='asis'}

# Spatial Error Model 
# y_growth = (I - rho_W)^-1(b0 + b1_r) + (I - rho_W)^-1e
# s.t. r = M * y_growth 

cache_path = glue(proj_path, "rural_sem_ml.rds")
if (file.exists(cache_path)) {
  specif <- readRDS(cache_path)
} else {
  specif <- list()
  for (j in 1:length(W_list_alt_rural)){
    W <- W_list_alt_rural[[j]]
    specif[[j]] <- list()
    for (i in rf){
      df_reg <- rdf[rural_places, c(i, paste0(i, "_", twg))] %>% 
        `colnames<-`(c(i, twg))
      fm <- as.formula(paste(i, "~", twg))
      specif[[j]][[i]] <- errorsarlm(fm, data = df_reg, listw = W, zero.policy = TRUE)
      }
    }
    saveRDS(specif, dp$util$mkdir(cache_path))
  }

for (j in 1:length(params$w_names)){
  cat('####', params$w_names[j], ' \n')
  specif[[j]] %>% 
    `names<-`(coef_names) %>%  
    {msummary(., 
              statistic = c("conf.int", "{std.error}", "{p.value}"), 
              estimate = "{estimate}{stars}", 
              coef_map = c("(Intercept)" = "(Intercept)",
                           "trade_weighted_growth" = "Trade Weighted Growth",
                           "lambda" = "rho") )} %>% 
    kable_styling(font_size = 11, bootstrap_options = c("hover", "condensed")) %>% print()
  cat('\n')
}

```
::: 


# regression - nonrural only


## W edge list - nonrural
```{r}

cache_path = glue(proj_path, "W_list_alt_nonrural.rds")
  if (file.exists(cache_path)) {
    W_list_alt_nonrural <- readRDS(cache_path)
  } else {
    W_list_alt_nonrural <- list()
    W_list <- dp$imped_list(
      pn_dim = nonrural_places, 
      dmat = dp$call_dist_mat(year = params$year) )
    for (j in 1:length(W_list)){
      W_list_alt_nonrural[[j]] <- W_list[[j]] %>% 
        mat2listw(style="M", zero.policy=TRUE) 
    }
    saveRDS(W_list_alt_nonrural, dp$util$mkdir(cache_path))
  }

```


## OLS - normalized trade

```{r results='asis'}

# y_growth = b0 + b1_r + e
# s.t. r = M * y_growth

# Row normalized trade values
cache_path = glue(proj_path, "nonrural_ols.rds")
if (file.exists(cache_path)) {
  specif <- readRDS(cache_path)
} else {
  specif <- list()
  for (i in rf){
    df_reg <- rdf[nonrural_places, c(i, paste0(i, "_", twg))] %>% 
      `colnames<-`(c(i, twg))
    fm <- as.formula(paste(i, "~", twg))
    specif[[i]] <- lm(fm, data = df_reg)
  }
  saveRDS(specif, dp$util$mkdir(cache_path))
}

  specif %>%
  `names<-`(coef_names) %>%  
  {msummary(., 
            statistic = c("conf.int", "{std.error}", "{p.value}"), 
            estimate = "{estimate}{stars}", 
            coef_map = c("(Intercept)" = "(Intercept)",
                         "trade_weighted_growth" = "Trade Weighted Growth") )} %>% 
    kable_styling(font_size = 11, bootstrap_options = c("hover", "condensed")) %>% print()

  
```



## OLS - naive spatial lag
::: panel-tabset 
```{r results='asis'}

# y_growth = b0 + b1_r + b2_d + e
# s.t. r = M * y_growth and  d = W * y_growth

cache_path = glue(proj_path, "nonrural_ols_naive.rds")
if (file.exists(cache_path)) {
  specif <- readRDS(cache_path)
} else {
  specif <- list()  
  for (j in 1:length(W_list)){
    specif[[j]] <- list()
    for (i in rf){
      w <- {W_list[[j]][nonrural_places, nonrural_places] %*% as.matrix(rdf[nonrural_places, i])}
      df_reg <- rdf[nonrural_places, c(i, paste0(i, "_", twg))] %>% 
        {cbind(., w)} %>% 
        as.data.frame() %>% 
        `colnames<-`(c(i, twg, "spatially_weighted_growth"))
      fm <- as.formula(paste(i, "~", twg, "+", "spatially_weighted_growth"))
      specif[[j]][[i]] <- lm(fm, data = df_reg)
      }
  }
  saveRDS(specif, dp$util$mkdir(cache_path))
}

for (j in 1:length(params$w_names)){
  cat('####', params$w_names[j], ' \n')
   specif[[j]] %>% 
      `names<-`(coef_names) %>%  
      {msummary(., 
                statistic = c("conf.int", "{std.error}", "{p.value}"), 
                estimate = "{estimate}{stars}", 
                coef_map = c("(Intercept)" = "(Intercept)",
                             "trade_weighted_growth" = "Trade Weighted Growth", 
                             "spatially_weighted_growth" = "Space Weighted Growth") )} %>% 
    kable_styling(font_size = 11, bootstrap_options = c("hover", "condensed")) %>% print()
  cat('\n')
}


```
::: 


## Spatial Lag Model (max likelihood)
::: panel-tabset 
```{r results='asis'}

# Spatial Lag Model or Spatial Autoregressive Model
# y_growth = (I - lambda_W)^-1(b0 + b1_r + e)
# s.t. r = M * y_growth 

cache_path = glue(proj_path, "nonrural_slm_ml.rds")
if (file.exists(cache_path)) {
  specif <- readRDS(cache_path)
} else {
  specif <- list()
  for (j in 1:length(W_list_alt_nonrural)){
    W <- W_list_alt_nonrural[[j]]
    specif[[j]] <- list()
    for (i in rf){
      df_reg <- rdf[nonrural_places, c(i, paste0(i, "_", twg))] %>% 
        `colnames<-`(c(i, twg))
      fm <- as.formula(paste(i, "~", twg))
      specif[[j]][[i]] <- lagsarlm(fm, data = df_reg, listw = W, zero.policy = TRUE)
      }
    }
    saveRDS(specif, dp$util$mkdir(cache_path))
  }

for (j in 1:length(params$w_names)){
  cat('####', params$w_names[j], ' \n')
  specif[[j]] %>% 
    `names<-`(coef_names) %>%  
    {msummary(., 
              statistic = c("conf.int", "{std.error}", "{p.value}"), 
              estimate = "{estimate}{stars}", 
              coef_map = c("(Intercept)" = "(Intercept)",
                           "trade_weighted_growth" = "Trade Weighted Growth",
                           "rho" = "lambda") )} %>% 
    kable_styling(font_size = 11, bootstrap_options = c("hover", "condensed")) %>% print()
  cat('\n')
}

```
::: 


## Spatial Error Model (max likelihood)
::: panel-tabset 
```{r results='asis'}

# Spatial Error Model 
# y_growth = (I - rho_W)^-1(b0 + b1_r) + (I - rho_W)^-1e
# s.t. r = M * y_growth 

cache_path = glue(proj_path, "nonrural_sem_ml.rds")
if (file.exists(cache_path)) {
  specif <- readRDS(cache_path)
} else {
  specif <- list()
  for (j in 1:length(W_list_alt_nonrural)){
    W <- W_list_alt_nonrural[[j]]
    specif[[j]] <- list()
    for (i in rf){
      df_reg <- rdf[nonrural_places, c(i, paste0(i, "_", twg))] %>% 
        `colnames<-`(c(i, twg))
      fm <- as.formula(paste(i, "~", twg))
      specif[[j]][[i]] <- errorsarlm(fm, data = df_reg, listw = W, zero.policy = TRUE)
      }
    }
    saveRDS(specif, dp$util$mkdir(cache_path))
  }

for (j in 1:length(params$w_names)){
  cat('####', params$w_names[j], ' \n')
  specif[[j]] %>% 
    `names<-`(coef_names) %>%  
    {msummary(., 
              statistic = c("conf.int", "{std.error}", "{p.value}"), 
              estimate = "{estimate}{stars}", 
              coef_map = c("(Intercept)" = "(Intercept)",
                           "trade_weighted_growth" = "Trade Weighted Growth",
                           "lambda" = "rho") )} %>% 
    kable_styling(font_size = 11, bootstrap_options = c("hover", "condensed")) %>% print()
  cat('\n')
}

```
:::


# In progress

## Moran's I
```{r}

if(FALSE){
  #   wl <- intersect(rownames(W), pn) %>% {W[., .]} %>% mat2listw() 
  # #lagsarlm(i ~ trade_weighted_growth, data=df_reg, listw = wl, method="eigen", quiet=FALSE)
  # lagsarlm(population_gr ~ trade_weighted_growth, data=df_reg, listw = W_list_alt[[j]], method="eigen", quiet=FALSE, zero.policy=TRUE)
  #  moran.test(x = df_reg$population_gr, listw = W_list_alt[[j]], zero.policy = T)
  #  moran.plot(x = df_reg$population_gr, listw = W_list_alt[[j]], zero.policy = T)
  # 
  # moran.test(x = df_reg$population_gr, listw = wl) 
  #  moran.plot(x = df_reg$population_gr, listw = wl)

for (j in 1:length(W_list_alt)){
      for (i in rf){
        moran.test(x = df[df$place %in% pn, ][[i]], listw = W_list_alt[[j]], zero.policy = T) %>% print()
      }
}

mspec <- list()
      for (i in rf){
        mspec[[i]] <- moran.test(x = df[df$place %in% pn, ][[i]], listw = W_list_alt[[j]], zero.policy = T) %>% tidy()
        print(mspec[[i]])
      }

mspec <- data.frame()
      for (i in rf){
        mspec <- moran.test(x = rdf[[i]], listw = W_list_alt[[j]], zero.policy = T) %>% tidy() %>% rbind(mspec,.)
      }
mspec

#Morans I has theoretical limits from -1 for perfect dispersion or repulsion, to +1 for perfect clustering, with 0 indicting a perfectly random pattern.
}
```


```{r}

spec_mods <- c(
  "ols.rds",
  "ols_tradecap_weight.rds",
  "ols_naive.rds",
  
  "slm_ml.rds",
  # "slm_iv.rds",
  
  "sem_ml.rds",
  "sem_tradecap_weight.rds",
  
  "sarar.rds",
  
  "slm_rdumi.rds",
  "sem_rdumi.rds",

  # "regime_ols.rds",
  # "regime_slm.rds",
  # "regime_sem.rds",
  
  "rural_ols.rds",
  "rural_ols_naive.rds",
  "rural_slm_ml.rds", 
  "rural_sem_ml.rds",
  
  "nonrural_ols.rds",
  "nonrural_ols_naive.rds",
  "nonrural_slm_ml.rds",
  "nonrural_sem_ml.rds"
  ) 

estdf <- data.frame()
for(i in spec_mods){
    x <- glue(proj_path, i) %>% 
      readRDS() %>% 
      spec_df(spec_list = ., mn = gsub(".rds*", "", i), wm = if(!is.null(names(.))){"none"} )
    estdf <- rbind(estdf, x)
}

estdf %>% 
  filter(model %in% c("ols", "ols_naive", "slm_ml", "sem_ml")) %>% 
  #filter(wmatrix %in% c("none", "inverse", "qnn", "knn", "proximity")) %>% 
  coef_point_plot(error = F)

```









